### 2024-11-30 paper/paper.qmd
Updated models sections to the paper

---
title: "When Risk Hits the Road: Does Danger Drive a Return?"
subtitle: "Motor Theft and Collisions in Toronto Are Driven by [Neighborhood Clustering] and [Environmental Conditions]"
author: 
  - Yingke He
thanks: "Code and data are available at: [https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main](https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main)."
date: today
date-format: long
abstract: "This study examines motor vehicle theft and traffic collisions in Toronto, focusing on spatial and temporal patterns, recovery outcomes, and contributing factors. [The analysis reveals that thefts are geographically clustered, with recovery rates differing by location type, while collisions are more frequent under poor visibility and adverse road conditions.] These findings underscore the need for targeted interventions, including enhanced surveillance in high-theft areas, infrastructure improvements for road safety, and optimized strategies for vehicle recovery. Such measures are critical for fostering safer and more secure urban mobility in Toronto."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction {#sec-intro}

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Source

## Data Measurement and Limitations

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome Variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}

```

Talk way more about it.

## Predictor Variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model {#sec-model}

The main idea of this composite risk score model is to calculate an index of risk for owning and riding a motorbike. The modeling strategy has two primary objectives. The first objective is to estimate the likelihood of theft \( P(\text{Theft}) \) and collision \( P(\text{Collision}) \) under diverse environmental and situational conditions using logistic regression models. The second objective is to combine these probabilities into a unified Risk Index, providing actionable insights into the risk for owning and riding a motorbike

The models were run in R [@citeR] using the `stats` package to implement logistic regression models predicting theft and collision probabilities. The theft model incorporates predictor variables such as neighborhood ID, premises type, report hour, location type, and report day, while the collision model includes neighborhood ID, road surface condition, lighting condition, and traffic control.

Both models estimate the log-odds of their respective probabilities using the predictors,  enabling reliable predictions under varying environmental and situational conditions. Model diagnostics and background details, including variable descriptions and performance metrics, are included in [Appendix @sec-model-details].

## Model Set-up

The logistic regression model for theft probability takes the form of the following equation:

\begin{align*}
\log\left(\frac{P(\text{Theft})}{1 - P(\text{Theft})}\right) &= \beta_0 + \beta_1 \cdot \text{HOOD}_{158} + \beta_2 \cdot \text{Premises Type} + \beta_3 \cdot \text{Report Hour} \\
&\quad + \beta_4 \cdot \text{Location Type} + \beta_5 \cdot \text{Report Day} + \epsilon
\end{align*}

This model utilizes the following predictor variables:

- **Neighborhood ID** (`hood_158`): Unique identifier for the neighborhood.
- **Premises Type** (`premises_type`): Type of premises where the theft occurred like parking lot, garage
- **Report Hour** (`REPORT_HOUR`):  Hour of the day the theft was reported.
- **Location Type** (`location`):Location of the theft like indoors and outdoors.
- **Report Day** (`REPORT_DOW`): Day of the Week Offence was Reported.

The logistic regression model for collision probability takes the form of the following equation:

\begin{align*}
\log\left(\frac{P(\text{Collision})}{1 - P(\text{Collision})}\right) &= \beta_0 + \beta_1 \cdot \text{HOOD}_{158} + \beta_2 \cdot \text{Road Surface Condition} + \beta_3 \cdot \text{Light} \\
&\quad + \beta_4 \cdot \text{Traffic Control} + \epsilon
\end{align*}

This model utilizes the following predictor variables:

- **Neighborhood ID** (`hood_158`): Unique identifier for the neighborhood.
- **Road Surface Condition** (`road_conditions`): Road surface condition like wet and dry.
- **Lighting Condition** (`lighting_conditions`): Lighting conditions at the time of the collision like dark or clear.
- **Traffic Control** (`traffic_control`):Type of traffic control present like stop signs or traffic signals
After fitting both models, combine the probabilities into a weighted risk index:

\begin{align*}
\text{Risk Index} &= w_1 \cdot P(\text{Theft}) + w_2 \cdot P(\text{Collision})
\end{align*}

\begin{align*}
\text{where weights } w_1 \text{and } w_2 \text{ are adjust based on the relative importance of theft and collision risks.}
\end{align*}

## Model justification

Logistic regression was selected for this analysis due to its simplicity and effectiveness in estimating probabilities for binary outcomes, such as theft and collision. This statistical method is well-suited for modeling the log-odds of an event's occurrence as a linear combination of predictor variables, providing a clear framework for understanding the relationships between independent variables and the dependent outcome. Its ability to estimate odds ratios allows for an intuitive interpretation of the impact of predictor variables, making it particularly effective for identifying and quantifying risks in diverse scenarios.

This model's strengths include its flexibility in handling both categorical and continuous predictors, enabling the integration of a wide range of variables such as neighborhood characteristics, premises type, lighting conditions, road surface conditions, time of day, and environmental conditions. Additionally, logistic regression’s ability to include interaction terms and manage potential multicollinearity enhances its effectiveness in capturing complex relationships between predictors and the likelihood of thefts and collisions. Utilizing these features enables the model to deliver a thorough evaluation of the factors affecting theft and collision probabilities, addressing the first objectives of the modelling strategy.

After estimating the probabilities of theft and collision, these probabilities are combined into a composite Risk Index using a weighted formula. This Risk Index offers a unified measure of risk, allowing for comparative evaluations of the relative safety of motorbike ownership and usage across varying contexts.

The modeling approach is justified by logistic regression’s ability to deliver clear, interpretable, and statistically reliable risk estimates. Its statistical efficiency, with its capacity to manage complex, large datasets, makes it an ideal choice for accurately predicting theft and collision risk, while the composite Risk Index facilitates a clearer understanding of the risks associated with motorbike ownership and usage.

## Model Assumtions and Validations

To validate the use of logistic regression models, the four main assumptions: binary nature of the outcome, linearity of the logit, indipendence of observations, and lack of multicollinearity will be assessed [@kononen2011logistic]. To validate the use of a composite Risk Index, the correlation of the index with observed outcomes will be assess using a Pearson correlation test.

### Logistic Regression Models

#### Binary Nature of the Outcome

A fundamental assumption of logistic regression is that the response variable is binary or dichotomous, meaning it can take on only two possible outcomes [@nick2007logistic]. This assumption is satisfied in both models. In the theft model, the response variable indicates whether a theft occurred, with a value of 1 representing the occurrence of theft and 0 representing its absence. Similarly, in the collision model, the response variable denotes whether a collision occurred, where 1 represents the occurrence of a collision and 0 indicates no collision.

The logistic regression model does not directly predict 1 or 0. Instead, it estimates the probability that the response variable equals 1, given the predictor variables. For instance, the theft model outputs the probability of theft occurring (P(Theft)), and the collision model outputs the probability of a collision occurring (P(Collision)). These probabilities range between 0 and 1, allowing for detailed predictions of the likelihood of the respective events.

#### Independence of Observations

A second assumption of logistic regression is the independence of observations, meaning that all outcomes within the sample must be separate from one another [@stoltzfus2011logistic]. This implies that the data should not contain repeated measurements or outcomes that are inherently correlated.

For this analysis, the data has been checked to ensure that it contains no duplicate measures, with each observation corresponding to a distinct event, such as a theft or collision.

However, it is acknowledged that the data includes groupings by neighborhood, as indicated by the variable Neighborhood ID (`hood_158`). Neighborhood groupings may introduce some correlation due to shared environmental or socio-demographic factors. Despite this, the variable Neighborhood ID is included in the models as a predictor to account for such shared characteristics explicitly. Additionally, the other predictors, such as Road Surface Condition, Lighting Condition, and Traffic Control for the collision model, and Premises Type, Report Hour, Location Type, and Report Day for the theft model, represent independent situational or environmental variables tied to specific events.

Given the heterogeneity across neighborhoods and the distinct nature of the predictors, any residual correlation within groupings is expected to be minimal and not comparable to the stronger dependency seen in matched data, such as multiple observations from the same household or individual. Accordingly, it is believed that the independence assumption of logistic regression is sufficiently satisfied for this analysis.

#### Linear Relationship in the Log-Odds

## ```{r}
# Load necessary libraries
library(dplyr)
library(here)

dataset <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))

# Fit the logistic regression model for theft
model <- glm(outcome ~ REPORT_HOUR + premises_type + location + REPORT_DOW, 
             family = binomial, data = dataset)

# Create bins for REPORT_HOUR
dataset$hour_bin <- cut(dataset$REPORT_HOUR, breaks = 10)  # Divide into 10 bins

# Calculate mean REPORT_HOUR and predicted probabilities for each bin
bin_summary <- dataset %>%
  group_by(hour_bin) %>%
  summarise(mean_hour = mean(REPORT_HOUR, na.rm = TRUE),
            mean_prob = mean(predict(model, type = "response"), na.rm = TRUE))

# Plot the relationship between mean REPORT_HOUR and predicted probabilities
plot(bin_summary$mean_hour, bin_summary$mean_prob, 
     xlab = "Mean Report Hour (Binned)", 
     ylab = "Mean Predicted Probability", 
     main = "Linearity Check for REPORT_HOUR",
     pch = 20, col = "blue")
lines(lowess(bin_summary$mean_hour, bin_summary$mean_prob), col = "red", lwd = 2)

```

#### Absence of Multicollinearity

### Composite Risk Index

#### Pearson Correlation Test

# Results {#sec-result}

Our results are summarized in @tbl-modelresults.

```{r}

```

```{r}

```

## First Result Point

## Second Result Point

## Third Result Point

# Discussion {#sec-discussion}

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}

```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}

```

\newpage

# References

### 2024-11-30 paper/paper.qmd
updated the introduction section

@@
 
-Overview paragraph
+Decisions surrounding motorbike ownership and usage carry significant implications for personal safety and financial liability.
+Recent statistics highlight the increased risks associated with owning and riding motorbikes, including a heightened likelihood of theft and collisions compared to other vehicles [@Yasmin2016].
+These risks are influenced by various factors such as geographic location, road conditions, time of day, and type of motorbike, making it essential to develop tools that effectively assess and mitigate these dangers.[can add one more sentence + citations]
 
-Estimand paragraph
+This study introduces a composite risk score model to assess the risks associated with owning and riding a motorbike.
+Using data from Open Data Toronto, the model evaluates two critical events: motorbike theft and collisions.
+Logistic regression is employed to estimate the probabilities of these events, which are then combined into a single, interpretable composite risk score.
+This metric, incorporating factors such as neighborhood characteristics, road and lighting conditions, and time of day, is designed to guide motorbike users, insurers, and policymakers in risk assessment and decision-making, while informing strategies to mitigate these risks.
 
-Results paragraph
+The primary estimand of the analysis is the composite risk score, derived from the individual probabilities of motorbike theft and collision.
+This score is calculated using predictor variables such as neighborhood characteristics, road and lighting conditions, time of day, and other contextual factors, which were selected for their documented relevance to motorbike-related risks.
 
-Why it matters paragraph
+This analysis confirms and extends three key findings: (1) material properties significantly influence motor efficiency, with high-conductivity materials yielding higher energy savings; (2) geometric optimization of motor components enhances torque generation and reduces energy losses; and (3) operational conditions, such as load and temperature, exhibit nonlinear effects, underscoring the need for adaptive motor control strategies.
+Furthermore, our projections suggest that integrating advanced materials and control systems could improve motor efficiency by up to 15% by 2050, aligning with global energy transition goals [To be updated..].
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The structure of the paper is organized as follows: following @sec-intro, @sec-data outlines the data collection and preprocessing procedures, along with a detailed description of the outcome variable and the predictor variables used in the analysis.
+@sec-model introduces the logistic regression models applied to estimate the probabilities of theft and collision, as well as the method used to combine these probabilities into a composite risk score.
+@sec-result then presents the main findings, including insights into how different factors contribute to the risks of owning and riding a motorbike.
+Finally, @sec-discussion interprets the results, highlighting significant trends and implications for motorbike risk assessment, and concludes with a discussion on the limitations of the analysis and future research directions.
 
@@
 
-## Overview
-
 We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
 
-Overview text
+[Libraries To be updated...]
+
+Details about the data cleaning process and the criteria for variable selection are provided in Appendix A.
 
@@
 
+This study utilized two datasets published by the Toronto Police Service.
+The first dataset focuses on motor vehicle collisions involving killed or seriously injured persons (KSI), while the second examines thefts from motor vehicles.
+
+The Motor Vehicle Collisions dataset includes all reported incidents in which a person was either killed or seriously injured since 2006.
+It offers detailed information about each collision, such as the type of incident, the severity of injuries, and the location of the event, when available.
+Additionally, the dataset includes fields for both the old 140 and new 158 neighborhood structures in Toronto, allowing for flexible neighborhood-level analysis across different definitions.
+
+The Theft from Motor Vehicle dataset contains all reported occurrences of thefts from vehicles, categorized by reported date.
+These offences are classified based on the value of the stolen items, distinguishing between theft under and theft over thresholds.
+Each occurrence number may include multiple rows, representing the various offences associated with a single event.
+The dataset excludes "unfounded" occurrences, adhering to Statistics Canada’s definition that these events were determined not to have occurred or been attempted.
+Like the KSI dataset, this dataset includes fields for both the old and new neighborhood structures, enabling comprehensive geographic analyses of theft trends.
+
+[Add variable types and trends]
+
 ## Data Measurement and Limitations
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The process of translating real-world phenomena into entries in the dataset involves several stages.
+When a traffic collision or theft occurs, it is reported to law enforcement through various channels, such as emergency calls, online submissions, or in-person reports.
+Police officers or administrative personnel document the event details, including date, location, type of incident, and additional attributes such as severity or value of stolen items.
+These records are then digitized and aggregated into structured datasets, with fields organized to support analysis and reporting.
+However, during this process, certain changes or context-specific information may be lost, and the data ultimately reflects a structured summary of the events rather than their full complexity.
+
+The datasets from Open Data Toronto did not specify the exact methods used for data collection, which may introduce some uncertainty regarding the consistency and reliability of the recorded events.
+Additionally, for privacy reasons, the locations of crime occurrences have been deliberately offset to the nearest road intersection node.
+This may result in discrepancies when analyzing counts by division or neighborhood, as the reported locations may not reflect the exact sites of the occurrences.
+
+Some coordinate information in the datasets appears as “0, 0,” indicating that the specific location was either not validated or could not be geocoded.
+In such cases, a general division or neighborhood association may still be provided, but for invalid or external locations, the designation “NSA” (“Not Specified Area”) is used.
+Furthermore, the Toronto Police Service does not guarantee the accuracy, completeness, or timeliness of the data, which may lead to potential misinterpretations or incomplete analyses.
+
+Additional details about the dataset are available in the datasheet, accessible through the repository linked to this paper.
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The outcome variable of this analysis is the Risk Index, a composite metric that integrates the probabilities of two underlying events: theft and collision.
 
@@
 ```
-
-Talk more about it.
 
@@
 
-Talk way more about it.
-
 ## Predictor Variables
 
-Add graphs, tables and text.
+### Collision Probability
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+#### Neighborhood ID
+#### Premises Type
+#### Report Hour 
+#### Location Type
+#### Report Day
+
+### Theft Probability
+
+#### Neighborhood ID
+#### Road Surface Condition
+#### Lighting Condition
+#### Traffic Control
 
@@
 
-#### Independence of Observations
-
-A second assumption of logistic regression is the independence of observations, meaning that all outcomes within the sample must be separate from one another [@stoltzfus2011logistic].
-This implies that the data should not contain repeated measurements or outcomes that are inherently correlated.
-
-For this analysis, the data has been checked to ensure that it contains no duplicate measures, with each observation corresponding to a distinct event, such as a theft or collision.
-
-However, it is acknowledged that the data includes groupings by neighborhood, as indicated by the variable Neighborhood ID (`hood_158`).
-Neighborhood groupings may introduce some correlation due to shared environmental or socio-demographic factors.
-Despite this, the variable Neighborhood ID is included in the models as a predictor to account for such shared characteristics explicitly.
-Additionally, the other predictors, such as Road Surface Condition, Lighting Condition, and Traffic Control for the collision model, and Premises Type, Report Hour, Location Type, and Report Day for the theft model, represent independent situational or environmental variables tied to specific events.
-
-Given the heterogeneity across neighborhoods and the distinct nature of the predictors, any residual correlation within groupings is expected to be minimal and not comparable to the stronger dependency seen in matched data, such as multiple observations from the same household or individual.
-Accordingly, it is believed that the independence assumption of logistic regression is sufficiently satisfied for this analysis.
-
 #### Linear Relationship in the Log-Odds
 
-## ```{r}
-# Load necessary libraries
-library(dplyr)
-library(here)
+[Identifying Continuous Variables in Your Models:
+For LR Model 1 (Theft Model):
+Continuous Variable:
+REPORT_HOUR (Hour of the day the theft was reported
+For LR Model 2 (Collision Model):
+No Continuous Variables]
 
-dataset <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
-
-# Fit the logistic regression model for theft
-model <- glm(outcome ~ REPORT_HOUR + premises_type + location + REPORT_DOW, 
-             family = binomial, data = dataset)
-
-# Create bins for REPORT_HOUR
-dataset$hour_bin <- cut(dataset$REPORT_HOUR, breaks = 10)  # Divide into 10 bins
-
-# Calculate mean REPORT_HOUR and predicted probabilities for each bin
-bin_summary <- dataset %>%
-  group_by(hour_bin) %>%
-  summarise(mean_hour = mean(REPORT_HOUR, na.rm = TRUE),
-            mean_prob = mean(predict(model, type = "response"), na.rm = TRUE))
-
-# Plot the relationship between mean REPORT_HOUR and predicted probabilities
-plot(bin_summary$mean_hour, bin_summary$mean_prob, 
-     xlab = "Mean Report Hour (Binned)", 
-     ylab = "Mean Predicted Probability", 
-     main = "Linearity Check for REPORT_HOUR",
-     pch = 20, col = "blue")
-lines(lowess(bin_summary$mean_hour, bin_summary$mean_prob), col = "red", lwd = 2)
-
-```
+[Conclusion:
+Total Number of Plots: 1
+Draw a binning plot for REPORT_HOUR in LR Model 1 to assess the linear relationship]
 
 #### Absence of Multicollinearity
+
+[The easiest way to check for multicollinearity in your logistic regression models is to calculate the Variance Inflation Factor (VIF) for your predictors.
+This provides a straightforward and interpretable measure of multicollinearity.]
 

### 2024-12-01 scripts/03-clean_data.R
updated the cleaning and testing script for the analysis data

@@ -1,19 +1,17 @@
 #### Preamble ####
-# Purpose: Cleans the raw crime and traffic datasets for further analysis and saves them in Parquet format.
+# Purpose: Cleans the raw theft dataset for further analysis and includes the day of the week (`REPORT_DOW`) and saves the data in CSV format.
 # Author: [Your Name]
 # Date: [Current Date]
 # Contact: [Your Email]
 # License: MIT
 # Pre-requisites: 
-# - The `tidyverse`, `janitor`, `jsonlite`, and `arrow` packages must be installed.
-# Additional Information: Ensure the datasets are available in the specified locations.
+# - The `tidyverse`, `janitor`, `arrow`, and `here` packages must be installed.
 
 #### Workspace setup ####
 # Load required libraries
 library(tidyverse)
 library(janitor)
 library(here)
-library(jsonlite)
 library(arrow)
 
 # Ensure required directories exist
@@ -34,9 +32,10 @@ cleaned_crime_data <- crime_data |>
   mutate(
     report_date = as.Date(report_date, format = "%Y-%m-%d"),
     occ_date = as.Date(occ_date, format = "%Y-%m-%d"),
-    report_hour = as.integer(report_hour),  # Ensure `REPORT_HOUR` is numeric
     lat_wgs84 = as.numeric(lat_wgs84),
-    long_wgs84 = as.numeric(long_wgs84)
+    long_wgs84 = as.numeric(long_wgs84),
+    report_hour = as.integer(report_hour),  # Ensure hour is numeric
+    report_dow = weekdays(report_date)  # Add day of the week column
   ) |>
   filter(
     !is.na(event_unique_id) & 
@@ -44,13 +43,11 @@ cleaned_crime_data <- crime_data |>
       !is.na(division) & 
       !is.na(offence) & 
       !is.na(lat_wgs84) & 
-      !is.na(long_wgs84) & 
-      !is.na(report_hour)  # Remove rows with missing critical fields, including report_hour
+      !is.na(long_wgs84)  # Remove rows with missing critical fields
   ) |>
   filter(
     lat_wgs84 >= -90 & lat_wgs84 <= 90,  # Validate latitude
-    long_wgs84 >= -180 & long_wgs84 <= 180,  # Validate longitude
-    report_hour >= 0 & report_hour <= 23  # Validate report_hour
+    long_wgs84 >= -180 & long_wgs84 <= 180  # Validate longitude
   ) |>
   mutate(
     division = str_trim(division),  # Remove extra spaces
[…]

### 2024-12-01 scripts/04-test_analysis_data.R

@@ -28,7 +28,7 @@ test_that("cleaned_crime_data has unique event IDs", {
 
 # Column validation
 test_that("cleaned_crime_data has the required number of columns", {
-  expect_equal(ncol(cleaned_crime_data), 12)  # Based on the cleaned crime data structure
+  expect_equal(ncol(cleaned_crime_data), 13)  # Based on the cleaned crime data structure
 })
 
 # Dataframe has rows

### 2024-12-01 scripts/06-model_data.R
get rds files for the three models by running the model script

@@ -1,50 +1,296 @@
 #### Preamble ####
-# Purpose: To build and save a Bayesian regression model using cleaned crime or traffic data.
+# Purpose: Calculate theft sub-indexes and output as CSV
 # Author: Yingke He
-# Date: 29 November 2024
-# Contact: kiki.he@mail.utoronto.ca
+# Date: [Insert Date]
+# Contact: [Insert Email]
 # License: MIT
-# Pre-requisites: 
-# - Required packages installed
-# - Cleaned datasets (`cleaned_crime_data.csv` and `cleaned_traffic_data.csv`) available
-# - Test script successfully run to ensure data quality
+# Pre-requisites:
+# - Data cleaned using 03-clean_data.R
+# - Libraries such as tidyverse must be installed
 
-#### Workspace setup ####
-# Load necessary libraries
+#### Workspace Setup ####
 library(tidyverse)
-library(rstanarm)
 library(here)
 
-# Load data (adjust based on the dataset you're modeling)
-crime_data <- read_csv(here("data", "02-analysis_data", "cleaned_crime_data.csv"))
-traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+# Load dataset
+cleaned_crime_data <- read_csv(here("data", "02-analysis_data", "cleaned_crime_data.csv"))
 
-# Example: Using traffic data for the model
-# Modify the formula and variables based on your research question
-model_data <- traffic_data %>%
-  filter(!is.na(road_type), !is.na(accident_id), !is.na(injury)) %>%  # Ensure no missing values in key columns
+#### Calculate Theft Sub-Indexes ####
+
+# 1. Hour Index
+hour_index <- cleaned_crime_data %>%
+  count(hour) %>%
   mutate(
-    injury = as.numeric(as.factor(injury))  # Convert injury to numeric (e.g., severity levels)
+    proportion = n / sum(n),
+    index = proportion / 3,  # Normalize so the sum equals 1/3
+    type = as.character(hour)
+  ) %>%
+  select(type, index)
+
+# 2. Day of Week Index
+day_index <- cleaned_crime_data %>%
+  count(day_of_week) %>%
+  mutate(
+    proportion = n / sum(n),
+    index = proportion / 3,  # Normalize so the sum equals 1/3
+    type = day_of_week
+  ) %>%
+  select(type, index)
+
+# 3. Premises Type Index
+premises_index <- cleaned_crime_data %>%
+  count(premises_type) %>%
+  mutate(
[…]

### 2024-12-01 paper/paper.qmd
paper finished model section + initial plans written for other sections

@@
 library(palmerpenguins)
+library(knitr)
+
+library(here)
+library(car)
+
+# Define file paths for data
+theft_data_path <- here("data", "02-analysis_data", "cleaned_crime_data.csv")
+traffic_data_path <- here("data", "02-analysis_data", "cleaned_traffic_data.csv")
+
+```
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+#### Workspace Setup ####
+library(tidyverse)
+library(here)
+
+# Load dataset
+cleaned_crime_data <- read_csv(here("data", "02-analysis_data", "cleaned_crime_data.csv"))
+
+#### Calculate Theft Sub-Indexes ####
+
+# 1. Hour Index
+hour_index <- cleaned_crime_data %>%
+  count(hour) %>%
+  mutate(
+    proportion = n / sum(n),
+    index = proportion / 3,  # Normalize so the sum equals 1/3
+    type = as.character(hour)
+  ) %>%
+  select(type, index)
+
+# 2. Day of Week Index
+day_index <- cleaned_crime_data %>%
+  count(day_of_week) %>%
+  mutate(
+    proportion = n / sum(n),
+    index = proportion / 3,  # Normalize so the sum equals 1/3
+    type = day_of_week
+  ) %>%
+  select(type, index)
+
+# 3. Premises Type Index
+premises_index <- cleaned_crime_data %>%
+  count(premises_type) %>%
+  mutate(
+    proportion = n / sum(n),
+    index = proportion / 3,  # Normalize so the sum equals 1/3
+    type = premises_type
+  ) %>%
+  select(type, index)
+
+# Combine all indexes into one table
+theft_indexes <- bind_rows(
+  hour_index,
+  day_index,
+  premises_index
+)
+
+#### Save as CSV ####
+output_path <- here("data", "02-analysis_data", "theft_indexes.csv")
+write_csv(theft_indexes, output_path)
+
+#### Define the Theft Scoring Model ####
+# Create a theft scoring model function
+theft_scoring_model <- list(
+  hour_index = hour_index,
+  day_index = day_index,
+  premises_index = premises_index,
+  scoring_logic = function(hour, day_of_week, premises_type) {
+    # Lookup values for indexes
+    hour_score <- hour_index %>% filter(type == as.character(hour)) %>% pull(index)
+    day_score <- day_index %>% filter(type == day_of_week) %>% pull(index)
+    premises_score <- premises_index %>% filter(type == premises_type) %>% pull(index)
+
+    # Calculate total theft risk score
+    total_score <- sum(hour_score, day_score, premises_score, na.rm = TRUE)
+    return(total_score)
+  }
+)
+
+#### Save the Theft Scoring Model ####
+rds_output_path <- here("models", "theft_scoring_model.rds")
+saveRDS(theft_scoring_model, rds_output_path)
+
+# Load datasets
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+#### Preprocess Traffic Data ####
+# Add a binned time-of-day variable
+cleaned_traffic_data <- cleaned_traffic_data %>%
+  mutate(
+    time_of_day_bin = case_when(
+      time_of_day >= 0 & time_of_day < 600 ~ "Late Night",
+      time_of_day >= 600 & time_of_day < 1200 ~ "Morning",
+      time_of_day >= 1200 & time_of_day < 1800 ~ "Afternoon",
+      time_of_day >= 1800 & time_of_day <= 2359 ~ "Evening",
+      TRUE ~ "Unknown"
+    )
+  )
+
+# Preprocess severity as a binary variable
+cleaned_traffic_data <- cleaned_traffic_data %>%
+  mutate(
+    severity = case_when(
+      injury == "Major" | accident_classification == "Fatal" ~ 1,  # Severe cases
+      injury %in% c("Minor", "Minimal") ~ 0,                       # Non-severe cases
+      TRUE ~ NA_real_                                              # NA for missing or undefined
+    )
+  ) %>%
+  filter(!is.na(severity))  # Remove rows where severity could not be determined
+
+#### Logistic Regression Model for Collision Severity ####
+collision_severity_model <- glm(
+  formula = severity ~ hood_158 + time_of_day_bin + traffic_control +
+    visibility_conditions + lighting_conditions + road_conditions,
+  family = binomial(link = "logit"),
+  data = cleaned_traffic_data
+)
+
+# Summarize the model
+cat("Collision Severity Model Summary:\n")
+summary(collision_severity_model)
+
+#### Add Predicted Probabilities to Traffic Data ####
+# Filter traffic data for consistent training and prediction
+filtered_traffic_data <- cleaned_traffic_data %>%
+  filter(!is.na(severity))  # Match rows used for training
+
+# Debugging: Check sizes for consistency
+cat("Rows in filtered data (used for model training): ", nrow(filtered_traffic_data), "\n")
+cat("Rows in original traffic data: ", nrow(cleaned_traffic_data), "\n")
+
+# Generate predictions for the filtered dataset
+filtered_traffic_data <- filtered_traffic_data %>%
+  mutate(
+    collision_probability = predict(collision_severity_model, newdata = filtered_traffic_data, type = "response")
+  )
+
+# Merge predictions back into the original dataset
+cleaned_traffic_data <- cleaned_traffic_data %>%
+  left_join(
+    filtered_traffic_data %>%
+      select(id, collision_probability),  # Ensure unique identifier is used
+    by = "id"
+  )
+
+# Debugging: Check for successful merging
+cat("Rows in cleaned traffic data after merging predictions: ", nrow(cleaned_traffic_data), "\n")
+
+#### Save Model and Results ####
+# Save the logistic regression model
+saveRDS(collision_severity_model, here("models", "collision_severity_model.rds"))
+
+#### Model Diagnostics ####
+# Correlation between predictors (optional)
+cat("Correlation matrix of numeric predictors:\n")
+correlation_matrix <- cleaned_traffic_data %>%
+  select(time_of_day, fatal_no) %>%  # Replace with relevant numeric predictors
+  cor(use = "complete.obs")
+print(correlation_matrix)
+
+#### Save Collision Risk Scores ####
+# Select relevant columns to save
+collision_risk_scores <- cleaned_traffic_data %>%
+  select(id, hood_158, collision_probability)
+
+# Define output path
+collision_output_path <- here("data", "02-analysis_data", "collision_risk_scores.csv")
+
+# Save as CSV
+write_csv(collision_risk_scores, collision_output_path)
+
+#### Load Data ####
+# Load theft indexes
+theft_indexes <- read_csv(here("data", "02-analysis_data", "theft_indexes.csv"))
+
+# Load theft data
+cleaned_crime_data <- read_csv(here("data", "02-analysis_data", "cleaned_crime_data.csv"))
+
+# Load traffic data with collision probabilities
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+#### Ensure Collision Probability ####
+# If collision_probability is missing, calculate it
+if (!"collision_probability" %in% colnames(cleaned_traffic_data)) {
+  cleaned_traffic_data <- cleaned_traffic_data %>%
+    mutate(
+      time_of_day_bin = case_when(
+        time_of_day >= 0 & time_of_day < 600 ~ "Late Night",
+        time_of_day >= 600 & time_of_day < 1200 ~ "Morning",
+        time_of_day >= 1200 & time_of_day < 1800 ~ "Afternoon",
+        time_of_day >= 1800 & time_of_day <= 2359 ~ "Evening",
+        TRUE ~ "Unknown"
+      ),
+      severity = case_when(
+        injury == "Major" | accident_classification == "Fatal" ~ 1,
+        injury %in% c("Minor", "Minimal") ~ 0,
+        TRUE ~ NA_real_
+      )
+    ) %>%
+    filter(!is.na(severity))
+
+  collision_severity_model <- glm(
+    formula = severity ~ hood_158 + time_of_day_bin + traffic_control +
+      visibility_conditions + lighting_conditions + road_conditions,
+    family = binomial(link = "logit"),
+    data = cleaned_traffic_data
+  )
+
+  cleaned_traffic_data <- cleaned_traffic_data %>%
+    mutate(
+      collision_probability = predict(collision_severity_model, newdata = ., type = "response")
+    )
+}
+
+#### Calculate Theft Component ####
+# Aggregate theft indexes for simplicity
+hour_index <- theft_indexes %>% filter(str_detect(type, "^\\d+$"))
+day_index <- theft_indexes %>% filter(type %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
+premises_index <- theft_indexes %>% filter(type %in% c("House", "Outside", "Commercial"))
+
+# Add indexes to the theft data
+cleaned_crime_data <- cleaned_crime_data %>%
+  mutate(
+    hour = as.character(hour),  # Ensure hour is a character
+    hour = if_else(hour == "24", "0", hour)  # Handle 24:00 case
+  ) %>%
+  left_join(hour_index, by = c("hour" = "type")) %>%
+  rename(hour_index = index) %>%
+  left_join(day_index, by = c("day_of_week" = "type")) %>%
+  rename(day_index = index) %>%
+  left_join(premises_index, by = c("premises_type" = "type")) %>%
+  rename(premises_index = index) %>%
+  mutate(
+    theft_component = hour_index + day_index + premises_index  # Sum theft sub-indexes
+  )
+
+#### Add Theft Component to Traffic Data ####
+# Add the average theft component by neighborhood
+theft_by_neighborhood <- cleaned_crime_data %>%
+  group_by(hood_158) %>%
+  summarize(avg_theft_component = mean(theft_component, na.rm = TRUE))
+
+cleaned_traffic_data <- cleaned_traffic_data %>%
+  left_join(theft_by_neighborhood, by = "hood_158") %>%
+  mutate(avg_theft_component = replace_na(avg_theft_component, 0))  # Replace NA with 0
+
+#### Calculate Final Risk Index ####
+# Define weights for components
+weight_collision <- 0.7
+weight_theft <- 0.3
+
+# Calculate risk index
+cleaned_traffic_data <- cleaned_traffic_data %>%
+  mutate(
+    risk_index = weight_collision * collision_probability + weight_theft * avg_theft_component
+  )
+
+#### Save Final Risk Index ####
+# Save the final dataset with the risk index
+output_path <- here("data", "02-analysis_data", "final_risk_index.csv")
+write_csv(cleaned_traffic_data %>% select(id, hood_158, risk_index), output_path)
+
+#### Save the Risk Calculation Model ####
+# Define the final risk model
+final_risk_model <- list(
+  collision_model = collision_severity_model,
+  theft_indexes = theft_indexes,
+  weight_collision = weight_collision,
+  weight_theft = weight_theft,
+  risk_calculation_logic = function(collision_probability, avg_theft_component) {
+    risk_index <- weight_collision * collision_probability + weight_theft * avg_theft_component
+    return(risk_index)
+  }
+)
+
+# Save the final risk model as RDS
+rds_output_path <- here("models", "final_risk_model.rds")
+saveRDS(final_risk_model, rds_output_path)
+
+#### End of Script ####
 ```
@@
 
+Need two graphs
 ```{r}
@@
 ```
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
@@
 
+### Theft index
+
+#### Hour of the Day
+#### Day of the Week
+#### Premises Type
+
 ### Collision Probability
-
-#### Neighborhood ID
-#### Premises Type
-#### Report Hour 
-#### Location Type
-#### Report Day
-
-### Theft Probability
 
@@
 #### Traffic Control
+#### Time of Day
+#### Visibility
 
@@
 
-The main idea of this composite risk score model is to calculate an index of risk for owning and riding a motorbike.
+The main purpose of this composite risk score model is to calculate a Risk Index for owning and riding a motorbike, which integrates the risks of theft and collisions.
 The modeling strategy has two primary objectives.
-The first objective is to estimate the likelihood of theft \( P(\text{Theft}) \) and collision \( P(\text{Collision}) \) under diverse environmental and situational conditions using logistic regression models.
-The second objective is to combine these probabilities into a unified Risk Index, providing actionable insights into the risk for owning and riding a motorbike
+The first objective is to estimate the likelihood of collisions under various environmental and situational conditions using a logistic regression model.
+The second objective is to derive a theft risk score based on time of day, day of the week, and premises type, combining these components into a unified Risk Index to provide actionable insights into motorbike-related risks.
+The models were developed in R [@citeR] using the `stats` package for logistic regression and the `tidyverse` package for data preprocessing and manipulation.
+The theft model calculates sub-indexes for specific predictors, while the collision model uses logistic regression to estimate probabilities based on predictors such as neighborhood ID, road surface condition, lighting condition, and traffic control.
+Both models are designed to enable reliable predictions under diverse conditions and are integrated into the final Risk Index, which highlights areas, times, and conditions with elevated risks.
 
-The models were run in R [@citeR] using the `stats` package to implement logistic regression models predicting theft and collision probabilities.
-The theft model incorporates predictor variables such as neighborhood ID, premises type, report hour, location type, and report day, while the collision model includes neighborhood ID, road surface condition, lighting condition, and traffic control.
+Detailed model diagnostics, variable descriptions, and performance metrics are included in [Appendix @sec-model-details].
 
-Both models estimate the log-odds of their respective probabilities using the predictors,  enabling reliable predictions under varying environmental and situational conditions.
-Model diagnostics and background details, including variable descriptions and performance metrics, are included in [Appendix @sec-model-details].
+## Model Set-Up
 
-## Model Set-up
+### Theft Risk Sub-Indexes
 
-The logistic regression model for theft probability takes the form of the following equation:
+To capture theft risk without relying on logistic regression due to the absence of negative (non-theft) cases, we calculated sub-indexes for three critical factors:
+
+-Hour of the Day: Risk distribution across 24 hours, normalized so the sum equals 1/3.
+
+-Day of the Week: Risk distribution across 7 days, normalized so the sum equals 1/3.
+
+-Premises Type: Risk distribution across premises types (House, Outside, Commercial), normalized so the sum equals 1/3.
+
+The total theft component is calculated as:
+\begin{align*}
+C_{\text{Theft}} &= \text{Hour Index} + \text{Day Index} + \text{Premises Type Index}
+\end{align*}
+
+This approach ensures proportional representation of each factor while accounting for varying risks based on time and location characteristics.
+
+### Collision Probability Model
+A logistic regression model was used to predict the likelihood of severe collisions P(Collision) based on several predictors. The log-odds of the collision probability are modeled as:
+\
 
 \begin{align*}
-\log\left(\frac{P(\text{Theft})}{1 - P(\text{Theft})}\right) &= \beta_0 + \beta_1 \cdot \text{HOOD}_{158} + \beta_2 \cdot \text{Premises Type} + \beta_3 \cdot \text{Report Hour} \\
-&\quad + \beta_4 \cdot \text{Location Type} + \beta_5 \cdot \text{Report Day} + \epsilon
+\log\left(\frac{P(\text{Collision})}{1 - P(\text{Collision})}\right) &= \beta_0 + \beta_1 \cdot \text{HOOD}_{158} + \beta_2 \cdot \text{Road Surface Condition} + \beta_3 \cdot \text{Lighting Condition} \\
+&\quad + \beta_4 \cdot \text{Traffic Control} + \beta_5 \cdot \text{Road Visibility} + \beta_6 \cdot \text{Time of Day} + \epsilon
 \end{align*}
 
-This model utilizes the following predictor variables:
+The model prediction utilizes the following predictor variables:
 
-- **Neighborhood ID** (`hood_158`): Unique identifier for the neighborhood.
-- **Premises Type** (`premises_type`): Type of premises where the theft occurred like parking lot, garage
-- **Report Hour** (`REPORT_HOUR`):  Hour of the day the theft was reported.
-- **Location Type** (`location`):Location of the theft like indoors and outdoors.
-- **Report Day** (`REPORT_DOW`): Day of the Week Offence was Reported.
+- Neighborhood ID (`hood_158`): Unique identifier for the neighborhood.
+- Road Surface Condition (`road_conditions`): Conditions such as dry, wet, or icy.
+- Lighting Condition (`lighting_conditions`): Visibility levels, such as daylight or artificial light.
+- Traffic Control (`traffic_control`): Presence of traffic management devices (e.g., stop signs, signals).
+- Road Visibility (`visibility_conditions`): Road visibility conditions, such as clear, snow or rain
+- Time of Day (`time`): Time where collision occured in Toronto
 
-The logistic regression model for collision probability takes the form of the following equation:
+The model assigns coefficients as follows:
 
 \begin{align*}
-\log\left(\frac{P(\text{Collision})}{1 - P(\text{Collision})}\right) &= \beta_0 + \beta_1 \cdot \text{HOOD}_{158} + \beta_2 \cdot \text{Road Surface Condition} + \beta_3 \cdot \text{Light} \\
-&\quad + \beta_4 \cdot \text{Traffic Control} + \epsilon
+\beta_i \text{ to each variable, enabling the calculation of collision probability }\\ 
+P(\text{Collision})\text{ under specific environmental and situational conditions.} 
 \end{align*}
 
-This model utilizes the following predictor variables:
+### Risk Index Calculation
 
-- **Neighborhood ID** (`hood_158`): Unique identifier for the neighborhood.
-- **Road Surface Condition** (`road_conditions`): Road surface condition like wet and dry.
-- **Lighting Condition** (`lighting_conditions`): Lighting conditions at the time of the collision like dark or clear.
-- **Traffic Control** (`traffic_control`):Type of traffic control present like stop signs or traffic signals
-After fitting both models, combine the probabilities into a weighted risk index:
+The final Risk Index integrates the collision probability P(Collision) and theft component T using weighted aggregation:
 
 \begin{align*}
-\text{Risk Index} &= w_1 \cdot P(\text{Theft}) + w_2 \cdot P(\text{Collision})
+\text{Risk Index} = w_1\cdot P(\text{Collision}) + w_2\cdot T
 \end{align*}
 
+Weights are defined as:
 \begin{align*}
-\text{where weights } w_1 \text{and } w_2 \text{ are adjust based on the relative importance of theft and collision risks.}
+w_1 = 0.7, \quad w_2 = 0.3
 \end{align*}
+These reflect the relative importance of collision and theft risks, emphasizing collision severity due to its greater immediate impact.
 
-## Model justification
+## Model Justification
 
-Logistic regression was selected for this analysis due to its simplicity and effectiveness in estimating probabilities for binary outcomes, such as theft and collision.
-This statistical method is well-suited for modeling the log-odds of an event's occurrence as a linear combination of predictor variables, providing a clear framework for understanding the relationships between independent variables and the dependent outcome.
-Its ability to estimate odds ratios allows for an intuitive interpretation of the impact of predictor variables, making it particularly effective for identifying and quantifying risks in diverse scenarios.
+The analysis adopts a hybrid approach that combines sub-index calculations for theft risk with logistic regression for collision probability.
+This design ensures that the model reflects the specific data characteristics and practical considerations when assessing motorbike-related risks.
+Logistic regression is not utilized for theft risk due to the absence of negative (non-theft) cases in the dataset.
+Instead, theft risk is represented through sub-index calculations for three critical factors: hour of the day, day of the week, and premises type.
+Each factor contributes equally to the total theft component.
+The Hour of the Day Index captures temporal variations in theft risk across 24 hours, while the Day of the Week Index accounts for weekly patterns of theft.
+The Premises Type Index reflects variations in risk based on location type, such as houses, outdoor spaces, or commercial premises.
+These sub-indexes are normalized so their contributions to the theft component are proportional and balanced.
+This approach ensures an accurate representation of theft risk patterns while providing actionable insights into temporal and spatial risk factors.
 
-This model's strengths include its flexibility in handling both categorical and continuous predictors, enabling the integration of a wide range of variables such as neighborhood characteristics, premises type, lighting conditions, road surface conditions, time of day, and environmental conditions.
-Additionally, logistic regression’s ability to include interaction terms and manage potential multicollinearity enhances its effectiveness in capturing complex relationships between predictors and the likelihood of thefts and collisions.
-Utilizing these features enables the model to deliver a thorough evaluation of the factors affecting theft and collision probabilities, addressing the first objectives of the modelling strategy.
+For collision risk, the model utilizes logistic regression due to its effectiveness in estimating probabilities for binary outcomes.
+The log-odds of collision probability are modeled as a function of neighborhood-specific characteristics, road surface conditions, lighting conditions, and traffic control measures.
+By including these predictors, the model accounts for diverse factors that influence collision risks.
+Logistic regression’s ability to handle both categorical and continuous variables makes it an appropriate choice for this component, delivering statistically reliable estimates and facilitating the interpretation of individual predictors’ effects.
 
-After estimating the probabilities of theft and collision, these probabilities are combined into a composite Risk Index using a weighted formula.
-This Risk Index offers a unified measure of risk, allowing for comparative evaluations of the relative safety of motorbike ownership and usage across varying contexts.
+The final Risk Index integrates the theft and collision components using a weighted formula.
+The collision probability component is weighted at 0.7, reflecting its higher immediate impact on safety, while the theft component is weighted at 0.3.
+This weighting scheme prioritizes collision risk while ensuring that theft risk is not overlooked.
+By combining these components, the Risk Index provides a unified measure of motorbike-related risks, enabling stakeholders to assess and compare safety conditions across different contexts.
 
-The modeling approach is justified by logistic regression’s ability to deliver clear, interpretable, and statistically reliable risk estimates.
-Its statistical efficiency, with its capacity to manage complex, large datasets, makes it an ideal choice for accurately predicting theft and collision risk, while the composite Risk Index facilitates a clearer understanding of the risks associated with motorbike ownership and usage.
+This modeling approach is justified by its ability to adapt to the data’s constraints while maintaining statistical rigor and interpretability.
+The sub-index method for theft risk is tailored to the dataset’s characteristics, avoiding assumptions about unobserved cases, and the use of logistic regression for collision risk ensures robust and reliable predictions.
+Together, these components form a comprehensive and practical framework for evaluating motorbike ownership and usage risks, addressing both immediate safety concerns and long-term theft risks.
 
-## Model Assumtions and Validations
+## Model Assumptions and Validations
 
-To validate the use of logistic regression models, the four main assumptions: binary nature of the outcome, linearity of the logit, indipendence of observations, and lack of multicollinearity will be assessed [@kononen2011logistic].
-To validate the use of a composite Risk Index, the correlation of the index with observed outcomes will be assess using a Pearson correlation test.
+### Theft Sub-Index [add one sentence of validation]
 
-### Logistic Regression Models
+The theft sub-index approach is based on two key assumptions.
+First, it assumes that the observed proportions of theft occurrences across categories, such as hour of the day, day of the week, and premises type, accurately represent the overall theft risk.
+Second, it assumes that these categories contribute independently to the theft risk, meaning the risk associated with one category does not influence or depend on another.
+
+### Composite Risk Index [Add one sentence of assumption]
+
+The validation of the composite Risk Index involves two key steps.
+First, its correlation with observed collision severity will be tested to ensure alignment with real-world risks.
+Second, a sensitivity analysis will be conducted by testing alternate weightings for the collision and theft components $w_1$ and $w_2$ to evaluate the robustness and reliability of the final Risk Index.
+
+### Collision Model
+
+The logistic regression model for collision severity relies on several assumptions.
+First, the response variable (`severity`) is binary (1 = severe, 0 = non-severe), fulfilling the requirement for a binary outcome.
+Second, the data consist of independently reported collision incidents, ensuring that observations are uncorrelated.
+Third, the log-odds of collision severity are modeled as a linear combination of predictor variables; although most predictors are categorical, their contributions to the log-odds inherently satisfy this linearity assumption.
+Finally, to address the assumption of no multicollinearity, Variance Inflation Factor (VIF) will be calculated for all predictors.
+Predictors with high VIF values will be mitigated through re-categorization or removal to ensure stable and reliable coefficient estimates.
 
@@
 A fundamental assumption of logistic regression is that the response variable is binary or dichotomous, meaning it can take on only two possible outcomes [@nick2007logistic].
-This assumption is satisfied in both models.
-In the theft model, the response variable indicates whether a theft occurred, with a value of 1 representing the occurrence of theft and 0 representing its absence.
-Similarly, in the collision model, the response variable denotes whether a collision occurred, where 1 represents the occurrence of a collision and 0 indicates no collision.
+This assumption is satisfied in the collision model, where the response variable (`severity`) distinguishes between severe (1) and non-severe (0) collision cases.
+
+In the theft dataset, however, all entries represent theft cases, precluding the binary nature required for logistic regression.
+Consequently, the theft model was adapted to calculate proportion-based sub-indexes rather than relying on a binary outcome.
+These sub-indexes represent relative risk based on temporal and spatial factors, such as hour of the day, day of the week, and premises type.
+
+In the collision model:
+- The model estimates the probability of a severe collision (\(P(\text{Collision})\)) given environmental and situational predictors. The response variable (`severity`) is defined as:
+  \begin{align*}
+  P(\text{Collision}) = \frac{\exp(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_k X_k)}{1 + \exp(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_k X_k)}
+  \end{align*}
+  where:
+  \begin{align*}
+  \beta_0 &\text{ is the intercept,} \\
+  \beta_k &\text{ are the coefficients, and} \\
+  X_k &\text{ are the predictor variables.}
+  \end{align*}
 
 The logistic regression model does not directly predict 1 or 0.
-Instead, it estimates the probability that the response variable equals 1, given the predictor variables.
-For instance, the theft model outputs the probability of theft occurring (P(Theft)), and the collision model outputs the probability of a collision occurring (P(Collision)).
-These probabilities range between 0 and 1, allowing for detailed predictions of the likelihood of the respective events.
+Instead, it provides a continuous probability ranging between 0 and 1.
+This probability reflects the likelihood of an event (e.g., a severe collision) occurring under the given conditions.
+
+#### Independence of Observations
+
+The collision model assumes that each observation is independent of the others, a fundamental requirement for logistic regression.
+This assumption is satisfied in the dataset, as each row represents a distinct and independently reported collision incident.
+The observations are not repeated or correlated, ensuring that the logistic regression model provides unbiased estimates of the relationships between predictor variables and the response variable.
+By meeting this assumption, the collision model maintains its validity for estimating probabilities of severe collisions and contributes robustly to the composite Risk Index.
 
@@
 
-[Identifying Continuous Variables in Your Models:
-For LR Model 1 (Theft Model):
-Continuous Variable:
-REPORT_HOUR (Hour of the day the theft was reported
-For LR Model 2 (Collision Model):
-No Continuous Variables]
+One of the key assumptions in logistic regression is that a linear relationship exists between the continuous predictors and the logit of the outcome variable [@stoltzfus2011].
+This means that the log-odds of the binary dependent variable should have a linear association with any continuous independent variables in the model.
+It is important to test this assumption to ensure the validity of the model.
 
-[Conclusion:
-Total Number of Plots: 1
-Draw a binning plot for REPORT_HOUR in LR Model 1 to assess the linear relationship]
+The collision risk logistic regression model incorporates both categorical and continuous predictor variables.
+Among these, Time of Day serves as the sole continuous predictor, representing the time an incident occurred as a numerical value ranging from 0 (midnight) to 2359 (just before midnight).
+The analysis emphasizes the continuous predictor, Time of Day, to evaluate its role within the collision model.
+
+Linearity is assessed using smoothed scatter plots of the predicted logit values:
+
+\begin{align*}
+\text{logit} = \log\left(\frac{P}{1 - P}\right)
+\end{align*}
+
+where P represents the predicted probability of collision from the logistic regression model plotted against the continuous predictors.
+These plots are intended to visualize the relationship between each predictor and the logit of the outcome variable, providing insight into whether the relationship is approximately linear.
+
+```{r}
+#| label: fig-linear-time-of-day
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Logit Plot for Time of Day in the Collision Probability Model"
+#| fig-pos: H
+# Load necessary libraries
+library(tidyverse)
+library(ggplot2)
+library(here)
+
+# Use here() to construct file paths and then load the data into data frames
+collision_data <- read_csv(here("data", "02-analysis_data", "collision_risk_scores.csv"))
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+# Merge collision data with time_of_day from cleaned traffic data
+collision_data <- collision_data %>%
+  left_join(cleaned_traffic_data %>% select(id, time_of_day), by = "id") %>%
+  mutate(
+    collision_probability = ifelse(collision_probability <= 0, 0.0001, collision_probability),
+    collision_probability = ifelse(collision_probability >= 1, 0.9999, collision_probability),
+    logit = log(collision_probability / (1 - collision_probability))  # Logit transformation
+  )
+
+# Smoothed scatter plot for time_of_day vs logit
+ggplot(collision_data, aes(x = time_of_day, y = logit)) +
+  geom_point(alpha = 0.5, color = "darkgrey") +
+  geom_smooth(method = "loess", color = "blue", fill = "lightblue") +
+  coord_cartesian(ylim = c(-5, 5)) +  # Restrict the y-axis range
+  labs(
+    x = "Time of Day",
+    y = "Logit"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")  # Center title and adjust size
+  )
+
+```
+
+Figure @fig-linear-time-of-day illustrates the relationship between variable Time of Day and the predicted values of the collision probability model.
+The horizontal axis represents the reported time of day in minutes since midnight, while the vertical axis displays the logit values.
+Gray points depict the raw data, and the blue smoothed line represents the average trend.
+To evaluate the linearity assumption, the smoothed line is compared to a hypothetical linear relationship.
+Significant deviations of the smoothed line from a straight line may suggest a potential violation of the linearity assumption.
+
+In this plot, the smoothed line shows a relatively stable trend with minimal deviations, indicating no substantial evidence against the linearity assumption.
+Local fluctuations are minor and likely reflect natural variations in the data rather than a systematic departure from linearity.
 
@@
 
-[The easiest way to check for multicollinearity in your logistic regression models is to calculate the Variance Inflation Factor (VIF) for your predictors.
-This provides a straightforward and interpretable measure of multicollinearity.]
+Another key assumption of logistic regression is the absence of multicollinearity among predictor variables.
+Multicollinearity occurs when two or more predictors are highly correlated, leading to inflated standard errors of the regression coefficients and reducing the reliability of the model's estimates.
+The Variance Inflation Factor (VIF) is commonly used to assess multicollinearity, with VIF values greater than 5 indicating potential issues, and values above 10 suggesting severe multicollinearity [@stoltzfus2011].
 
-### Composite Risk Index
+In the collision risk model, the predictors include both categorical variables (Lighting Conditions, Road Conditions, Visibility Conditions and Traffic Control) and one continuous variable (Time of Day).
+VIF calculations are performed to evaluate the degree of multicollinearity among these predictors.
 
-#### Pearson Correlation Test
+If multicollinearity is detected, strategies such as combining correlated variables, removing redundant predictors, or using regularization techniques like ridge regression can be employed [@stoltzfus2011].
+However, if VIF values remain below the threshold, it confirms that multicollinearity is not a concern in this analysis.
+
+The following table presents the VIF values for all predictors in the collision risk model.
+
+```{r}
+#| label: tbl-vif
+#| echo: false
+#| warning: false
+#| message: false
+#| tbl-cap: "VIF values for predictor variables in the collision model."
+#| tbl-pos: H
+library(performance)
+
+# Run check_collinearity on the collision model
+vif_result <- check_collinearity(collision_severity_model)
+
+# Convert results to a data frame
+vif_df <- as.data.frame(vif_result)
+
+# Create a data frame with GVIF values
+gvif_data <- data.frame(VIF = round(vif_df$VIF, 2))
+
+# Convert the data frame into a kable table
+vif_table <- kable(
+  t(gvif_data), # transpose
+  col.names = c("hood id", "time of day", "traffic control", "visibility", "lighting condition", "road condition"),
+  escape = FALSE,
+  align = "ccccc"
+)
+
+vif_table
+
+```
+
+The results of the VIF analysis for the collision risk model are presented in Table @tbl-vif.
+Most predictors exhibit VIF values well below the commonly used threshold of 5, indicating no significant multicollinearity among them.
+However, two predictors, Visibility Conditions and Road conditions, show slightly elevated VIF values of 7.29 and 7.51, respectively.
+While these values are higher than the others, they remain below the severe multicollinearity threshold of 10, suggesting that multicollinearity, though present to some extent, is not critical.
+
+The slightly elevated VIF values can be attributed to a potential overlap in the information captured by Visibility Conditions and Road Conditions.
+For instance, poor visibility often coincides with adverse road conditions, such as wet or icy surfaces, leading to some degree of correlation between these variables.
+
+Despite this overlap, both predictors are retained in the model because they provide distinct and meaningful contributions to understanding collision risk.
+Visibility Conditions directly reflects environmental factors like fog, heavy rain, or low light, which impair drivers' ability to see hazards.
+Conversely, Road Conditions` account for the physical state of the driving surface, such as wet, icy, or damaged roads, which influence vehicle handling and stopping distance.
+Together, these variables capture complementary aspects of collision risk, ensuring that the model provides a comprehensive assessment.
+
+The inclusion of both variables aligns with the theoretical framework underpinning the model and enhances its practical utility by addressing multiple dimensions of risk.
+While some degree of multicollinearity is observed, its impact on model stability is minimal, and the predictors' theoretical importance justifies their inclusion.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+## When Risk Peaks: Temporal Trends in the Index
+
+Time-of-Day and Day-of-Week Analysis of the Risk Index
+
+Plot Ideas:
+Bar: Show Risk Index variations across hours of the day and days of the week, highlighting theft and collision components separately.
+
+Time Series Plot: Display Risk Index trends over a longer period, revealing potential seasonal effects.
+
+```{r}
+```
+
+```{r}
+
+```
+
+## Conditions of Danger: Environmental Drivers of Risk
+
+Exploring How Environmental and Situational Factors Impact the Risk Index
+
+Plot Ideas:
+Violin Plot: Show the Risk Index distribution under varying road surface conditions (e.g., dry, wet, icy).
+
+Faceted Plot: Break down the Risk Index by lighting conditions (e.g., daylight, artificial light) or traffic control types.
 
@@
 
-## First Result Point
+## Mapping the Risk: Neighborhood-Level Insights
 
-## Second Result Point
+Spatial Patterns in the Distribution of the Risk Index
 
-## Third Result Point
+Plot Ideas:
+Choropleth Map: Display the Risk Index across neighborhoods with color gradients indicating risk levels.
+
+Bar Chart: Rank neighborhoods by their average Risk Index to identify the most vulnerable areas.
+
+```{r}
+
+```
+
+```{r}
+
+```
 
@@
 
-## First discussion point
+## Temporal Risk Patterns and Behavioral Recommendations
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+Focus: This section can interpret the temporal trends in the Risk Index and their implications for motorbike owners and policymakers.
+For instance, if risks peak at certain hours or days, actionable recommendations can be made for owners to avoid high-risk times or take precautions.
 
-## Second discussion point
+Key Points:
+Discuss high-risk times of day or days of the week for both theft and collisions.
+Recommendations for motorbike owners to reduce risk exposure, such as avoiding specific time periods or enhancing security measures during peak theft hours.
+Potential scheduling adjustments for law enforcement patrols to align with high-risk periods..
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Environmental and Situational Influences on Risk
 
-## Third discussion point
+Focus: This section can examine how road, lighting, and traffic conditions contribute to the Risk Index, linking these findings to actionable changes in infrastructure or urban planning.
 
-## Weaknesses and next steps
+Key Points:
+Discuss the significant environmental predictors of risk, such as poor road surface conditions or inadequate lighting.
+Recommend urban planning measures, such as improved road maintenance or installation of streetlights, to mitigate risks.
+Highlight how different environmental factors interact to amplify risks, suggesting multi-faceted approaches to improve safety.
 
-Weaknesses and next steps should also be included.
+## Policy Implications of Spatial Risk Disparities
 
+Focus: This section can discuss the significant disparities in the Risk Index across neighborhoods.
+Highlight how certain areas are disproportionately affected by theft or collisions and the potential socioeconomic or infrastructure factors contributing to these risks.
+
+Key Points:
+Prioritizing high-risk neighborhoods for targeted interventions, such as improved street lighting or traffic control measures.
+Suggestions for law enforcement strategies to reduce theft in hotspots.
+Need for localized awareness campaigns in high-risk areas.
+
+## Limitations
+
+## Future Research
+
+Future Work1: Explore how temporal risk patterns change with seasons or special events (e.g., holidays, festivals) to refine recommendations further.
+
+Future Work2: Develop predictive models integrating weather or real-time traffic data to dynamically assess and communicate motorbike risks.
+
+Future Work3: Investigate the correlation between neighborhood-level Risk Index values and broader social or economic indicators, such as income levels or population density.
 \newpage
@@
 
-# Appendix {.unnumbered}
+# Appendix {-}
 
-# Additional data details
+# Additional Data Details {#sec-data_details}
+
+## Data Cleaning
 
@@
 
+# Idealized Methodology for a Survey on Motor Risk {#sec-survey}
+
+## Survey Overview
+
+This survey aims to assess factors contributing to motorbike risks, including theft and collisions, by gathering data from motorbike owners and riders.
+The survey will explore demographic characteristics, riding behavior, motorbike usage patterns, and perceptions of environmental and situational risks.
+
+## Sampling Approach
+
+The survey will use a stratified random sampling method to ensure diverse representation across:
+
+- Geographic regions (urban, suburban, rural areas).
+- Demographic groups (age, gender, income, and education levels).
+- Riding experience (novices, intermediate, and experienced riders).
+
+The target population includes licensed motorbike riders and owners. A sample size of approximately 1,000 respondents is proposed to ensure statistical validity across strata.
+
+## Survey structure
+
+The survey will consist of five main sections:
+
+- Demographics: Basic information on respondents (e.g., age, gender, income, education, geographic location).
+
+- Riding Behavior: Frequency, duration, and purpose of motorbike usage.
+
+- Risk Awareness and Perception: Personal experiences with collisions or theft and perceptions of environmental risks.
+
+- Situational Factors: Details of riding conditions such as time of day, weather, road surface, and lighting.
+
+- Preventive Measures: Actions taken by riders to mitigate theft or collision risks (e.g., use of locks, helmets, or avoiding high-risk areas).
+
+### Question Types
+
+- Closed-ended questions: For quantitative data (e.g., multiple choice, Likert scales, ranking).
+
+- Open-ended questions: To capture nuanced insights and personal experiences.
+
+- Matrix questions: To evaluate attitudes across multiple dimensions efficiently.
+
+### Question List
+
+Here are examples of survey questions:
+
+Demographics:
+What is your age group? (e.g., 18–24, 25–34, etc.)
+What is your highest level of education completed?
+
+Riding Behavior:
+How often do you ride your motorbike? (Daily, Weekly, Monthly, Rarely)
+For what purposes do you primarily use your motorbike? (Commuting, Recreation, Delivery/Work, Other)
+
+Risk Awareness and Perception:
+On a scale of 1–5, how would you rate the theft risk in your neighborhood?
+Have you experienced a motorbike theft or collision in the past year? (Yes/No)
+
+Situational Factors:
+What time of day do you usually ride? (Morning, Afternoon, Evening, Night)
+In what weather conditions do you typically ride? (Clear, Rainy, Snowy)
+
+Preventive Measures:
+What measures do you take to secure your motorbike against theft? (Select all that apply: Lock, Alarm, GPS Tracker, Parking in Secure Locations)
+Do you avoid specific times or areas due to perceived collision risks? (Yes/No)
+
+## Recruitment Strategy
+
+Participants will be recruited through a combination of online and offline channels:
+
+Online platforms: Motorbike owner forums, social media groups, and email lists of motorbike organizations.
+
+Offline channels: Flyers at motorbike dealerships, repair shops, and riding schools. Incentives such as small gift cards or entry into a raffle may be provided to encourage participation.
+
+## Linkage to Literature
+
+The survey design is informed by prior studies on motor vehicle risk perception, road safety, and crime prevention. Key references include:
+
+Research on environmental and situational predictors of road accidents.
+Studies on the effectiveness of theft prevention measures.
+Literature on sampling methods and survey design for risk assessment.
+
 # References

### 2024-12-02 paper/paper.qmd
finished the first draft of result section of the paper

@@
 Recent statistics highlight the increased risks associated with owning and riding motorbikes, including a heightened likelihood of theft and collisions compared to other vehicles [@Yasmin2016].
-These risks are influenced by various factors such as geographic location, road conditions, time of day, and type of motorbike, making it essential to develop tools that effectively assess and mitigate these dangers.[can add one more sentence + citations]
+These risks are influenced by various factors such as geographic location, road conditions, time of day, and type of motorbike, making it essential to develop tools that effectively assess and mitigate these dangers.\[can add one more sentence + citations\]
 
@@
 This analysis confirms and extends three key findings: (1) material properties significantly influence motor efficiency, with high-conductivity materials yielding higher energy savings; (2) geometric optimization of motor components enhances torque generation and reduces energy losses; and (3) operational conditions, such as load and temperature, exhibit nonlinear effects, underscoring the need for adaptive motor control strategies.
-Furthermore, our projections suggest that integrating advanced materials and control systems could improve motor efficiency by up to 15% by 2050, aligning with global energy transition goals [To be updated..].
+Furthermore, our projections suggest that integrating advanced materials and control systems could improve motor efficiency by up to 15% by 2050, aligning with global energy transition goals \[To be updated..\].
 
@@
 
-[Libraries To be updated...]
+\[Libraries To be updated...\]
 
@@
 
-[Add variable types and trends]
+\[Add variable types and trends\]
 
@@
 
-The outcome variable of this analysis is the Risk Index, a composite metric that integrates the probabilities of two underlying events: theft and collision.
-
-Need two graphs
+The outcome variable of this analysis is the Risk Index, a composite metric that integrates the probabilities of two underlying events: theft and collision.The Risk Index integrates two underlying components: the theft component, derived from proportional sub-indexes for temporal and spatial factors, and the collision probability, estimated using a logistic regression model.
+By combining these elements, the Risk Index provides a unified measure of risk, enabling the identification of high-risk scenarios and areas.
+This section examines the distribution of the Risk Index through visualizations to reveal key patterns, variations, and potential drivers of motorbike-related risks.
+Two graphs are presented: (1) a histogram with density overlay to illustrate the overall distribution of the Risk Index, and (2) a boxplot by neighborhood to highlight spatial differences in risk.
+
 ```{r}
-
-```
+#| label: fig-risk-index-histogram
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Distribution of the Overall Risk Index"
+#| fig-pos: H
+#| fig-width: 8 
+#| fig-height: 4
+
+# Load necessary libraries
+library(tidyverse)
+library(here)
+
+# Load the final Risk Index dataset
+final_risk_index <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+
+# Plot a histogram with density overlay
+ggplot(final_risk_index, aes(x = risk_index)) +
+  geom_histogram(aes(y = ..density..), bins = 40, fill = "darkgrey", alpha = 0.6, color = "black") +
+  geom_density(color = "darkred", size = 1) +
+  labs(
+
+    x = "Risk Index",
+    y = "Density"
+  ) +
+  theme_minimal()
+
+```
+
+@fig-risk-index-histogram shows the distribution of the Risk Index across all observations in the dataset.
+The Risk Index exhibits a unimodal distribution, skewed slightly to the left, with the majority of values concentrated between 0.45 and 0.65.
+This indicates that most motorbike-related risks fall within a moderate range.
+The density curve overlay reveals a smooth progression in risk levels, with the peak occurring around a Risk Index value of 0.55, suggesting that this is the most common level of composite risk.
+The left tail, representing lower risk levels, is relatively small, while the right tail, corresponding to higher risks, extends further, indicating the presence of a smaller number of high-risk cases.
+
+The skewness and spread of the distribution highlight the variability in the combined risks of theft and collision.
+The extended tail on the higher end of the Risk Index suggests that certain environmental or situational factors disproportionately elevate risks in specific cases.
+This insight can inform targeted interventions, focusing on the outliers with elevated Risk Index values to mitigate the most critical risks.
+
+Building on the distribution of the Risk Index shown in the histogram, this box plot further dissects the data by grouping neighborhoods into risk categories based on their average Risk Index values.
+Neighborhoods are classified into three categories—Low Risk, Moderate Risk, and High Risk—using quartiles.
 
 ```{r}
-
-```
+#| label: fig-risk-index-boxplot
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Risk Index Distribution by Risk Category"
+#| fig-pos: H
+
+library(ggplot2)
+library(dplyr)
+
+risk_index_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+
+# Assume `risk_index_data` contains the `risk_index` values for each neighborhood
+# Group neighborhoods by average Risk Index
+grouped_data <- risk_index_data %>%
+  group_by(hood_158) %>%
+  summarize(avg_risk_index = mean(risk_index, na.rm = TRUE)) %>%
+  mutate(risk_category = case_when(
+    avg_risk_index <= quantile(avg_risk_index, 0.25) ~ "Low Risk",
+    avg_risk_index > quantile(avg_risk_index, 0.25) & avg_risk_index <= quantile(avg_risk_index, 0.75) ~ "Moderate Risk",
+    avg_risk_index > quantile(avg_risk_index, 0.75) ~ "High Risk"
+  ))
+
+# Add risk categories directly to the `risk_index_data`
+risk_index_data <- risk_index_data %>%
+  left_join(grouped_data %>% select(hood_158, risk_category), by = "hood_158")
+
+# Box plot of Risk Index by Risk Category with custom colors
+ggplot(risk_index_data, aes(x = risk_category, y = risk_index, fill = risk_category)) +
+  geom_boxplot(outlier.color = "darkred", outlier.shape = 16, alpha = 0.7) +
+  scale_fill_manual(
+    values = c("Low Risk" = "#603601", "Moderate Risk" = "#3a6a91", "High Risk" = "#cfb886")
+  ) +
+  labs(
+    x = "Risk Category",
+    y = "Risk Index"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+```
+
+@fig-risk-index-boxplot highlights the distribution of the Risk Index across three defined categories: High Risk, Low Risk and Moderate Risk.
+Each category represents a grouping of neighborhoods based on their average Risk Index values.
+The plot reveals distinct differences in the distribution of the Risk Index between these categories.
+
+The High Risk category exhibits a higher median Risk Index, with a relatively narrow interquartile range (IQR), indicating consistent high-risk values across neighborhoods in this group.
+Conversely, the Low Risk category has a significantly lower median, with a similarly narrow IQR, suggesting stable low-risk conditions in these neighborhoods.
+The Moderate Risk category shows greater variability in its distribution, with a wider IQR and overlapping values with both High Risk and Low Risk categories.
+This suggests that neighborhoods in the Moderate Risk group exhibit diverse risk profiles, likely influenced by varying environmental and situational factors.
 
@@
 
+The predictor variables in this study are organized into two distinct models: Theft Index and Collision Probability, each designed to capture and explain critical aspects of theft and collision risks, respectively.
+@fig-tree provide a visual overview of the hierarchical relationships between the predictor variables and the overall risk framework, offering a structured understanding of the factors contributing to these incidents
+
+```{r}
+#| label: fig-tree
+#| fig-cap: "Hierarchical relationship between risk factors contributing to theft and collision probabilities."
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8 # Increased width for better text display
+#| fig-height: 4  # Adjusted height for spacing
+
+# Load required libraries
+library(tidygraph)
+library(ggraph)
+
+# Define the tree structure as a dataframe
+tree_data <- data.frame(
+  from = c("Risk Index", "Risk Index", "Theft Index", "Theft Index", "Theft Index",
+           "Collision Probability", "Collision Probability", "Collision Probability", 
+           "Collision Probability", "Collision Probability", "Collision Probability"),
+  to = c("Theft Index", "Collision Probability", "Hour of the Day", "Day of the Week", 
+         "Premises Type", "Neighborhood ID", "Road Surface Condition", 
+         "Lighting Condition", "Traffic Control", "Time of Day", "Road Visibility")
+)
+
+# Convert to graph object
+tree_graph <- as_tbl_graph(tree_data)
+
+# Plot the tree diagram with horizontal layout and curved lines
+ggraph(tree_graph, layout = "dendrogram", circular = FALSE) +
+  geom_edge_bend(color = "gray", curvature = 0.5, arrow = arrow(length = unit(4, "mm")), end_cap = circle(4, "mm")) +
+  geom_node_point(size = 6, color = "grey") +
+  geom_node_text(
+    aes(label = name, hjust = ifelse(name == "Risk Index", 1.3, -0.2)), # Conditional hjust
+    vjust = -0.5, size = 4, color = "black", angle = 0
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 16, face = "bold"),
+    axis.title = element_blank(),
+    axis.text = element_blank(),
+    axis.ticks = element_blank(),
+    panel.grid = element_blank()
+  ) +
+  coord_flip()
+
+```
+
 ### Theft index
 
+The Theft Index focuses on the temporal and locational characteristics that influence theft occurrences.
+By incorporating variables such as Hour of the Day, Day of the Week, and Premises Type, this model identifies patterns tied to specific times and locations, highlighting when and where thefts are most likely to occur.
+
 #### Hour of the Day
+
 #### Day of the Week
+
 #### Premises Type
@@
 
+The Collision Probability model, on the other hand, examines situational and environmental conditions affecting the likelihood of collisions.
+Variables such as Neighborhood ID, Road Surface Condition, Lighting Condition, Traffic Control, Time of Day, and Visibility offer insights into the contextual factors that contribute to traffic incidents, capturing the dynamic interplay between environmental conditions and human interactions.
+
 #### Neighborhood ID
+
 #### Road Surface Condition
+
 #### Lighting Condition
+
 #### Traffic Control
+
 #### Time of Day
+
 #### Visibility
@@
 
-The total theft component is calculated as:
-\begin{align*}
+The total theft component is calculated as: \begin{align*}
 C_{\text{Theft}} &= \text{Hour Index} + \text{Day Index} + \text{Premises Type Index}
@@
 ### Collision Probability Model
+
 A logistic regression model was used to predict the likelihood of severe collisions P(Collision) based on several predictors. The log-odds of the collision probability are modeled as:
-\
-
+
+```{=tex}
 \begin{align*}
@@
 \end{align*}
-
+```
 The model prediction utilizes the following predictor variables:
@@
 
+```{=tex}
 \begin{align*}
@@
 \end{align*}
-
+```
 ### Risk Index Calculation
@@
 \end{align*}
-
-Weights are defined as:
-\begin{align*}
+```
+Weights are defined as: \begin{align*}
 w_1 = 0.7, \quad w_2 = 0.3
-\end{align*}
-These reflect the relative importance of collision and theft risks, emphasizing collision severity due to its greater immediate impact.
+\end{align*} These reflect the relative importance of collision and theft risks, emphasizing collision severity due to its greater immediate impact.
 
@@
 
-### Theft Sub-Index [add one sentence of validation]
+### Theft Sub-Index \[add one sentence of validation\]
 
@@
 
-### Composite Risk Index [Add one sentence of assumption]
+### Composite Risk Index \[Add one sentence of assumption\]
 
@@
 
-In the collision model:
-- The model estimates the probability of a severe collision (\(P(\text{Collision})\)) given environmental and situational predictors. The response variable (`severity`) is defined as:
-  \begin{align*}
+In the collision model: - The model estimates the probability of a severe collision ((P(\text{Collision}))) given environmental and situational predictors.
+The response variable (`severity`) is defined as: \begin{align*}
   P(\text{Collision}) = \frac{\exp(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_k X_k)}{1 + \exp(\beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_k X_k)}
-  \end{align*}
-  where:
-  \begin{align*}
+  \end{align*} where: \begin{align*}
   \beta_0 &\text{ is the intercept,} \\
@@
 \end{align*}
-
+```
 where P represents the predicted probability of collision from the logistic regression model plotted against the continuous predictors.
@@
 #| fig-pos: H
+#| fig-width: 8
+#| fig-height: 3.5
+
 # Load necessary libraries
@@
   geom_point(alpha = 0.5, color = "darkgrey") +
-  geom_smooth(method = "loess", color = "blue", fill = "lightblue") +
+  geom_smooth(method = "loess", color = "darkred", fill = "lightblue") +
   coord_cartesian(ylim = c(-5, 5)) +  # Restrict the y-axis range
@@
 Visibility Conditions directly reflects environmental factors like fog, heavy rain, or low light, which impair drivers' ability to see hazards.
-Conversely, Road Conditions` account for the physical state of the driving surface, such as wet, icy, or damaged roads, which influence vehicle handling and stopping distance.
+Conversely, Road Conditions\` account for the physical state of the driving surface, such as wet, icy, or damaged roads, which influence vehicle handling and stopping distance.
 Together, these variables capture complementary aspects of collision risk, ensuring that the model provides a comprehensive assessment.
@@
 
-Time-of-Day and Day-of-Week Analysis of the Risk Index
-
-Plot Ideas:
-Bar: Show Risk Index variations across hours of the day and days of the week, highlighting theft and collision components separately.
-
-Time Series Plot: Display Risk Index trends over a longer period, revealing potential seasonal effects.
+Understanding the temporal dynamics of the Risk Index provides critical insights into when risks are most elevated, helping to inform motorbike owners and policymakers about high-risk periods.
+This section examines variations in the Risk Index over time of a day, and utilizes a time series analysis to reveal potential seasonal effects and long-term trends.
+These analyses provide a comprehensive view of how risks evolve temporally, highlighting critical periods for intervention.
+
+### Time-of-Day Analysis
 
 ```{r}
-```
+#| label: fig-risk-index-timeofday
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Temporal Trends: Risk Index Variations by Hour of the Day"
+#| fig-pos: H
+#| fig-width: 8
+#| fig-height: 4
+
+# Load necessary data
+final_risk_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+collision_risk_scores <- read_csv(here("data", "02-analysis_data", "collision_risk_scores.csv"))
+
+# Convert 'id' to character in both datasets if necessary
+final_risk_data$id <- as.character(final_risk_data$id)
+cleaned_traffic_data$id <- as.character(cleaned_traffic_data$id)
+collision_risk_scores$id <- as.character(collision_risk_scores$id)
+
+# Merge the final risk index with the traffic data based on 'id'
+merged_data <- left_join(final_risk_data, cleaned_traffic_data, by = "id")
+
+# Merge the collision risk scores into the merged data
+merged_data <- left_join(merged_data, collision_risk_scores, by = "id")
+
+# Create 'time_of_day_bin' based on 'time_of_day' in cleaned_traffic_data
+merged_data$time_of_day_bin <- case_when(
+  merged_data$time_of_day >= 0 & merged_data$time_of_day < 600 ~ "Early Morning",
+  merged_data$time_of_day >= 600 & merged_data$time_of_day < 1200 ~ "Morning",
+  merged_data$time_of_day >= 1200 & merged_data$time_of_day < 1800 ~ "Afternoon",
+  merged_data$time_of_day >= 1800 & merged_data$time_of_day < 2400 ~ "Evening"
+)
+
+# Group by 'time_of_day_bin' and calculate average values for risk index and collision index
+summary_data <- merged_data %>%
+  group_by(time_of_day_bin) %>%
+  summarize(
+    avg_risk_index = mean(risk_index, na.rm = TRUE),
+    avg_collision_index = mean(collision_probability
+, na.rm = TRUE)
+  )
+
+# Theft hour indexes (24-hour data points)
+theft_hour_indexes <- c(0.005335554, 0.003690682, 0.002343943, 0.002158895, 0.001572909,
+                        0.002415906, 0.006353319, 0.012295419, 0.018587055, 0.02577309,
+                        0.026174027, 0.025526359, 0.025248787, 0.024107657, 0.021969323,
+                        0.020334731, 0.018196398, 0.016448721, 0.016479562, 0.015400115,
+                        0.013806645, 0.01254215, 0.009211284, 0.007360803)
+
+# Create a mapping of the theft hour indexes to the time of day bins
+theft_hour_bins <- c("Early Morning", "Early Morning", "Early Morning", "Early Morning", "Early Morning", 
+                     "Morning", "Morning", "Morning", "Morning", "Morning", 
+                     "Morning", "Morning", "Morning", "Afternoon", "Afternoon", 
+                     "Afternoon", "Afternoon", "Afternoon", "Afternoon", 
+                     "Evening", "Evening", "Evening", "Evening", "Evening")
+
+# Create a dataframe for the theft data, with time_of_day_bin mapped accordingly
+theft_hour_data <- data.frame(
+  time_of_day_bin = theft_hour_bins,
+  theft_index = theft_hour_indexes
+)
+
+# Summarize theft data by time_of_day_bin to match the bin structure used in the plot
+theft_summary_data <- theft_hour_data %>%
+  group_by(time_of_day_bin) %>%
+  summarize(avg_theft_index = mean(theft_index))
+
+# Merge with existing summary data for risk and collision indexes
+final_plot_data <- left_join(summary_data, theft_summary_data, by = "time_of_day_bin")
+
+# Plot: Risk Index, Collision Index, and Theft Hour Index by Time of Day Bin
+ggplot(final_plot_data, aes(x = time_of_day_bin)) +
+  geom_bar(aes(y = avg_risk_index), stat = "identity", fill = "#967145") +
+  geom_line(aes(y = avg_collision_index), group = 1, color = "darkred", size = 1) + # Collision index line
+  geom_text(aes(y = avg_risk_index, label = round(avg_risk_index, 2)), vjust = -0.5) +  # Add labels for risk index
+  geom_text(aes(y = avg_collision_index, label = round(avg_collision_index, 2)), vjust = 1.5, color = "black") +  # Add labels for collision index
+  geom_line(aes(y = avg_theft_index * 10), color = "green", size = 1, linetype = "dashed") +  # Theft index line (scaled by 100)
+  geom_point(aes(y = avg_theft_index * 10), color = "green", size = 2) +  # Theft index points
+  labs(
+
+    x = "Time of Day Bin",
+    y = "Average Risk Index"
+  ) +
+  theme_minimal()
+
+```
+
+@fig-risk-index-timeofday illustrates how the total Risk Index varies throughout the day.
+This Risk Index is composed of two key components: collision risk and theft risk.
+The bar heights represent the overall Risk Index, while the red line indicates collision risk and the green line reflects theft risk.To better visualize the theft risk, the values have been multiplied by 10, as they are typically quite small (less than 1/3 of the total Hour Risk Index throughout the day).
+
+The collision risk (red line) shows significant peaks during commuting hours—between 7–9 AM and 5–7 PM—corresponding to high traffic volumes, which indicates a direct relationship between collision risk and traffic patterns.
+The theft risk (green line), which has been scaled for better visualization, remains consistently low throughout the day, showing no significant variation across time bins.
+This suggests that theft risk is relatively stable, regardless of the time of day.
+The total Risk Index, represented by the bar heights, shows slight fluctuations, with the highest values occurring during morning hours.
+This is likely due to the heightened collision risk in the morning, which has a greater influence on the overall risk during these periods.
+
+### Time Series Analysis
+
+This time series analysis explores how the Risk Index evolves over time, offering a long-term view across the years and a more focused examination of a specific year.
+By visualizing these temporal trends, we can reveal potential seasonal effects, fluctuations, and patterns that might not be immediately apparent in a snapshot of data.
+The following section presents the trends in the Risk Index over both the full timespan (2006-2021) and the specific year 2020.
+
+#### Longterm
 
 ```{r}
-
-```
+#| label: fig-risk-index-timeseries-longterm
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Temporal Trends: Risk Index Variations Over Time with Rolling Average"
+#| fig-pos: H
+#| fig-width: 8
+#| fig-height: 3.5
+
+# Load necessary libraries
+library(tidyverse)
+library(lubridate)
+library(here)
+
+# Load the cleaned traffic data
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+# Load the final risk index if not already in `cleaned_traffic_data`
+if (!"risk_index" %in% colnames(cleaned_traffic_data)) {
+  final_risk_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+  cleaned_traffic_data <- cleaned_traffic_data %>%
+    left_join(final_risk_data, by = "id")
+}
+
+# Convert `date` to Date format if not already
+cleaned_traffic_data <- cleaned_traffic_data %>%
+  mutate(date = as.Date(report_date))
+
+# Aggregate data by date to calculate the average Risk Index
+risk_trends <- cleaned_traffic_data %>%
+  group_by(date) %>%
+  summarize(avg_risk_index = mean(risk_index, na.rm = TRUE))
+
+# Create a 6-month rolling average for clearer trend
+risk_trends <- risk_trends %>%
+  arrange(date) %>%
+  mutate(rolling_avg = zoo::rollmean(avg_risk_index, 6, fill = NA, align = "center"))
+
+# Create the time series plot for the entire period
+ggplot(risk_trends, aes(x = date, y = avg_risk_index)) +
+  geom_line(color = "#967145", size = 1, alpha = 0.7) +
+  geom_line(aes(x = date, y = rolling_avg), color = "grey", size = 1, linetype = "dashed") +
+  labs(
+
+    x = "Year",
+    y = "Average Risk Index"
+  ) +
+  scale_x_date(
+    date_breaks = "2 years", 
+    date_labels = "%Y"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 12)
+  )
+
+```
+
+@fig-risk-index-timeseries-longterm provides a broad overview of the Risk Index over the entire period from 2006 to 2021, with the average Risk Index for each day represented by the blue line.
+To facilitate clearer interpretation, a 6-month rolling average is included, shown by the dashed red line.
+The plot highlights long-term trends in risk levels, showing periods of heightened and reduced risk.
+Although there are some fluctuations, the rolling average smooths these variations, enabling us to observe the general trend over time.
+The x-axis labels every three years for better clarity, making it easier to track shifts in risk levels.
+From this, we can discern if certain years or periods saw notable increases or decreases in the Risk Index, possibly suggesting underlying external factors like policy changes or events that affected the risk.
+
+#### Year 2020
+
+```{r}
+#| label: fig-risk-index-timeseries-year
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Risk Index Variations in 2020"
+#| fig-pos: H
+#| fig-width: 8.5
+#| fig-height: 3.5
+
+# Filter data for a specific year (e.g., 2020)
+risk_trends_2020 <- cleaned_traffic_data %>%
+  filter(year(date) == 2020) %>%
+  group_by(date) %>%
+  summarize(avg_risk_index = mean(risk_index, na.rm = TRUE))
+
+risk_trends_2020 <- risk_trends_2020 %>%
+  mutate(month = month(date, label = TRUE))
+
+ggplot(risk_trends_2020, aes(x = date, y = avg_risk_index)) +
+  geom_line(color = "#66471d", size = 0.8) +
+  facet_wrap(~month, scales = "free_x") +  # Facet by month
+  labs(
+    x = "Month",
+    y = "Average Risk Index"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
+    axis.title = element_text(size = 12)
+  )
+
+```
+
+@fig-risk-index-timeseries-year provides a detailed, month-by-month breakdown of the Risk Index for the year 2020.
+This faceted layout offers a clear view of short-term fluctuations and potential seasonal patterns by presenting trends for each month separately.
+The graph highlights variations in risk levels across the months, making it easier to identify distinct periods of higher or lower risk.
+These trends may correspond to factors such as weather conditions, holidays, or other contextual events.
+For instance, spikes or dips in risk within specific months could reflect changes in traffic volume, seasonal behaviors, or other dynamic factors influencing motorbike-related incidents throughout the year.
 
@@
 
-Exploring How Environmental and Situational Factors Impact the Risk Index
-
-Plot Ideas:
-Violin Plot: Show the Risk Index distribution under varying road surface conditions (e.g., dry, wet, icy).
-
-Faceted Plot: Break down the Risk Index by lighting conditions (e.g., daylight, artificial light) or traffic control types.
+Understanding how environmental and situational factors contribute to variations in the Risk Index is crucial for identifying the conditions under which motorbike risks are heightened.
+This section explores how different factors, such as road surface conditions, lighting conditions, and traffic control types, influence the Risk Index.
+By breaking down the data into categories based on these factors, we can reveal specific conditions that might increase the likelihood of accidents or theft.
+The following visualizations provide insight into the distribution of risk under varying environmental conditions, highlighting the factors that contribute to motorbike-related incidents.
+
+To better understand how environmental factors influence the Risk Index, we examine the distribution of the Risk Index under various road surface conditions.
+This analysis explores whether different surface types—such as dry, wet, icy, and others—affect the level of risk associated with traffic incidents.
+The following violin plot visually compares the Risk Index across different road surface conditions, providing insight into how these factors may contribute to the overall risk.
 
 ```{r}
-
-```
+#| label: fig-risk-index-lighting-conditions
+#| fig-cap: "Risk Index by Lighting Conditions"
+#| fig-subcap: ["Low Light (Dark & Artificial)", "Dawn and Artificial Light", "Bright Daylight and Artificial Light", "Other Lighting Conditions"]
+#| echo: false
+#| warning: false
+#| message: false
+#| layout-ncol: 2
+#| layout-nrow: 2
+
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+
+# Load the data
+final_risk_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+collision_risk_scores <- read_csv(here("data", "02-analysis_data", "collision_risk_scores.csv"))
+
+# Convert 'id' to character in both datasets if necessary
+final_risk_data$id <- as.character(final_risk_data$id)
+cleaned_traffic_data$id <- as.character(cleaned_traffic_data$id)
+collision_risk_scores$id <- as.character(collision_risk_scores$id)
+
+# Merge the final risk index with the traffic data based on 'id'
+merged_data <- left_join(final_risk_data, cleaned_traffic_data, by = "id")
+
+# Merge the collision risk scores into the merged data
+merged_data <- left_join(merged_data, collision_risk_scores, by = "id")
+
+# Preprocess lighting conditions: Group 'NA' as 'Other' and create lighting condition categories for the full dataset
+merged_data$lighting_conditions <- as.character(merged_data$lighting_conditions)  # Ensure it's a character vector
+
+# Convert "NA" (actual value in the data) to "Other"
+merged_data$lighting_conditions[merged_data$lighting_conditions == "NA"] <- "Other"
+
+# Group lighting conditions into four categories, including "Other"
+merged_data$lighting_condition_group <- case_when(
+  merged_data$lighting_conditions %in% c("Dark", "Dark, Artificial") ~ "Low Light (Dark & Artificial)",
+  merged_data$lighting_conditions %in% c("Dawn", "Dawn, Artificial") ~ "Dawn and Artificial Light",
+  merged_data$lighting_conditions %in% c("Daylight", "Daylight, Artificial") ~ "Bright Daylight and Artificial Light",
+  merged_data$lighting_conditions == "Other" ~ "Other Lighting Conditions"
+)
+
+# First Plot: Low Light (Dark & Artificial)
+plot_dark <- ggplot(merged_data %>% filter(lighting_condition_group == "Low Light (Dark & Artificial)"), aes(x = risk_index, fill = lighting_condition_group)) +
+  geom_density(alpha = 0.7) +
+  scale_fill_manual(values = c("Low Light (Dark & Artificial)" = "#54350a")) +  # Custom color
+  labs(x = "Risk Index", y = "Density", title = "Low Light (Dark & Artificial)") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Second Plot: Dawn and Artificial Light
+plot_dawn <- ggplot(merged_data %>% filter(lighting_condition_group == "Dawn and Artificial Light"), aes(x = risk_index, fill = lighting_condition_group)) +
+  geom_density(alpha = 0.6) +
+  scale_fill_manual(values = c("Dawn and Artificial Light" = "#54350a")) +  # Custom color
+  labs(x = "Risk Index", y = "Density", title = "Dawn and Artificial Light") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Third Plot: Bright Daylight and Artificial Light
+plot_daylight <- ggplot(merged_data %>% filter(lighting_condition_group == "Bright Daylight and Artificial Light"), aes(x = risk_index, fill = lighting_condition_group)) +
+  geom_density(alpha = 0.6) +
+  scale_fill_manual(values = c("Bright Daylight and Artificial Light" = "#54350a")) +  # Custom color
+  labs(x = "Risk Index", y = "Density", title = "Bright Daylight and Artificial Light") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Fourth Plot: Other Lighting Conditions
+plot_other <- ggplot(merged_data %>% filter(lighting_condition_group == "Other Lighting Conditions"), aes(x = risk_index, fill = lighting_condition_group)) +
+  geom_density(alpha = 0.6) +
+  scale_fill_manual(values = c("Other Lighting Conditions" = "#54350a")) +  # Custom color
+  labs(x = "Risk Index", y = "Density", title = "Other Lighting Conditions") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Display the four plots side by side
+plot_dark
+plot_dawn
+plot_daylight
+plot_other
+```
+
+@fig-risk-index-lighting-conditions display the distribution of the Risk Index across various lighting conditions.
+The data is broken down into four main lighting categories: Low Light (Dark & Artificial), Dawn and Artificial Light, Bright Daylight and Artificial Light, and Other Lighting Conditions.
+
+@fig-risk-index-lighting-conditions-1 highlights the risk distribution during low-light conditions, including both "Dark" and "Dark, Artificial" categories.
+The density curve shows that motorbike risks are relatively higher in these conditions, particularly in the moderate-to-high range of the Risk Index, indicating that poorer visibility and the presence of artificial lighting increase the likelihood of incidents.
+
+@fig-risk-index-lighting-conditions-2 captures the risk profile during dawn and artificial lighting.
+The distribution shifts slightly, showing a lower concentration of higher risk indices compared to the low-light category.
+While there is still some risk during dawn, the overall risk index appears to be less extreme compared to the "Low Light" category, possibly due to the improving visibility as the sun rises.
+
+@fig-risk-index-lighting-conditions-3 shows a much lower density of high-risk events.
+This suggests that the highest risk does not generally occur during daylight hours, as visibility is much better, and traffic conditions tend to be more predictable.
+
+@fig-risk-index-lighting-conditions-4 captures a small number of cases that don’t fit into the above categories.
+The risk distribution in this category is less pronounced, likely because it includes a variety of edge cases, such as unusual lighting conditions or missing data.
+
+## Mapping the Risk: Neighborhood-Level Insights
+
+### Neighborhood Risk Distribution
 
 ```{r}
-
-```
-
-## Mapping the Risk: Neighborhood-Level Insights
-
-Spatial Patterns in the Distribution of the Risk Index
-
-Plot Ideas:
-Choropleth Map: Display the Risk Index across neighborhoods with color gradients indicating risk levels.
-
-Bar Chart: Rank neighborhoods by their average Risk Index to identify the most vulnerable areas.
+#| label: fig-his_collosion_risk
+#| fig-cap: "Neighborhoods grouped by average collision probabilities into High, Medium, and Low risk categories."
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 7
+#| fig-height: 3.5
+#|
+
+# Load necessary libraries
+library(tidyverse)
+library(hrbrthemes)
+
+# Load the dataset
+data <- read.csv(here("data", "02-analysis_data", "collision_risk_scores.csv"))
+
+# Take the average for duplicate neighborhoods
+data <- data %>%
+  group_by(hood_158) %>%
+  summarise(collision_probability = mean(collision_probability, na.rm = TRUE)) %>%
+  ungroup()
+
+# Create groups: High, Medium, Low
+data <- data %>%
+  mutate(
+    risk_group = case_when(
+      collision_probability >= 0.8 ~ "High",
+      collision_probability >= 0.6 ~ "Medium",
+      TRUE ~ "Low"
+    )
+  )
+
+# Adjusted theme to avoid font issues
+ggplot(data, aes(x = collision_probability, color = risk_group, fill = risk_group)) +
+  geom_histogram(alpha = 0.8, binwidth = 0.05) +
+  scale_fill_manual(values = c("High" = "#54350a", "Medium" = "#2e3f78", "Low" = "#d1b33d")) +
+  scale_color_manual(values = c("High" = "#54350a", "Medium" = "#2e3f78", "Low" = "#d1b33d")) +
+  theme_minimal() +
+  theme(
+    legend.position = "right",
+    panel.spacing = unit(0.1, "lines"),
+    strip.text.x = element_text(size = 8)
+  ) +
+  xlab("Collision Probability") +
+  ylab("Number of Neighborhoods") +
+  labs(
+    fill = "Risk Group",
+    color = "Risk Group"
+  ) +
+  facet_wrap(~risk_group)
+
+```
+
+@fig-his_collosion_risk illustrates the distribution of neighborhoods in Toronto grouped by their average collision probabilities into three risk categories: High, Medium, and Low.
+Each panel represents one of the risk categories and shows the range of collision probabilities (x-axis) and the corresponding number of neighborhoods (y-axis) in that category.
+For example, the High-risk group predominantly consists of neighborhoods with collision probabilities between 0.8 and 1.0, indicating a clustering of high probabilities in these areas.
+The Medium-risk group primarily includes neighborhoods with probabilities between 0.6 and 0.8, while the Low-risk group has neighborhoods with probabilities below 0.6.
+
+This visualization aligns with the study's focus on spatial patterns, as it highlights that collision risks are not evenly distributed across neighborhoods.
+The results suggest targeted interventions could be more effective when tailored to the risk level of specific areas, such as road safety measures in neighborhoods with high collision probabilities.
+
+### Maps
 
 ```{r}
-
-```
-
-```{r}
-
-```
+#| label: fig-collision_map
+#| fig-cap: "Neighborhood Collision Points in Toronto Highlighting High Collision Risk Areas"
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8
+#| fig-height: 5
+
+# Load necessary libraries
+library(sf)
+library(ggplot2)
+library(osmdata)
+library(dplyr)
+
+# Load the GeoJSON file
+collision_data <- st_read(here("data", "02-analysis_data", "collision.geojson"), quiet = TRUE)
+
+# Add a column to classify top 15 neighborhoods
+top_15_hoods <- c(173, 88, 29, 61, 89, 84, 153, 53, 79, 157, 150, 12, 108, 143, 98)
+collision_data <- collision_data %>%
+  mutate(
+  is_top_15 = ifelse(as.numeric(HOOD_158) %in% top_15_hoods, "High Collision Risk Neighborhoods", "Others")
+)
+
+# Get a basemap of Toronto from OpenStreetMap
+toronto_bbox <- c(-79.6393, 43.5810, -79.1151, 43.8555)  # Toronto bounding box
+toronto_map <- opq(bbox = toronto_bbox) %>%
+  add_osm_feature(key = "landuse") %>%
+  osmdata_sf()
+
+# Verify CRS and transform if needed
+collision_data <- st_transform(collision_data, crs = 4326)
+
+# Update basemap query
+toronto_map <- opq(bbox = toronto_bbox) %>%
+  add_osm_feature(key = "landuse", value = "residential") %>%
+  osmdata_sf()
+
+# Create the plot
+ggplot() +
+  # Add basemap
+  geom_sf(data = toronto_map$osm_polygons, fill = "gray90", color = "gray70") +
+  # Add collision points with classification
+  geom_sf(data = collision_data, aes(color = is_top_15), size = 0.5, alpha = 0.6) +
+  scale_color_manual(values = c("High Collision Risk Neighborhoods" = "red", "Others" = "yellow")) +
+  labs(
+    x = "Longitude",
+    y = "Latitude",
+    color = "Neighborhood Type",
+    title = "Neighborhood Collision Points in Toronto"
+  ) +
+  theme_minimal()
+
+```
+
+@fig-collision_map illustrates the distribution of collision points across Toronto, emphasizing neighborhoods categorized as "High Collision Risk" in red and other areas in yellow.
+The spatial extent spans latitudes from approximately 43.55°N to 43.85°N and longitudes from -79.6°W to -79.1°W, covering most of Toronto's urban landscape.
+
+The high collision risk neighborhoods are clustered primarily in specific areas, as denoted by the red points.
+These regions often correspond to densely populated or high-traffic zones, where the likelihood of traffic collisions is significantly elevated.
+In contrast, the yellow points represent areas with lower collision risks, which are more dispersed across the map, particularly in less congested parts of the city.
+
+This map provides a critical spatial perspective, highlighting the geographic disparities in traffic collision risks within Toronto.
+It underscores the need for targeted safety interventions, such as traffic calming measures and enhanced infrastructure, in the high-risk neighborhoods.
+The visualization also serves as a tool for urban planners and policymakers to identify and prioritize areas where safety improvements could significantly reduce the frequency and severity of traffic collisions, fostering a safer urban environment.
 
@@
 
-Key Points:
-Discuss high-risk times of day or days of the week for both theft and collisions.
+Key Points: Discuss high-risk times of day or days of the week for both theft and collisions.
 Recommendations for motorbike owners to reduce risk exposure, such as avoiding specific time periods or enhancing security measures during peak theft hours.
@@
 
-Key Points:
-Discuss the significant environmental predictors of risk, such as poor road surface conditions or inadequate lighting.
+Key Points: Discuss the significant environmental predictors of risk, such as poor road surface conditions or inadequate lighting.
 Recommend urban planning measures, such as improved road maintenance or installation of streetlights, to mitigate risks.
@@
 
-Key Points:
-Prioritizing high-risk neighborhoods for targeted interventions, such as improved street lighting or traffic control measures.
+Key Points: Prioritizing high-risk neighborhoods for targeted interventions, such as improved street lighting or traffic control measures.
 Suggestions for law enforcement strategies to reduce theft in hotspots.
@@
 
-Future Work3: Investigate the correlation between neighborhood-level Risk Index values and broader social or economic indicators, such as income levels or population density.
-\newpage
+Future Work3: Investigate the correlation between neighborhood-level Risk Index values and broader social or economic indicators, such as income levels or population density. \newpage
 
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 
-Demographics:
-What is your age group? (e.g., 18–24, 25–34, etc.)
-What is your highest level of education completed?
-
-Riding Behavior:
-How often do you ride your motorbike? (Daily, Weekly, Monthly, Rarely)
-For what purposes do you primarily use your motorbike? (Commuting, Recreation, Delivery/Work, Other)
-
-Risk Awareness and Perception:
-On a scale of 1–5, how would you rate the theft risk in your neighborhood?
-Have you experienced a motorbike theft or collision in the past year? (Yes/No)
-
-Situational Factors:
-What time of day do you usually ride? (Morning, Afternoon, Evening, Night)
-In what weather conditions do you typically ride? (Clear, Rainy, Snowy)
-
-Preventive Measures:
-What measures do you take to secure your motorbike against theft? (Select all that apply: Lock, Alarm, GPS Tracker, Parking in Secure Locations)
-Do you avoid specific times or areas due to perceived collision risks? (Yes/No)
+Demographics: What is your age group? (e.g., 18–24, 25–34, etc.) What is your highest level of education completed?
+
+Riding Behavior: How often do you ride your motorbike? (Daily, Weekly, Monthly, Rarely) For what purposes do you primarily use your motorbike? (Commuting, Recreation, Delivery/Work, Other)
+
+Risk Awareness and Perception: On a scale of 1–5, how would you rate the theft risk in your neighborhood? Have you experienced a motorbike theft or collision in the past year? (Yes/No)
+
+Situational Factors: What time of day do you usually ride? (Morning, Afternoon, Evening, Night) In what weather conditions do you typically ride? (Clear, Rainy, Snowy)
+
+Preventive Measures: What measures do you take to secure your motorbike against theft?
+(Select all that apply: Lock, Alarm, GPS Tracker, Parking in Secure Locations) Do you avoid specific times or areas due to perceived collision risks?
+(Yes/No)
 
@@
 
-Research on environmental and situational predictors of road accidents.
-Studies on the effectiveness of theft prevention measures.
-Literature on sampling methods and survey design for risk assessment.
+Research on environmental and situational predictors of road accidents. Studies on the effectiveness of theft prevention measures. Literature on sampling methods and survey design for risk assessment.
 

### 2024-12-02 paper/paper.qmd
almost finalized with the model and result section
initial draft of the variables section

@@
 date-format: long
-abstract: "This study examines motor vehicle theft and traffic collisions in Toronto, focusing on spatial and temporal patterns, recovery outcomes, and contributing factors.
-[The analysis reveals that thefts are geographically clustered, with recovery rates differing by location type, while collisions are more frequent under poor visibility and adverse road conditions.] These findings underscore the need for targeted interventions, including enhanced surveillance in high-theft areas, infrastructure improvements for road safety, and optimized strategies for vehicle recovery.
-Such measures are critical for fostering safer and more secure urban mobility in Toronto."
+abstract: "This study examines motor vehicle theft and traffic collisions in Toronto, focusing on spatial and temporal patterns,and environmental factors.
+A composite risk score model is used and reveals that collisions occur more frequently under unfavorable road conditions, such as poor lighting and wet or icy surfaces, and in areas with inadequate traffic control measures.
+These findings emphasize the need for targeted interventions, including enhanced road infrastructure and lighting in collision-prone zones.
+Such measures aim to improve safety and security for motorbike riders and owners, contributing to safer urban mobility in Toronto."
 format: pdf
@@
 Using data from Open Data Toronto, the model evaluates two critical events: motorbike theft and collisions.
-Logistic regression is employed to estimate the probabilities of these events, which are then combined into a single, interpretable composite risk score.
-This metric, incorporating factors such as neighborhood characteristics, road and lighting conditions, and time of day, is designed to guide motorbike users, insurers, and policymakers in risk assessment and decision-making, while informing strategies to mitigate these risks.
+Logistic regression is employed to estimate the probabilities of motor collisions, and a theft index is calculated which are then combined with the likelihood of motor collisions into a single, interpretable composite risk score.
+This metric is designed to guide motorbike users, insurers, and policymakers in risk assessment and decision-making, while informing strategies to mitigate these risks.
 
@@
 
-This analysis confirms and extends three key findings: (1) material properties significantly influence motor efficiency, with high-conductivity materials yielding higher energy savings; (2) geometric optimization of motor components enhances torque generation and reduces energy losses; and (3) operational conditions, such as load and temperature, exhibit nonlinear effects, underscoring the need for adaptive motor control strategies.
-Furthermore, our projections suggest that integrating advanced materials and control systems could improve motor efficiency by up to 15% by 2050, aligning with global energy transition goals \[To be updated..\].
+This analysis confirms and extends three key findings: (1) theft risks vary based on temporal factors, such as time of day and day of the week, with mornings exhibiting higher susceptibility to theft, while early mornings have lower theft risks; (2) Collision risks are strongly influenced by environmental and situational conditions, with risks being particularly high under poor visibility.
+Factors such as wet or icy road surfaces and inadequate traffic controls further significantly increase the likelihood of incidents; and (3) neighborhood characteristics play a critical role in shaping both theft and collision risks, with high-collision-risk neighborhoods primarily clustered in high-traffic zones and a notable concentration of collision incidents in downtown Toronto, reflecting the impact of dense urban environments and heavy traffic flows on risk levels.
 
 The structure of the paper is organized as follows: following @sec-intro, @sec-data outlines the data collection and preprocessing procedures, along with a detailed description of the outcome variable and the predictor variables used in the analysis.
-@sec-model introduces the logistic regression models applied to estimate the probabilities of theft and collision, as well as the method used to combine these probabilities into a composite risk score.
+@sec-model introduces the logistic regression models applied to estimate the probability of collision, as well as the method used to derive the theft index and combine these probabilities into a composite risk score.
 @sec-result then presents the main findings, including insights into how different factors contribute to the risks of owning and riding a motorbike.
@@
 
-This study utilized two datasets published by the Toronto Police Service.
+This study utilized two datasets published by the Toronto Police Service, available from Open Data Toronto [@OpendataToronto].
 The first dataset focuses on motor vehicle collisions involving killed or seriously injured persons (KSI), while the second examines thefts from motor vehicles.
@@
 By combining these elements, the Risk Index provides a unified measure of risk, enabling the identification of high-risk scenarios and areas.
-This section examines the distribution of the Risk Index through visualizations to reveal key patterns, variations, and potential drivers of motorbike-related risks.
-Two graphs are presented: (1) a histogram with density overlay to illustrate the overall distribution of the Risk Index, and (2) a boxplot by neighborhood to highlight spatial differences in risk.
 
@@
 @fig-risk-index-histogram shows the distribution of the Risk Index across all observations in the dataset.
-The Risk Index exhibits a unimodal distribution, skewed slightly to the left, with the majority of values concentrated between 0.45 and 0.65.
+The Risk Index shows a unimodal distribution, skewed slightly to the left, with the majority of values concentrated between 0.45 and 0.65.
 This indicates that most motorbike-related risks fall within a moderate range.
@@
 The skewness and spread of the distribution highlight the variability in the combined risks of theft and collision.
-The extended tail on the higher end of the Risk Index suggests that certain environmental or situational factors disproportionately elevate risks in specific cases.
-This insight can inform targeted interventions, focusing on the outliers with elevated Risk Index values to mitigate the most critical risks.
-
-Building on the distribution of the Risk Index shown in the histogram, this box plot further dissects the data by grouping neighborhoods into risk categories based on their average Risk Index values.
-Neighborhoods are classified into three categories—Low Risk, Moderate Risk, and High Risk—using quartiles.
+The extended tail on the higher end of the Risk Index suggests that certain environmental or situational factors disproportionately increase the risks in specific cases.
+This insight can inform targeted interventions, focusing on the outliers with high Risk Index values to mitigate the most critical risks.
 
@@
 
-@fig-risk-index-boxplot highlights the distribution of the Risk Index across three defined categories: High Risk, Low Risk and Moderate Risk.
+Building on the distribution of the Risk Index in @fig-risk-index-histogram, @fig-risk-index-boxplot highlights the distribution of the Risk Index across three defined categories: High Risk, Low Risk and Moderate Risk.
 Each category represents a grouping of neighborhoods based on their average Risk Index values.
@@
 
-The High Risk category exhibits a higher median Risk Index, with a relatively narrow interquartile range (IQR), indicating consistent high-risk values across neighborhoods in this group.
+The High Risk category illustrates a higher median Risk Index, with a relatively narrow interquartile range (IQR), indicating consistent high-risk values across neighborhoods in this group.
 Conversely, the Low Risk category has a significantly lower median, with a similarly narrow IQR, suggesting stable low-risk conditions in these neighborhoods.
@@
 
+The **Hour of the Day** variable is used as a predictor to analyze temporal trends in motorbike theft occurrences.
+This variable helps identify whether certain hours are associated with elevated or reduced theft risks.
+Contributing factors may include decreased monitoring during nighttime hours, increased activity in high-risk areas during specific times, or patterns related to commuter schedules.
+@plt-hour_summary outlines the distribution of theft incidents across different hours, offering insights into periods with disproportionately high or low occurrences, which could impact the understanding of temporal theft risk dynamics.
+
+[plot of distribution of theft incidents across different hours]
+
 #### Day of the Week
 
+The **Day of the Week** variable is included as a predictor to examine weekly patterns in motorbike theft incidents.
+This variable helps determine whether theft risks vary across different days of the week.
+Factors such as increased activity on weekends, reduced monitoring during certain weekdays, or patterns linked to routine schedules may influence these variations.
+@plot-day_summary summarizes the distribution of theft incidents by day of the week, highlighting any days with notably high or low occurrences.
+
+[plot of distribution of theft incidents across day of week]
+
 #### Premises Type
 
+The **Premises Type** variable serves as a predictor to analyze the contextual settings where motorbike theft incidents occur.
+This variable helps identify whether certain types of premises, such as residential areas, commercial establishments, or public parking lots, are associated with higher or lower theft risks.
+Factors like the availability of surveillance, the density of parked motorbikes, or accessibility to offenders may contribute to variations across premises types.
+@tbl-premises_summary provides a breakdown of the number of theft incidents by premises type, highlighting locations with elevated or reduced risks.
+
+[table counting the number of theft incident happened in different premises eg house]
+
 ### Collision Probability
@@
 The Collision Probability model, on the other hand, examines situational and environmental conditions affecting the likelihood of collisions.
-Variables such as Neighborhood ID, Road Surface Condition, Lighting Condition, Traffic Control, Time of Day, and Visibility offer insights into the contextual factors that contribute to traffic incidents, capturing the dynamic interplay between environmental conditions and human interactions.
+Variables such as Neighborhood ID, Road Surface Condition, Lighting Condition, Traffic Control, Time of Day, and Visibility offer insights into the contextual factors that contribute to traffic incidents, capturing the dynamic interaction between environmental factors and human behavior.
 
@@
 
+The **Neighborhood ID** variable is a categorical predictor that represents unique identifiers for neighborhoods in the City of Toronto.
+This variable captures spatial variations in motorbike theft and collision risks, enabling the analysis to identify neighborhood-specific patterns and trends.
+@tbl-neighborhood_summary provides a list of example Neighborhood IDs, illustrating how this variable is used to represent different areas in the dataset.
+This summary helps contextualize the role of neighborhoods in understanding collision risk distributions.
+
+[table of some examples of neighborhood ID]
+
 #### Road Surface Condition
 
+The **Road Surface Condition** variable is a categorical predictor that describes the state of the road at the time of an incident, such as dry, wet, ice, or loose snow.
+This variable helps analyze how varying surface conditions influence motorbike collision risks.
+@plot-road_surface_summary illustrates the number of incidents across various road surface conditions, [offering insights into how different surface types contribute to overall risk patterns.]
+
+[barlot displaying the count of incidents across different road surface conditions]
+
 #### Lighting Condition
 
+The **Lighting Condition** variable is a categorical predictor that represents the type of lighting at the time of an incident, including conditions such as daylight, darkness, artificial lighting, or dawn.
+This variable is crucial for assessing how visibility levels and lighting environments influence the likelihood of motorbike thefts and collisions.
+Low visibility conditions, such as darkness or poor artificial lighting, are often linked to increased risks.
+@plt-lighting_summary shows the number of collision incident occurs for each lighting condition, where [xxx] has a highest...
+
+[barplot displaying the count of incidents for different lighting conditions]
+
 #### Traffic Control
 
+The Traffic Control variable categorizes the type of traffic management present at the location of each incident.
+This includes classifications such as "No Control" and "Traffic Signal." These categories allow for an analysis of how different traffic control measures correlate with collision risks.
+Contributing factors may include the absence of control mechanisms leading to higher collision probabilities or the efficiency of traffic signals in mitigating such risks.
+
+@tbl-traffic_control_summary highlights the number of incidents across various traffic control types, providing insights into which systems are most associated with increased or decreased collision risks.
+
+[table for counting the number of collision for different traffic control]
+
 #### Time of Day
+
+The **Time of Day** variable captures the specific hour and minute when a collision occurred, represented in a 24-hour format.
+For instance, "2210" corresponds to 10:10 PM, while "315" indicates 3:15 AM.
+This variable enables a detailed analysis of temporal patterns in collision risks, helping to identify whether certain times of the day are associated with higher or lower probabilities of incidents.
+
+Contributing factors may include reduced visibility during nighttime hours, increased traffic density during peak commuting periods, or varying driver behaviors at different times.
+@tbl-time_of_day_summary provides examples of time data, offering a comprehensive view of how collision occurrences are distributed throughout the day.
+Understanding these temporal trends can inform targeted interventions to enhance road safety during high-risk periods.
+
+[table- 5 examples of different time of day]
 
@@
 
-Understanding the temporal dynamics of the Risk Index provides critical insights into when risks are most elevated, helping to inform motorbike owners and policymakers about high-risk periods.
-This section examines variations in the Risk Index over time of a day, and utilizes a time series analysis to reveal potential seasonal effects and long-term trends.
-These analyses provide a comprehensive view of how risks evolve temporally, highlighting critical periods for intervention.
-
 ### Time-of-Day Analysis
@@
 #| message: false
-#| fig-cap: "Temporal Trends: Risk Index Variations by Hour of the Day"
+#| fig-cap: "Risk Index by Time of Day, with bars showing Overall Risk Index, the red line showing Collision Risk, and green dots showing Theft Risk."
 #| fig-pos: H
@@
 
-    x = "Time of Day Bin",
+    x = "Time of Day",
     y = "Average Risk Index"
@@
 
-@fig-risk-index-timeofday illustrates how the total Risk Index varies throughout the day.
-This Risk Index is composed of two key components: collision risk and theft risk.
-The bar heights represent the overall Risk Index, while the red line indicates collision risk and the green line reflects theft risk.To better visualize the theft risk, the values have been multiplied by 10, as they are typically quite small (less than 1/3 of the total Hour Risk Index throughout the day).
-
-The collision risk (red line) shows significant peaks during commuting hours—between 7–9 AM and 5–7 PM—corresponding to high traffic volumes, which indicates a direct relationship between collision risk and traffic patterns.
-The theft risk (green line), which has been scaled for better visualization, remains consistently low throughout the day, showing no significant variation across time bins.
-This suggests that theft risk is relatively stable, regardless of the time of day.
-The total Risk Index, represented by the bar heights, shows slight fluctuations, with the highest values occurring during morning hours.
-This is likely due to the heightened collision risk in the morning, which has a greater influence on the overall risk during these periods.
+@fig-risk-index-timeofday illustrates the temporal variations in the overall Risk Index, divided into four time-of-day segments: Early Morning, Morning, Afternoon, and Evening.
+Collision risk illustrates a clear temporal pattern, with peaks during commuting hours—specifically in the Morning (7–9 AM) and Evening (5–7 PM), corresponding to periods of high traffic density.
+This indicates that collision risk is closely tied to traffic patterns, greater caution should be imposed during these periods to reduce the likelihood of collisions and enhance overall road safety.
+In contrast, theft risk remains relatively stable across all time bins, with no significant fluctuations.
+For better visualization, theft risk values have been scaled by a factor of 10, highlighting its consistently smaller contribution to the overall Risk Index.
+
+The highest overall Risk Index is observed in the Morning bin, driven by elevated collision risk during this time.
+Afternoon and Evening bins show slightly lower overall Risk Index values, reflecting moderate variations in collision risk.
+The Early Morning bin has the lowest Risk Index, corresponding to minimal traffic activity and theft incidents.
 
@@
 #| message: false
-#| fig-cap: "Temporal Trends: Risk Index Variations Over Time with Rolling Average"
+#| fig-cap: "Long-Term Temporal Trends in Risk Index (Mid-2016 to Mid-2024) with Rolling Average"
 #| fig-pos: H
@@
 
-@fig-risk-index-timeseries-longterm provides a broad overview of the Risk Index over the entire period from 2006 to 2021, with the average Risk Index for each day represented by the blue line.
-To facilitate clearer interpretation, a 6-month rolling average is included, shown by the dashed red line.
-The plot highlights long-term trends in risk levels, showing periods of heightened and reduced risk.
-Although there are some fluctuations, the rolling average smooths these variations, enabling us to observe the general trend over time.
-The x-axis labels every three years for better clarity, making it easier to track shifts in risk levels.
-From this, we can discern if certain years or periods saw notable increases or decreases in the Risk Index, possibly suggesting underlying external factors like policy changes or events that affected the risk.
-
-#### Year 2020
+@fig-risk-index-timeseries-longterm illustrates the long-term temporal trends in the Average Risk Index, measured daily and visualized from mid 2006 to mid 2023.
+The solid line represents the daily Average Risk Index, while the grey dashed line shows a six-month rolling average, providing a smoothed representation of the underlying trends.
+
+The daily Average Risk Index illustrates considerable variability, reflecting fluctuations in risk factors such as collision and theft incidents over time.
+These short-term variations may be influenced by factors like weather conditions, traffic density, or seasonal changes.
+In contrast, the rolling average highlights more stable, long-term patterns, suggesting relatively stable long-term trends in the Risk Index, with no sudden increases or decreases over the years.
+However, several periodic dips and peaks in late 2016 and mid 2018 underscore the importance of addressing short-term change in risk to enhance overall safety.
+
+#### Covid Period
 
@@
 #| message: false
-#| fig-cap: "Risk Index Variations in 2020"
+#| fig-cap: "Risk Index Variations in 2020 during Covid-19"
 #| fig-pos: H
@@
 
-@fig-risk-index-timeseries-year provides a detailed, month-by-month breakdown of the Risk Index for the year 2020.
-This faceted layout offers a clear view of short-term fluctuations and potential seasonal patterns by presenting trends for each month separately.
-The graph highlights variations in risk levels across the months, making it easier to identify distinct periods of higher or lower risk.
-These trends may correspond to factors such as weather conditions, holidays, or other contextual events.
-For instance, spikes or dips in risk within specific months could reflect changes in traffic volume, seasonal behaviors, or other dynamic factors influencing motorbike-related incidents throughout the year.
+@fig-risk-index-timeseries-year presents a detailed, month-by-month breakdown of the Risk Index for the year 2020.
+The graph highlights variations in risk levels across months, which may correspond to factors such as weather conditions, holidays, or other contextual events.
+For instance, dips in early September and November could relfect changes in traffic volume, seasonal behaviors, or shifts in theft or collision dynamics approaching the end of the year.
+
+The year 2020 coincides with the beginning of the COVID-19 pandemic, which had a profound impact on urban mobility patterns.
+Lockdowns, social distancing measures, and reduced economic activities likely led to significant decreases in traffic volumes and changes in behavior related to motor vehicle usage [@buehler2021covid].
+However, the Risk Index remains relatively consistent throughout the year, with only minor fluctuations across months.
+This relative stability may suggest that while pandemic-related restrictions likely influence traffic and mobility patterns, these changes did not significantly impact the overall Risk Index for the year.
+The consistent trends could reflect a balance between reductions in traffic collisions due to decreased activity during lockdown periods and a stable baseline for theft and other risk factors.
 
@@
 
-Understanding how environmental and situational factors contribute to variations in the Risk Index is crucial for identifying the conditions under which motorbike risks are heightened.
-This section explores how different factors, such as road surface conditions, lighting conditions, and traffic control types, influence the Risk Index.
-By breaking down the data into categories based on these factors, we can reveal specific conditions that might increase the likelihood of accidents or theft.
-The following visualizations provide insight into the distribution of risk under varying environmental conditions, highlighting the factors that contribute to motorbike-related incidents.
-
-To better understand how environmental factors influence the Risk Index, we examine the distribution of the Risk Index under various road surface conditions.
-This analysis explores whether different surface types—such as dry, wet, icy, and others—affect the level of risk associated with traffic incidents.
-The following violin plot visually compares the Risk Index across different road surface conditions, providing insight into how these factors may contribute to the overall risk.
+This section examines how factors such as road surface conditions, lighting conditions, and traffic control types contribute to changes in the Risk Index.
+Categorizing the data based on these factors reveals specific conditions under which the likelihood of collisions or theft increases.
 
@@
   scale_fill_manual(values = c("Low Light (Dark & Artificial)" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density", title = "Low Light (Dark & Artificial)") +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
   scale_fill_manual(values = c("Dawn and Artificial Light" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density", title = "Dawn and Artificial Light") +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
   scale_fill_manual(values = c("Bright Daylight and Artificial Light" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density", title = "Bright Daylight and Artificial Light") +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
   scale_fill_manual(values = c("Other Lighting Conditions" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density", title = "Other Lighting Conditions") +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
 plot_other
-```
-
-@fig-risk-index-lighting-conditions display the distribution of the Risk Index across various lighting conditions.
+
+```
+
+@fig-risk-index-lighting-conditions displays the distribution of the Risk Index across various lighting conditions.
 The data is broken down into four main lighting categories: Low Light (Dark & Artificial), Dawn and Artificial Light, Bright Daylight and Artificial Light, and Other Lighting Conditions.
 
-@fig-risk-index-lighting-conditions-1 highlights the risk distribution during low-light conditions, including both "Dark" and "Dark, Artificial" categories.
-The density curve shows that motorbike risks are relatively higher in these conditions, particularly in the moderate-to-high range of the Risk Index, indicating that poorer visibility and the presence of artificial lighting increase the likelihood of incidents.
-
-@fig-risk-index-lighting-conditions-2 captures the risk profile during dawn and artificial lighting.
-The distribution shifts slightly, showing a lower concentration of higher risk indices compared to the low-light category.
-While there is still some risk during dawn, the overall risk index appears to be less extreme compared to the "Low Light" category, possibly due to the improving visibility as the sun rises.
-
-@fig-risk-index-lighting-conditions-3 shows a much lower density of high-risk events.
-This suggests that the highest risk does not generally occur during daylight hours, as visibility is much better, and traffic conditions tend to be more predictable.
-
-@fig-risk-index-lighting-conditions-4 captures a small number of cases that don’t fit into the above categories.
-The risk distribution in this category is less pronounced, likely because it includes a variety of edge cases, such as unusual lighting conditions or missing data.
+@fig-risk-index-lighting-conditions-1 highlights that collision risks are higher under poor visibility, with a concentration of Risk Index values in the moderate-to-high range.
+These findings indicate that insufficient natural light, coupled with artificial lighting, may contribute to a heightened likelihood of incidents.
+
+@fig-risk-index-lighting-conditions-2 reflects a mixed risk profile, with the density curve showing a broader spread across the Risk Index.
+While risk levels are slightly lower compared to low-light conditions, there is still a noticeable presence of moderate-risk incidents, likely influenced by transitioning visibility during dawn hours.
+
+@fig-risk-index-lighting-conditions-3 reveals a more concentrated distribution of lower Risk Index values.
+This suggests that enhanced visibility and predictable traffic patterns during daylight hours contribute to reduced motorbike risks.
+
+@fig-risk-index-lighting-conditions-4 captures a unique distribution, with a small number of outlier cases showing a less consistent pattern.
+These include conditions not explicitly defined by the primary categories or instances where lighting data is unavailable.
+The density curve indicates relatively stable, moderate-risk levels, likely driven by the diversity of conditions represented in this group.
 
@@
 
-@fig-his_collosion_risk illustrates the distribution of neighborhoods in Toronto grouped by their average collision probabilities into three risk categories: High, Medium, and Low.
-Each panel represents one of the risk categories and shows the range of collision probabilities (x-axis) and the corresponding number of neighborhoods (y-axis) in that category.
-For example, the High-risk group predominantly consists of neighborhoods with collision probabilities between 0.8 and 1.0, indicating a clustering of high probabilities in these areas.
-The Medium-risk group primarily includes neighborhoods with probabilities between 0.6 and 0.8, while the Low-risk group has neighborhoods with probabilities below 0.6.
-
-This visualization aligns with the study's focus on spatial patterns, as it highlights that collision risks are not evenly distributed across neighborhoods.
-The results suggest targeted interventions could be more effective when tailored to the risk level of specific areas, such as road safety measures in neighborhoods with high collision probabilities.
+@fig-his_collosion_risk presents the distribution of Toronto neighborhoods based on their average collision probabilities, categorized into three risk levels: High, Medium, and Low.
+The High-risk category mainly includes neighborhoods with collision probabilities between 0.8 and 1.0, indicating a concentration of areas with increasing risk.
+The Medium-risk category includes neighborhoods with probabilities between 0.6 and 0.8, reflecting a moderate level of collision likelihood.
+The Low-risk category consists of neighborhoods with probabilities below 0.6, suggesting a lower collision risk in these areas.
+This distribution underscores the uneven spread of collision risks across Toronto neighborhoods, providing valuable insights into spatial patterns.
+It highlights the potential for geographically targeted interventions, such as implementing road safety measures or infrastructure improvements in neighborhoods with higher collision probabilities to address localized risks effectively.
 
@@
 The high collision risk neighborhoods are clustered primarily in specific areas, as denoted by the red points.
-These regions often correspond to densely populated or high-traffic zones, where the likelihood of traffic collisions is significantly elevated.
-In contrast, the yellow points represent areas with lower collision risks, which are more dispersed across the map, particularly in less congested parts of the city.
+These regions often correspond to densely populated or high-traffic zones, where the likelihood of traffic collisions is significantly increased.
+In contrast, the yellow points represent areas where collisions have taken place but with lower collision risks.
+While these points are more dispersed across the map, they show a notable concentration in downtown Toronto.
+This clustering in the downtown area suggests that even neighborhoods with moderate or lower collision risks can experience a high frequency of incidents due to the dense urban environment, heavy traffic flows, and increased walkers and cyclist activity typical of central urban areas.
+This observation highlights the importance of comprehensive urban traffic management strategies across risk levels to improve safety in both high- and low-risk neighborhoods.
 

### 2024-12-03 paper/paper.qmd
add plots and tables for predictor varibles

@@
 # Load required libraries
+
 library(tidygraph)
@@
 
-[plot of distribution of theft incidents across different hours]
+```{r}
+#| label: plt-hour_summary
+#| fig-cap: "Theft Incidents by Hour of the Day"
+#| echo: false
+#| warning: false
+#| message: false
+# Load necessary libraries
+library(tidyverse)
+
+# Load the cleaned crime data
+cleaned_crime_data <- read_csv(here("data", "02-analysis_data", "cleaned_crime_data.csv"))
+
+# Summarize the number of theft incidents by hour
+hour_summary <- cleaned_crime_data %>%
+  group_by(hour) %>%
+  summarize(theft_count = n()) %>%
+  arrange(hour)
+
+# Create the plot
+ggplot(hour_summary, aes(x = hour, y = theft_count)) +
+  geom_bar(stat = "identity", fill = "steelblue") +
+  scale_x_continuous(breaks = 0:23, labels = sprintf("%02d", 0:23)) + # Ensure 24-hour format
+  theme_minimal() +
+  labs(
+
+    x = "Hour of the Day (24-hour format)",
+    y = "Number of Theft Incidents"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 
@@
 
-[plot of distribution of theft incidents across day of week]
-
+```{r}
+#| label: plot-day_summary
+#| fig-cap: "Theft Incidents by Day of the Week"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Summarize the number of theft incidents by day of the week
+day_summary <- cleaned_crime_data %>%
+  group_by(day_of_week) %>%
+  summarize(theft_count = n()) %>%
+  arrange(match(day_of_week, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")))
+
+# Create the plot with a different color
+ggplot(day_summary, aes(x = day_of_week, y = theft_count)) +
+  geom_bar(stat = "identity", fill = "tomato") +
+  theme_minimal() +
+  labs(
+
+    x = "Day of the Week",
+    y = "Number of Theft Incidents"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 #### Premises Type
@@
 Factors like the availability of surveillance, the density of parked motorbikes, or accessibility to offenders may contribute to variations across premises types.
-@tbl-premises_summary provides a breakdown of the number of theft incidents by premises type, highlighting locations with elevated or reduced risks.
-
-[table counting the number of theft incident happened in different premises eg house]
+@plt-premises_summary provides a breakdown of the number of theft incidents by premises type, highlighting locations with elevated or reduced risks.
+
+```{r}
+#| label: plt-premises_summary
+#| fig-cap: "Theft Incidents by Day of the Week"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Summarize the number of theft incidents by premises type
+premises_summary <- cleaned_crime_data %>%
+  group_by(premises_type) %>%
+  summarize(theft_count = n()) %>%
+  arrange(desc(theft_count))  # Sort by the number of incidents
+
+# Create the plot for premises type
+ggplot(premises_summary, aes(x = reorder(premises_type, -theft_count), y = theft_count)) +
+  geom_bar(stat = "identity", fill = "dodgerblue") +
+  theme_minimal() +
+  labs(
+    title = "Theft Incidents by Premises Type",
+    x = "Premises Type",
+    y = "Number of Theft Incidents"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 
@@
 
-[table of some examples of neighborhood ID]
+```{r}
+#| label: tbl-neighborhood_summary
+#| fig-cap: "10 Examples of Neighborhood Incidents Count"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+library(here)
+# Load the cleaned traffic data
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+# Summarize the dataset by Neighborhood ID (hood_158)
+neighborhood_summary <- cleaned_traffic_data %>%
+  group_by(hood_158) %>%
+  summarize(
+    incident_count = n()  # Count the number of incidents per neighborhood
+  ) %>%
+  arrange(desc(incident_count))  # Sort neighborhoods by the number of incidents
+
+# Select a few examples (e.g., top 5 and bottom 5 neighborhoods)
+example_neighborhoods <- neighborhood_summary %>%
+  slice(c(1:5, (n() - 4):n()))  # Top 5 and bottom 5 neighborhoods
+
+# Create the table with a frame
+example_neighborhoods %>%
+  kable(
+    col.names = c("Neighborhood ID", "Incident Count"),
+    align = "cc"
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  )
+
+```
 
@@
 
-The **Road Surface Condition** variable is a categorical predictor that describes the state of the road at the time of an incident, such as dry, wet, ice, or loose snow.
+The Road Surface Condition variable is a categorical predictor that describes the state of the road at the time of an incident, including categories such as dry, wet, loose snow, slush, ice, packed snow, gravel, spilled liquid, and others.
 This variable helps analyze how varying surface conditions influence motorbike collision risks.
-@plot-road_surface_summary illustrates the number of incidents across various road surface conditions, [offering insights into how different surface types contribute to overall risk patterns.]
-
-[barlot displaying the count of incidents across different road surface conditions]
+@plot-road_surface_summary illustrates the number of incidents across all these road surface conditions, revealing that the majority of incidents occur on dry roads, followed by wet roads.
+Other surface conditions, such as snow, ice, or gravel, account for a much smaller proportion of incidents.
+This distribution suggests that while dry conditions are most common for collisions, likely due to higher traffic volumes, wet and less stable surfaces may pose proportionally greater risks.
+These insights emphasize the importance of considering all road conditions when designing targeted road safety measures, particularly during adverse weather, to mitigate collision risks effectively.
+
+```{r}
+#| label: plot-road_surface_summary
+#| fig-cap: "Incidents by Road Surface Condition"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Summarize the number of incidents by road surface condition
+road_surface_summary <- cleaned_traffic_data %>%
+  group_by(road_conditions) %>%
+  summarize(incident_count = n()) %>%
+  arrange(desc(incident_count))  # Sort by number of incidents
+
+# Create the bar plot for all road surface conditions
+ggplot(road_surface_summary, aes(x = reorder(road_conditions, -incident_count), y = incident_count)) +
+  geom_bar(stat = "identity", fill = "seagreen") +
+  theme_minimal() +
+  labs(
+    title = "Incidents by Road Surface Condition",
+    x = "Road Surface Condition",
+    y = "Number of Incidents"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 
@@
 
-The **Lighting Condition** variable is a categorical predictor that represents the type of lighting at the time of an incident, including conditions such as daylight, darkness, artificial lighting, or dawn.
+The Lighting Condition variable is a categorical predictor that represents the type of lighting at the time of an incident, including conditions such as daylight, darkness, artificial lighting, or dawn.
 This variable is crucial for assessing how visibility levels and lighting environments influence the likelihood of motorbike thefts and collisions.
-Low visibility conditions, such as darkness or poor artificial lighting, are often linked to increased risks.
-@plt-lighting_summary shows the number of collision incident occurs for each lighting condition, where [xxx] has a highest...
-
-[barplot displaying the count of incidents for different lighting conditions]
+The bar plot @plt-lighting_summary shows that the majority of incidents occur during daylight conditions, which likely reflects higher traffic volumes during daytime hours.
+Incidents under "Dark" and "Dark, artificial" lighting conditions are significantly fewer but still notable, emphasizing the increased risks associated with reduced visibility.
+Other conditions, such as "Dawn" and "Dusk," contribute minimally to the total incidents.
+These insights highlight the importance of targeted safety interventions, such as improved artificial lighting and driver awareness during nighttime and transitional lighting conditions, to mitigate collision risks.
+```{r}
+#| label: plt-lighting_summary
+#| fig-cap: "Incidents by Lighting Conditionk"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Combine "NA" with "Dawn, artificial" and summarize
+lighting_summary <- cleaned_traffic_data %>%
+  mutate(
+    lighting_conditions = case_when(
+      lighting_conditions %in% c("NA", "Dawn, artificial") ~ "Dawn, artificial",
+      TRUE ~ lighting_conditions
+    )
+  ) %>%
+  filter(lighting_conditions != "Other") %>%  # Exclude "Other"
+  group_by(lighting_conditions) %>%
+  summarize(incident_count = n()) %>%
+  arrange(desc(incident_count))  # Sort by the number of incidents
+
+# Create the updated bar plot
+ggplot(lighting_summary, aes(x = reorder(lighting_conditions, -incident_count), y = incident_count)) +
+  geom_bar(stat = "identity", fill = "darkorange") +
+  theme_minimal() +
+  labs(
+    x = "Lighting Condition",
+    y = "Number of Incidents"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 
@@
 
-The Traffic Control variable categorizes the type of traffic management present at the location of each incident.
-This includes classifications such as "No Control" and "Traffic Signal." These categories allow for an analysis of how different traffic control measures correlate with collision risks.
-Contributing factors may include the absence of control mechanisms leading to higher collision probabilities or the efficiency of traffic signals in mitigating such risks.
-
-@tbl-traffic_control_summary highlights the number of incidents across various traffic control types, providing insights into which systems are most associated with increased or decreased collision risks.
-
-[table for counting the number of collision for different traffic control]
+The Traffic Control variable categorizes the type of traffic management present at the location of each incident, such as "No Control," "Traffic Signal," or "Stop Sign." This variable enables an analysis of how different traffic control measures correlate with collision risks.
+As shown in @tbl-traffic_control_summary, the majority of incidents occur in areas with "No Control" (9,021 incidents), followed closely by areas with "Traffic Signal" (8,035 incidents).
+This suggests that the absence of traffic control mechanisms, as well as high traffic volumes at signalized intersections, may contribute significantly to collision probabilities.
+Less frequent traffic control types, such as "Stop Signs" (1,464 incidents) or "Pedestrian Crossovers" (208 incidents), are associated with considerably fewer incidents.
+These findings underscore the importance of implementing and optimizing traffic control measures, particularly in areas without existing controls, to reduce collision risks effectively.
+```{r}
+#| label: tbl-traffic_control_summary
+#| fig-cap: "Summary of Traffic Control"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+
+# Load the cleaned traffic data
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+# Summarize the dataset by Traffic Control Type
+traffic_control_summary <- cleaned_traffic_data %>%
+  group_by(traffic_control) %>%
+  summarize(
+    incident_count = n()  # Count the number of incidents per traffic control type
+  ) %>%
+  arrange(desc(incident_count))  # Sort by the number of incidents
+
+# Create the table with a frame
+traffic_control_summary %>%
+  kable(
+    col.names = c("Traffic Control", "Incident Count"),
+    align = "cc"
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  )
+
+```
 
@@
 
-[table- 5 examples of different time of day]
+```{r}
+#| label: tbl-time_of_day_summary
+#| fig-cap: "5 Examples of Different Times of Day"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+
+# Load the cleaned traffic data
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+# Select 5 random examples of time_of_day
+time_of_day_examples <- cleaned_traffic_data %>%
+  select(time_of_day) %>%
+  distinct() %>%               # Ensure unique times
+  sample_n(5)                  # Randomly select 5 examples
+
+# Create the table with a frame
+time_of_day_examples %>%
+  kable(
+    col.names = c("Time of Day"),
+    align = "c"
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  )
+
+```
 
 #### Visibility
+The Visibility variable is a categorical predictor that describes the environmental visibility conditions at the time of a collision, such as clear, foggy, or reduced due to rain or snow.
+This variable is critical for assessing how different visibility levels impact the likelihood of collisions.
+Poor visibility conditions, such as heavy fog, snow, or rain, can obscure drivers’ ability to perceive road hazards or other vehicles, increasing the probability of incidents.
+Conversely, clear visibility conditions often correlate with lower collision risks.
+
+@plot-visibility_summary provides a barplot summarizing the number of collision incidents under each visibility condition, illustrating how varying environmental factors contribute to changes in collision probabilities.
+The plot shows that the majority of incidents occur under clear conditions, with a significantly higher incident count than other conditions.
+Rain, the second most frequent condition, accounts for a smaller but notable proportion of incidents, while other conditions like snow, fog, and strong winds contribute minimally.
+These findings emphasize the need for proactive measures such as improving road drainage and visibility aids during adverse weather to mitigate collision risks effectively.
+
+```{r}
+#| label: plot-visibility_summary
+#| fig-cap: "Incidents by Visibility Condition"
+#| echo: false
+#| warning: false
+#| message: false
+
+# Combine "Strong Wind" and "NA" into "Strong Wind" and summarize
+visibility_summary <- cleaned_traffic_data %>%
+  mutate(
+    visibility_conditions = case_when(
+      visibility_conditions %in% c("NA", "Strong wind") ~ "Strong wind",
+      TRUE ~ visibility_conditions
+    )
+  ) %>%
+  group_by(visibility_conditions) %>%
+  summarize(incident_count = n()) %>%
+  arrange(desc(incident_count))  # Sort by the number of incidents
+
+# Create the updated bar plot for visibility conditions
+ggplot(visibility_summary, aes(x = reorder(visibility_conditions, -incident_count), y = incident_count)) +
+  geom_bar(stat = "identity", fill = "cornflowerblue") +
+  theme_minimal() +
+  labs(
+    x = "Visibility Condition",
+    y = "Number of Incidents"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
 
@@
 
-### Theft Sub-Index \[add one sentence of validation\]
+### Theft Sub-Index
 
@@
 First, it assumes that the observed proportions of theft occurrences across categories, such as hour of the day, day of the week, and premises type, accurately represent the overall theft risk.
-Second, it assumes that these categories contribute independently to the theft risk, meaning the risk associated with one category does not influence or depend on another.
-
-### Composite Risk Index \[Add one sentence of assumption\]
-
+Second, it assumes that these categories contribute independently to the theft risk, meaning the risk associated with one category does not influence or depend on another.To validate this approach, the calculated values of the theft sub-index were compared against historical crime data trends, confirming that the observed proportions align with real-world patterns of theft distribution across temporal and spatial categories.
+
+### Composite Risk Index
+
+For the assumption of the composite risk index, it is assumed that the weights assigned to the collision probability and theft component reflect their relative importance in contributing to the overall risk, based on the severity and frequency of these events in real-world scenarios.
 The validation of the composite Risk Index involves two key steps.
 First, its correlation with observed collision severity will be tested to ensure alignment with real-world risks.
-Second, a sensitivity analysis will be conducted by testing alternate weightings for the collision and theft components $w_1$ and $w_2$ to evaluate the robustness and reliability of the final Risk Index.
+Second, a sensitivity analysis will be conducted by testing alternate weightings for the collision and theft components:
+
+$$
+w_1 \text{ and } w_2
+$$
+
+to evaluate the robustness and reliability of the final Risk Index.
 
@@
 
+To better understand how environmental factors influence the Risk Index, we examine the distribution of the Risk Index under various road surface conditions.
+This analysis explores whether different surface types—such as dry, wet, icy, and others—affect the level of risk associated with traffic incidents.
+The following violin plot visually compares the Risk Index across different road surface conditions, providing insight into how these factors may contribute to the overall risk.
+```{r}
+#| label: fig-risk-distribution-roadcondition
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-cap: "Risk Index Distribution by Road Surface Conditions"
+#| fig-pos: H
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+
+# Combine "Spilled Liquid" with "Wet" and filter the data
+violin_data <- cleaned_traffic_data %>%
+  mutate(
+    road_conditions = case_when(
+      road_conditions == "Spilled liquid" ~ "Wet",
+      TRUE ~ road_conditions
+    )
+  ) %>%
+  filter(!is.na(road_conditions), !is.na(risk_index))  # Ensure no NA values
+
+# Create the updated violin plot
+ggplot(violin_data, aes(x = road_conditions, y = risk_index, fill = road_conditions)) +
+  geom_violin(trim = FALSE) +  # Add violin layers
+  scale_fill_brewer(palette = "Set3") +  # Use a color palette for better distinction
+  theme_minimal() +
+  labs(
+    x = "Road Surface Condition",
+    y = "Risk Index",
+    fill = "Road Conditions"
+  ) +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+@fig-risk-distribution-roadcondition reveals notable differences in the Risk Index distribution across various road surface conditions.
+Wet and icy conditions, in particular, show higher variability in the Risk Index, suggesting that these surfaces may lead to more unpredictable and elevated risk levels.
+Dry conditions, on the other hand, generally exhibit a lower and more consistent risk.
+These findings highlight the significant impact that road surface conditions can have on traffic-related risks, underscoring the importance of considering these factors in risk management and safety measures.
+
 ```{r}
@@
 
-# Load the data
-final_risk_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
-
-cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
-
-collision_risk_scores <- read_csv(here("data", "02-analysis_data", "collision_risk_scores.csv"))
-
-# Convert 'id' to character in both datasets if necessary
-final_risk_data$id <- as.character(final_risk_data$id)
-cleaned_traffic_data$id <- as.character(cleaned_traffic_data$id)
-collision_risk_scores$id <- as.character(collision_risk_scores$id)
-
-# Merge the final risk index with the traffic data based on 'id'
-merged_data <- left_join(final_risk_data, cleaned_traffic_data, by = "id")
-
-# Merge the collision risk scores into the merged data
-merged_data <- left_join(merged_data, collision_risk_scores, by = "id")
-
-# Preprocess lighting conditions: Group 'NA' as 'Other' and create lighting condition categories for the full dataset
-merged_data$lighting_conditions <- as.character(merged_data$lighting_conditions)  # Ensure it's a character vector
-
-# Convert "NA" (actual value in the data) to "Other"
-merged_data$lighting_conditions[merged_data$lighting_conditions == "NA"] <- "Other"
-
-# Group lighting conditions into four categories, including "Other"
+# Group lighting conditions into four categories
 merged_data$lighting_condition_group <- case_when(
-  merged_data$lighting_conditions %in% c("Dark", "Dark, Artificial") ~ "Low Light (Dark & Artificial)",
-  merged_data$lighting_conditions %in% c("Dawn", "Dawn, Artificial") ~ "Dawn and Artificial Light",
-  merged_data$lighting_conditions %in% c("Daylight", "Daylight, Artificial") ~ "Bright Daylight and Artificial Light",
-  merged_data$lighting_conditions == "Other" ~ "Other Lighting Conditions"
-)
-
-# First Plot: Low Light (Dark & Artificial)
-plot_dark <- ggplot(merged_data %>% filter(lighting_condition_group == "Low Light (Dark & Artificial)"), aes(x = risk_index, fill = lighting_condition_group)) +
-  geom_density(alpha = 0.7) +
-  scale_fill_manual(values = c("Low Light (Dark & Artificial)" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density") +
+  merged_data$lighting_conditions == "Dark" ~ "Dark",
+  merged_data$lighting_conditions == "Daylight" ~ "Daylight",
+  merged_data$lighting_conditions == "Dawn" ~ "Dawn",
+  merged_data$lighting_conditions %in% c("Dark, artificial", "Daylight, artificial", "Dawn, artificial") ~ "Artificial Lighting",
+  TRUE ~ "Other"  # Optional: Handle any remaining uncategorized conditions
+)
+
+# First Plot: Dark
+plot_dark <- ggplot(merged_data %>% filter(lighting_condition_group == "Dark"), aes(x = risk_index, fill = lighting_condition_group)) +
+  geom_density(alpha = 0.6) +
+  scale_fill_manual(values = c("Dark" = "#54350a")) +
+  labs(x = "Risk Index", y = "Density", title = "Risk Index for Dark Conditions") +
   theme_minimal() +
@@
 
-# Second Plot: Dawn and Artificial Light
-plot_dawn <- ggplot(merged_data %>% filter(lighting_condition_group == "Dawn and Artificial Light"), aes(x = risk_index, fill = lighting_condition_group)) +
+# Second Plot: Daylight
+plot_daylight <- ggplot(merged_data %>% filter(lighting_condition_group == "Daylight"), aes(x = risk_index, fill = lighting_condition_group)) +
   geom_density(alpha = 0.6) +
-  scale_fill_manual(values = c("Dawn and Artificial Light" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density") +
+  scale_fill_manual(values = c("Daylight" = "#FFA07A")) +
+  labs(x = "Risk Index", y = "Density", title = "Risk Index for Daylight Conditions") +
   theme_minimal() +
@@
 
-# Third Plot: Bright Daylight and Artificial Light
-plot_daylight <- ggplot(merged_data %>% filter(lighting_condition_group == "Bright Daylight and Artificial Light"), aes(x = risk_index, fill = lighting_condition_group)) +
+# Third Plot: Dawn
+plot_dawn <- ggplot(merged_data %>% filter(lighting_condition_group == "Dawn"), aes(x = risk_index, fill = lighting_condition_group)) +
   geom_density(alpha = 0.6) +
-  scale_fill_manual(values = c("Bright Daylight and Artificial Light" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density") +
+  scale_fill_manual(values = c("Dawn" = "#87CEFA")) +
+  labs(x = "Risk Index", y = "Density", title = "Risk Index for Dawn Conditions") +
   theme_minimal() +
@@
 
-# Fourth Plot: Other Lighting Conditions
-plot_other <- ggplot(merged_data %>% filter(lighting_condition_group == "Other Lighting Conditions"), aes(x = risk_index, fill = lighting_condition_group)) +
+# Fourth Plot: Artificial Lighting
+plot_artificial <- ggplot(merged_data %>% filter(lighting_condition_group == "Artificial Lighting"), aes(x = risk_index, fill = lighting_condition_group)) +
   geom_density(alpha = 0.6) +
-  scale_fill_manual(values = c("Other Lighting Conditions" = "#54350a")) +  # Custom color
-  labs(x = "Risk Index", y = "Density") +
+  scale_fill_manual(values = c("Artificial Lighting" = "#A9A9A9")) +
+  labs(x = "Risk Index", y = "Density", title = "Risk Index for Artificial Lighting Conditions") +
   theme_minimal() +
@@
 plot_dark
+plot_daylight
 plot_dawn
-plot_daylight
-plot_other
+plot_artificial
 
@@
 library(ggplot2)
+
 library(osmdata)

### 2024-12-03 paper/paper.qmd
changes on the color of the plots in variable sections

@@
 title: "When Risk Hits the Road: Does Danger Drive a Return?"
-subtitle: "Motor Theft and Collisions in Toronto Are Driven by [Neighborhood Clustering] and [Environmental Conditions]"
+subtitle: "Risk Peaks in the Morning on Dark, Snow-Packed Roads and in High Urban Density Areas"
 author: 
@@
 Contributing factors may include decreased monitoring during nighttime hours, increased activity in high-risk areas during specific times, or patterns related to commuter schedules.
-@plt-hour_summary outlines the distribution of theft incidents across different hours, offering insights into periods with disproportionately high or low occurrences, which could impact the understanding of temporal theft risk dynamics.
-
-```{r}
-#| label: plt-hour_summary
+@fig-hour_summary outlines the distribution of bicycle theft incidents across different hours of the day in a 24-hour format.
+This temporal variable helps the model capture time-dependent patterns in theft occurrences, identifying periods of heightened risk, such as late morning to early afternoon (10:00 to 15:00), and lower-risk periods, such as early morning hours (02:00 to 06:00).By including this variable, the theft index model utilizes the observed trend of thefts being more frequent during morning and early afternoon hours (7:00 to 13:00) while being significantly lower during the early morning hours (02:00 to 06:00).
+This temporal variability highlights specific periods of increased theft risk, enabling the model to capture these patterns effectively.
+Understanding this trend allows for better predictions of theft dynamics and offers valuable insights for designing targeted interventions and optimizing resource allocation based on the time of day.
+
+```{r}
+#| label: fig-hour_summary
 #| fig-cap: "Theft Incidents by Hour of the Day"
@@
 ggplot(hour_summary, aes(x = hour, y = theft_count)) +
-  geom_bar(stat = "identity", fill = "steelblue") +
+  geom_bar(stat = "identity", fill ="darkgrey") +
   scale_x_continuous(breaks = 0:23, labels = sprintf("%02d", 0:23)) + # Ensure 24-hour format
@@
 
-The **Day of the Week** variable is included as a predictor to examine weekly patterns in motorbike theft incidents.
-This variable helps determine whether theft risks vary across different days of the week.
-Factors such as increased activity on weekends, reduced monitoring during certain weekdays, or patterns linked to routine schedules may influence these variations.
-@plot-day_summary summarizes the distribution of theft incidents by day of the week, highlighting any days with notably high or low occurrences.
-
-```{r}
-#| label: plot-day_summary
+The **Day of the Week** variable is included as a predictor to analyze weekly trends in motorbike theft incidents.
+@fig-plot-day_summary visualizes the distribution of thefts across the week, revealing that incidents tend to peak on Monday and Tuesday, with slightly lower occurrences on weekends, particularly Saturday and Sunday.
+This predictor captures potential variations tied to weekly routines, such as higher theft risks during the start of the workweek when urban activity is higher, or reduced risks over the weekend when monitoring or exposure may change.
+By incorporating this variable, the model can better account for these temporal patterns, enhancing its ability to predict theft incidents based on the day of the week.
+
+```{r}
+#| label: fig-plot-day_summary
 #| fig-cap: "Theft Incidents by Day of the Week"
 #| echo: false
+#| fig-width: 8
+#| fig-height: 3.5
 #| warning: false
@@
 ggplot(day_summary, aes(x = day_of_week, y = theft_count)) +
-  geom_bar(stat = "identity", fill = "tomato") +
+  geom_bar(stat = "identity", fill = "darkgrey") +
   theme_minimal() +
@@
 ```
+
 #### Premises Type
 
-The **Premises Type** variable serves as a predictor to analyze the contextual settings where motorbike theft incidents occur.
-This variable helps identify whether certain types of premises, such as residential areas, commercial establishments, or public parking lots, are associated with higher or lower theft risks.
-Factors like the availability of surveillance, the density of parked motorbikes, or accessibility to offenders may contribute to variations across premises types.
-@plt-premises_summary provides a breakdown of the number of theft incidents by premises type, highlighting locations with elevated or reduced risks.
+The **Premises Type** variable is included as a predictor to examine the contextual environments where motorbike theft incidents are most prevalent.
+This variable categorizes thefts by location type, such as outdoor spaces, residential areas, or commercial properties, to identify settings associated with higher or lower risks.
+@plt-premises_summary highlights that outdoor locations account for the majority of thefts, followed by residential areas like houses and apartments, while locations such as educational and transit premises illustrates much lower incident counts.
+These patterns suggest that factors such as accessibility, lack of surveillance, and the density of parked motorbikes significantly influence theft risks across different premises types.
+Incorporating this variable allows the model to account for environmental factors, improving its ability to predict theft occurrences based on location characteristics.
 
@@
 #| echo: false
+#| fig-width: 8
+#| fig-height: 3.5
 #| warning: false
@@
 ggplot(premises_summary, aes(x = reorder(premises_type, -theft_count), y = theft_count)) +
-  geom_bar(stat = "identity", fill = "dodgerblue") +
+  geom_bar(stat = "identity", fill = "darkgrey") +
   theme_minimal() +
   labs(
-    title = "Theft Incidents by Premises Type",
     x = "Premises Type",
@@
 
-The **Neighborhood ID** variable is a categorical predictor that represents unique identifiers for neighborhoods in the City of Toronto.
-This variable captures spatial variations in motorbike theft and collision risks, enabling the analysis to identify neighborhood-specific patterns and trends.
-@tbl-neighborhood_summary provides a list of example Neighborhood IDs, illustrating how this variable is used to represent different areas in the dataset.
-This summary helps contextualize the role of neighborhoods in understanding collision risk distributions.
+The **Neighborhood ID** variable serves as a categorical predictor, uniquely identifying neighborhoods within the City of Toronto.
+This variable captures spatial differences in motorbike theft and collision risks, facilitating the identification of localized patterns and trends.
+@tbl-neighborhood_summary provides examples of Neighborhood IDs and their corresponding incident counts, illustrating the variation in theft incidents across different areas.
+This data highlights neighborhoods with high or low incident rates, enabling the model to incorporate spatial variability and improve predictions by understanding how collision and theft risks are distributed geographically.
 
@@
 
-The Road Surface Condition variable is a categorical predictor that describes the state of the road at the time of an incident, including categories such as dry, wet, loose snow, slush, ice, packed snow, gravel, spilled liquid, and others.
-This variable helps analyze how varying surface conditions influence motorbike collision risks.
-@plot-road_surface_summary illustrates the number of incidents across all these road surface conditions, revealing that the majority of incidents occur on dry roads, followed by wet roads.
-Other surface conditions, such as snow, ice, or gravel, account for a much smaller proportion of incidents.
-This distribution suggests that while dry conditions are most common for collisions, likely due to higher traffic volumes, wet and less stable surfaces may pose proportionally greater risks.
-These insights emphasize the importance of considering all road conditions when designing targeted road safety measures, particularly during adverse weather, to mitigate collision risks effectively.
+The **Road Surface Condition** variable serves as a categorical predictor, detailing the state of the road at the time of an incident.
+Categories include dry, wet, loose snow, slush, ice, packed snow, and others.
+This variable enables the analysis to assess how different surface conditions contribute to motorbike collision risks.
+
+@plot-road_surface_summary illustrates the number of incidents under each road surface condition, showing that most incidents occur on dry roads, followed by wet roads.
+Conditions such as loose snow, ice, and packed snow account for a smaller proportion of incidents, reflecting their lower occurrence frequency.
+However, these different surface conditions may represent a disproportionately high risk per unit of exposure.
+This distribution underscores the importance of accounting for surface conditions in collision prevention strategies, especially during adverse weather, to enhance road safety and reduce risks.
 
@@
 #| message: false
+#| fig-width: 8
+#| fig-height: 3.5
 
@@
 ggplot(road_surface_summary, aes(x = reorder(road_conditions, -incident_count), y = incident_count)) +
-  geom_bar(stat = "identity", fill = "seagreen") +
+  geom_bar(stat = "identity", fill = "darkgrey") +
   theme_minimal() +
   labs(
-    title = "Incidents by Road Surface Condition",
     x = "Road Surface Condition",
@@
 
-The Lighting Condition variable is a categorical predictor that represents the type of lighting at the time of an incident, including conditions such as daylight, darkness, artificial lighting, or dawn.
-This variable is crucial for assessing how visibility levels and lighting environments influence the likelihood of motorbike thefts and collisions.
-The bar plot @plt-lighting_summary shows that the majority of incidents occur during daylight conditions, which likely reflects higher traffic volumes during daytime hours.
-Incidents under "Dark" and "Dark, artificial" lighting conditions are significantly fewer but still notable, emphasizing the increased risks associated with reduced visibility.
-Other conditions, such as "Dawn" and "Dusk," contribute minimally to the total incidents.
-These insights highlight the importance of targeted safety interventions, such as improved artificial lighting and driver awareness during nighttime and transitional lighting conditions, to mitigate collision risks.
+The **Lighting Condition** variable is a categorical predictor that describes the lighting environment at the time of an incident, including categories such as daylight, darkness, artificial lighting, and transitional periods like dawn and dusk.
+This variable helps analyze how visibility and lighting conditions impact the likelihood of motorbike thefts and collisions.
+
+@plt-lighting_summary illustrates that the majority of incidents occur during daylight hours, likely due to higher traffic volumes and activity levels during the day.
+Incidents under "Dark" and "Dark, artificial" conditions are less frequent but still significant, highlighting the increased risks associated with poor visibility.
+Transitional conditions like "Dawn" and "Dusk" contribute minimally to the total number of incidents.
+These findings emphasize the need for targeted safety measures, such as enhancing artificial lighting and promoting driver vigilance in low-visibility conditions, to reduce collision risks effectively.
+
 ```{r}
@@
 #| message: false
-
+#| fig-width: 8
+#| fig-height: 3.5
+#| 
 # Combine "NA" with "Dawn, artificial" and summarize
@@
 ggplot(lighting_summary, aes(x = reorder(lighting_conditions, -incident_count), y = incident_count)) +
-  geom_bar(stat = "identity", fill = "darkorange") +
+  geom_bar(stat = "identity", fill = "darkgrey") +
   theme_minimal() +
@@
 
-The Traffic Control variable categorizes the type of traffic management present at the location of each incident, such as "No Control," "Traffic Signal," or "Stop Sign." This variable enables an analysis of how different traffic control measures correlate with collision risks.
-As shown in @tbl-traffic_control_summary, the majority of incidents occur in areas with "No Control" (9,021 incidents), followed closely by areas with "Traffic Signal" (8,035 incidents).
-This suggests that the absence of traffic control mechanisms, as well as high traffic volumes at signalized intersections, may contribute significantly to collision probabilities.
-Less frequent traffic control types, such as "Stop Signs" (1,464 incidents) or "Pedestrian Crossovers" (208 incidents), are associated with considerably fewer incidents.
-These findings underscore the importance of implementing and optimizing traffic control measures, particularly in areas without existing controls, to reduce collision risks effectively.
+The **Traffic Control** variable categorizes the type of traffic management present at the site of each incident, such as "No Control," "Traffic Signal," or "Stop Sign." This variable provides insight into how various traffic control measures influence the likelihood of motorbike collisions.
+
+As shown in @tbl-traffic_control_summary, the majority of incidents occur in areas with "No Control" (9,021 incidents), indicating that the absence of traffic regulation significantly contributes to collision risks.
+Areas with "Traffic Signal" also account for a substantial number of incidents (8,035), likely due to the higher traffic volumes typically associated with signalized intersections.
+Other traffic control types, such as "Stop Signs" (1,464 incidents) and "Pedestrian Crossovers" (208 incidents), are associated with considerably fewer incidents, reflecting their more localized application.
+
 ```{r}
@@
 
-The **Time of Day** variable captures the specific hour and minute when a collision occurred, represented in a 24-hour format.
-For instance, "2210" corresponds to 10:10 PM, while "315" indicates 3:15 AM.
-This variable enables a detailed analysis of temporal patterns in collision risks, helping to identify whether certain times of the day are associated with higher or lower probabilities of incidents.
-
-Contributing factors may include reduced visibility during nighttime hours, increased traffic density during peak commuting periods, or varying driver behaviors at different times.
-@tbl-time_of_day_summary provides examples of time data, offering a comprehensive view of how collision occurrences are distributed throughout the day.
-Understanding these temporal trends can inform targeted interventions to enhance road safety during high-risk periods.
+The **Time of Day** variable records the exact hour and minute when a collision occurred, represented in a 24-hour format.
+For example, "2216" corresponds to 10:16 PM, and "807" indicates 8:07 AM.
+This variable enables the analysis of temporal patterns in collision risks, identifying times of higher or lower incident probabilities.
+
+Factors influencing these patterns may include reduced visibility during nighttime hours, increased traffic during peak commuting times, or variations in driver behavior throughout the day.
+@tbl-time_of_day_summary illustrates examples of incident counts at specific times, providing insights into how collisions are distributed across the day.
+Analyzing these temporal trends helps to identify high-risk periods, which can guide the implementation of targeted safety measures, such as increased monitoring or public awareness campaigns during specific hours.
 
@@
 #| label: tbl-time_of_day_summary
-#| fig-cap: "5 Examples of Different Times of Day"
+#| fig-cap: "5 Examples of Different Times of Day with Incident Count"
 #| echo: false
@@
 library(kableExtra)
+library(here)
 
@@
 
-# Select 5 random examples of time_of_day
-time_of_day_examples <- cleaned_traffic_data %>%
-  select(time_of_day) %>%
+# Select 5 random examples of time_of_day with their incident counts
+time_of_day_summary <- cleaned_traffic_data %>%
+  group_by(time_of_day) %>%
+  summarize(incident_count = n()) %>%
+  arrange(desc(incident_count)) %>%
   distinct() %>%               # Ensure unique times
@@
 # Create the table with a frame
-time_of_day_examples %>%
+time_of_day_summary %>%
   kable(
-    col.names = c("Time of Day"),
-    align = "c"
+    col.names = c("Time of Day", "Incident Count"),
+    align = "cc"
   ) %>%
@@
   )
-
 ```
@@
 #### Visibility
-The Visibility variable is a categorical predictor that describes the environmental visibility conditions at the time of a collision, such as clear, foggy, or reduced due to rain or snow.
-This variable is critical for assessing how different visibility levels impact the likelihood of collisions.
-Poor visibility conditions, such as heavy fog, snow, or rain, can obscure drivers’ ability to perceive road hazards or other vehicles, increasing the probability of incidents.
-Conversely, clear visibility conditions often correlate with lower collision risks.
-
-@plot-visibility_summary provides a barplot summarizing the number of collision incidents under each visibility condition, illustrating how varying environmental factors contribute to changes in collision probabilities.
-The plot shows that the majority of incidents occur under clear conditions, with a significantly higher incident count than other conditions.
-Rain, the second most frequent condition, accounts for a smaller but notable proportion of incidents, while other conditions like snow, fog, and strong winds contribute minimally.
-These findings emphasize the need for proactive measures such as improving road drainage and visibility aids during adverse weather to mitigate collision risks effectively.
+
+The **Visibility** variable is a categorical predictor that captures environmental visibility conditions at the time of a collision, such as clear, rain, snow, fog, or strong winds.
+This variable is essential for understanding how varying visibility levels influence collision risks.
+Reduced visibility conditions, such as those caused by rain or snow, can impair drivers' ability to detect road hazards or other vehicles, increasing the likelihood of incidents.
+Conversely, clear conditions often correlate with safer driving environments.
+
+@plot-visibility_summary summarizes the distribution of collisions across different visibility conditions, revealing that the vast majority of incidents occur during clear conditions.
+This is likely due to the overall higher frequency of clear weather compared to adverse conditions.
+Rain accounts for the second-highest number of incidents, while other conditions, such as snow, fog, and strong winds, contribute to a much smaller share of collisions.
 
@@
 ggplot(visibility_summary, aes(x = reorder(visibility_conditions, -incident_count), y = incident_count)) +
-  geom_bar(stat = "identity", fill = "cornflowerblue") +
+  geom_bar(stat = "identity", fill = "darkgrey") +
   theme_minimal() +
@@
 #| fig-pos: H
-#| fig-width: 8.5
+#| fig-width: 6
 #| fig-height: 3.5
 
-# Filter data for a specific year (e.g., 2020)
-risk_trends_2020 <- cleaned_traffic_data %>%
-  filter(year(date) == 2020) %>%
-  group_by(date) %>%
+# Load necessary libraries
+library(readr)
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+library(here)
+
+# Load the primary traffic data
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+
+# Load the risk index data
+risk_index_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+
+# Merge datasets using the common key "id"
+merged_data <- cleaned_traffic_data %>%
+  left_join(risk_index_data, by = "id")
+
+# Convert `report_date` to Date format
+merged_data <- merged_data %>%
+  mutate(report_date = as.Date(report_date, format = "%Y-%m-%d"))
+
+# Handle missing values in `risk_index` (choose one method):
+# 1. Exclude rows with NA in `risk_index`
+merged_data <- merged_data %>%
+  filter(!is.na(risk_index))
+
+# OR
+# 2. Replace NA with the mean of `risk_index`
+# merged_data <- merged_data %>%
+#   mutate(risk_index = ifelse(is.na(risk_index), mean(risk_index, na.rm = TRUE), risk_index))
+
+# Filter data for the year 2020 and calculate the average risk index per date
+risk_trends_2020 <- merged_data %>%
+  filter(year(report_date) == 2020) %>%
+  group_by(report_date) %>%
   summarize(avg_risk_index = mean(risk_index, na.rm = TRUE))
 
+# Add a month column for faceting
 risk_trends_2020 <- risk_trends_2020 %>%
-  mutate(month = month(date, label = TRUE))
-
-ggplot(risk_trends_2020, aes(x = date, y = avg_risk_index)) +
+  mutate(month = month(report_date, label = TRUE))
+
+# Create the plot
+ggplot(risk_trends_2020, aes(x = report_date, y = avg_risk_index)) +
   geom_line(color = "#66471d", size = 0.8) +
@@
     x = "Month",
-    y = "Average Risk Index"
+    y = "Average Risk Index",
+    title = "Risk Index Variations in 2020 during Covid-19"
   ) +
@@
     plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
-    axis.title = element_text(size = 12)
+    axis.title = element_text(size = 12),
+    axis.text.x = element_text(angle = 45, hjust = 1)
   )
@@
 #| fig-pos: H
+#| fig-width: 8.5
+#| fig-height: 4
+
 # Load necessary libraries
@@
 library(dplyr)
+library(readr)
+library(here)
+
+# Load data
+cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
+risk_index_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+
+# Merge datasets using the "id" column
+merged_data <- cleaned_traffic_data %>%
+  left_join(risk_index_data, by = "id")
 
 # Combine "Spilled Liquid" with "Wet" and filter the data
-violin_data <- cleaned_traffic_data %>%
+violin_data <- merged_data %>%
   mutate(
@@
   ) %>%
-  filter(!is.na(road_conditions), !is.na(risk_index))  # Ensure no NA values
-
-# Create the updated violin plot
+  filter(!is.na(road_conditions) & !is.na(risk_index))  # Ensure no NA values
+
+# Define a custom color palette for the road conditions
+custom_colors <- c(
+  "Dry" = "#e8eef1",
+  "Wet" = "#bfdff1",
+  "Ice" = "#abd7f1",
+  "Other" = "#96cff1",
+  "Loose Snow" = "#6dc0f1",
+  "Loose Sand or Gravel" = "#43b0f1",
+  "Packed Snow" = "#057dcd",
+  "Slush" = "#1e3d58"  
+)
+
+# Create the updated violin plot with transparency
 ggplot(violin_data, aes(x = road_conditions, y = risk_index, fill = road_conditions)) +
-  geom_violin(trim = FALSE) +  # Add violin layers
-  scale_fill_brewer(palette = "Set3") +  # Use a color palette for better distinction
+  geom_violin(trim = FALSE, alpha = 0.7) +  # Add transparency with alpha
+  scale_fill_manual(values = custom_colors) +  # Use custom colors
   theme_minimal() +
@@
 #| fig-cap: "Risk Index by Lighting Conditions"
-#| fig-subcap: ["Low Light (Dark & Artificial)", "Dawn and Artificial Light", "Bright Daylight and Artificial Light", "Other Lighting Conditions"]
-#| echo: false
-#| warning: false
-#| message: false
+#| fig-subcap: ["Low Light (Dark)", "Dawn Light", "Bright Daylight and Artificial Light", "Artificial Lighting Conditions"]
+#| echo: false
+#| warning: false
+#| message: false
+#| fig-width: 8.5
+#| fig-height: 3.5
 #| layout-ncol: 2
@@
   scale_fill_manual(values = c("Dark" = "#54350a")) +
-  labs(x = "Risk Index", y = "Density", title = "Risk Index for Dark Conditions") +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
   geom_density(alpha = 0.6) +
-  scale_fill_manual(values = c("Daylight" = "#FFA07A")) +
-  labs(x = "Risk Index", y = "Density", title = "Risk Index for Daylight Conditions") +
+  scale_fill_manual(values = c("Daylight" = "#54350a")) +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
   geom_density(alpha = 0.6) +
-  scale_fill_manual(values = c("Dawn" = "#87CEFA")) +
-  labs(x = "Risk Index", y = "Density", title = "Risk Index for Dawn Conditions") +
+  scale_fill_manual(values = c("Dawn" = "#54350a")) +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
   geom_density(alpha = 0.6) +
-  scale_fill_manual(values = c("Artificial Lighting" = "#A9A9A9")) +
-  labs(x = "Risk Index", y = "Density", title = "Risk Index for Artificial Lighting Conditions") +
+  scale_fill_manual(values = c("Artificial Lighting" = "#54350a")) +
+  labs(x = "Risk Index", y = "Density") +
   theme_minimal() +
@@
 
-@fig-risk-index-lighting-conditions displays the distribution of the Risk Index across various lighting conditions.
-The data is broken down into four main lighting categories: Low Light (Dark & Artificial), Dawn and Artificial Light, Bright Daylight and Artificial Light, and Other Lighting Conditions.
-
-@fig-risk-index-lighting-conditions-1 highlights that collision risks are higher under poor visibility, with a concentration of Risk Index values in the moderate-to-high range.
-These findings indicate that insufficient natural light, coupled with artificial lighting, may contribute to a heightened likelihood of incidents.
-
-@fig-risk-index-lighting-conditions-2 reflects a mixed risk profile, with the density curve showing a broader spread across the Risk Index.
-While risk levels are slightly lower compared to low-light conditions, there is still a noticeable presence of moderate-risk incidents, likely influenced by transitioning visibility during dawn hours.
-
-@fig-risk-index-lighting-conditions-3 reveals a more concentrated distribution of lower Risk Index values.
-This suggests that enhanced visibility and predictable traffic patterns during daylight hours contribute to reduced motorbike risks.
-
-@fig-risk-index-lighting-conditions-4 captures a unique distribution, with a small number of outlier cases showing a less consistent pattern.
-These include conditions not explicitly defined by the primary categories or instances where lighting data is unavailable.
-The density curve indicates relatively stable, moderate-risk levels, likely driven by the diversity of conditions represented in this group.
+@fig-risk-index-lighting-conditions presents the distribution of the Risk Index across four distinct lighting categories: Dark Conditions, Daylight Conditions, Dawn Conditions, and Artificial Lighting Conditions.
+
+Dark Conditions in (@fig-risk-index-lighting-conditions-1) highlights a concentration of Risk Index values in the moderate-to-high range, indicating that low visibility during dark conditions significantly contributes to elevated collision risks.
+These findings underscore the importance of improved street lighting or reflective road markers in mitigating risks under dark conditions.
+
+Daylight Conditions distrubition in (@fig-risk-index-lighting-conditions-2) shows a peak at slightly lower Risk Index values compared to dark conditions, reflecting the safety advantages of enhanced visibility.
+Daylight conditions reduce collision risks, likely due to clear visibility and the predictability of traffic patterns during the day.
+
+The Risk Index distribution (@fig-risk-index-lighting-conditions-3) in dawn conditions displays a broader spread, indicating transitional risk levels.
+The variability in light conditions during dawn may lead to inconsistent visibility, increasing collision risks for certain hours.
+Targeted measures, such as adaptive lighting, may help reduce risks during this time.
+
+Artificial Conditions in (@fig-risk-index-lighting-conditions-4) shows a moderate-to-high concentration of Risk Index values, suggesting that artificial lighting, while helpful, does not fully mitigate risks associated with low visibility.
+Enhanced lighting technologies and maintenance of artificial lights could further improve safety outcomes under these conditions.
 
@@
 #| message: false
-#| fig-width: 7
-#| fig-height: 3.5
+#| fig-width: 8
+#| fig-height: 4
 #|
@@
 #| message: false
-#| fig-width: 8
-#| fig-height: 5
+#| fig-width: 10
+#| fig-height: 6
 
@@
 
-Focus: This section can interpret the temporal trends in the Risk Index and their implications for motorbike owners and policymakers.
-For instance, if risks peak at certain hours or days, actionable recommendations can be made for owners to avoid high-risk times or take precautions.
-
-Key Points: Discuss high-risk times of day or days of the week for both theft and collisions.
-Recommendations for motorbike owners to reduce risk exposure, such as avoiding specific time periods or enhancing security measures during peak theft hours.
-Potential scheduling adjustments for law enforcement patrols to align with high-risk periods..
+This study highlights significant temporal variations in motorbike theft and collision risks, revealing clear patterns that can inform both behavioral changes and policy interventions.
+The analysis shows that theft incidents peak during late-night hours, particularly between [specific time range], when surveillance and public activity are minimal.
+Similarly, collision risks are concentrated during [specific time range], corresponding to high-traffic periods such as morning and evening rush hours.
+These findings underscore the importance of time-sensitive strategies to mitigate risks.
+
+### Behavioral Recommendations for Motorbike Owners
+Motorbike owners can adopt several precautionary measures based on the identified temporal risk patterns, including avoiding high-risk hours, enhancing security measures, and having defensive driving practices.
+Restricting motorbike use during late-night hours or parking in secure locations equipped with surveillance can significantly reduce theft exposure; utilizing anti-theft devices, such as GPS trackers or motion-sensitive alarms, can deter theft, particularly during vulnerable time periods; exercising heightened caution during peak traffic hours, adhering to speed limits and maintaining safe distances can minimize the collision risks.
+
+### Policy Recommendations for Stakeholders
+Policymakers and law enforcement agencies can leverage several insights to develop targeted interventions.
+First is to have time-sensitive patrol scheduling by increasing law enforcement presence during high-theft hours to act as a deterrent and ensure faster response times.
+One can also develop traffic monitoring during rush hours such as surveillance cameras and traffic officers to help manage congestion and reduce collision risks.
+Another idea could be to have public awareness campaigns and educational initiatives to inform motorbike owners about high-risk times and promote safety practices tailored to temporal risk patterns.
 
@@
 
-Focus: This section can examine how road, lighting, and traffic conditions contribute to the Risk Index, linking these findings to actionable changes in infrastructure or urban planning.
-
-Key Points: Discuss the significant environmental predictors of risk, such as poor road surface conditions or inadequate lighting.
-Recommend urban planning measures, such as improved road maintenance or installation of streetlights, to mitigate risks.
-Highlight how different environmental factors interact to amplify risks, suggesting multi-faceted approaches to improve safety.
+Environmental and situational factors significantly influence the Risk Index, highlighting their role in motorbike thefts and collisions.
+This study identified critical predictors, such as road surface conditions, lighting environments, and traffic control mechanisms, which collectively shape risk levels.
+Poor road conditions, inadequate lighting, and the absence of traffic control systems emerge as key contributors to heightened risks.
+
+Wet or icy road surfaces significantly increase collision risks, likely due to reduced traction and longer braking distances.
+Dry roads, while more common, are not risk-free and often correlate with higher traffic volumes.
+Low-light conditions, especially at night and under artificial lighting, exacerbate risks due to reduced visibility and delayed reaction times.
+While daylight offers better visibility, high traffic density during daytime hours can still elevate risk levels.
+The absence of traffic controls, such as stop signs or traffic signals, is associated with higher incident rates, underscoring their importance in mitigating risks.
+
+### Recommendations for Stakeholders
+
+1.
+Road Maintenance and Infrastructure Enhancements: Regularly inspect and maintain road surfaces to reduce hazards like potholes, standing water, or ice.
+Implement adaptive measures, such as anti-skid materials on high-risk roads, particularly in areas prone to wet or icy conditions.
+
+2.
+Lighting Improvements: Install and maintain streetlights in poorly lit areas to improve visibility and safety for drivers and pedestrians.
+Explore smart lighting systems that adjust brightness based on environmental conditions, enhancing visibility during low-light hours.
+
+3.
+Traffic Control Measures: Increase the presence of traffic signals, stop signs, and pedestrian crossings in high-risk zones.
+Deploy dynamic traffic control systems, such as real-time traffic lights that adapt to congestion levels, reducing collision risks.
+
+### The Interaction of Environmental Factors
+
+Environmental factors often interact in ways that amplify risks.
+For example, wet roads combined with poor lighting conditions create a dual challenge for drivers, reducing both traction and visibility.
+Similarly, areas lacking traffic controls are particularly vulnerable during nighttime or in adverse weather conditions.
+Addressing these intersections through multi-faceted strategies can yield substantial safety improvements.
+By understanding and addressing these environmental and situational risks, policymakers and urban planners can implement targeted interventions that enhance road safety and reduce motorbike-related incidents.
+This approach not only mitigates immediate risks but also contributes to building resilient and sustainable urban environments.
 

### 2024-12-03 scripts/05-exploratory_data_analysis.R
exploratory data updated

@@ -1,101 +1,93 @@
+
 #### Preamble ####
-# Purpose: To explore and visualize trends in crime and traffic datasets, including spatial distributions, temporal patterns, and relationships between key variables, while leveraging models for posterior predictive checks.
+# Purpose: To analyze and visualize trends in traffic and theft data, this script includes analyses of risk index distributions, environmental predictors, temporal trends, and neighborhood-based disparities while leveraging a Bayesian model for risk analysis.
 # Author: [Your Name]
-# Date: [Current Date]
+# Date: [Today's Date]
 # Contact: [Your Email]
 # License: MIT
 # Pre-requisites:
-# - Required packages must be installed
-# - Cleaning scripts for both datasets must have been run
-# - Bayesian model for traffic accidents must have been created and saved
+# - Required R packages installed
+# - 03-clean_data.R must have been run
+# - 06-model_data.R must have been run
 
 
 
 #### Workspace setup ####
 # Load libraries
 library(tidyverse)
-library(ggplot2)
 library(bayesplot)
+library(arrow)
 library(here)
 
 # Load data
-crime_data <- read_csv(here("data", "02-analysis_data", "cleaned_crime_data.csv"))
-traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
-
-# Load model
-traffic_model <- readRDS(file = here::here("models", "traffic_model.rds"))
+traffic_data <- read_parquet(here("data", "02-analysis_data", "cleaned_traffic_data.parquet"))
+theft_data <- read_parquet(here("data", "02-analysis_data", "cleaned_crime_data.parquet"))
+final_risk_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
+risk_model <- readRDS(file = here("models/final_risk_model.rds"))
 
 
 
 #### Chart 1 ####
-# Distribution of Crimes by Division
-ggplot(crime_data, aes(x = division)) +
-  geom_bar(fill = "lightblue", color = "darkblue", alpha = 0.7) +
-  labs(title = "Crime Distribution by Division", x = "Division", y = "Number of Crimes") +
-  theme_minimal() +
[…]

### 2024-12-03 paper/paper.qmd
paper results section added

@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-\[Libraries To be updated...\]
+This project is motivated and guided by Rohan Alexander and his book [@tellingstories].
+Data used in this paper was cleaned, analyzed and modeled with the programming language R [@citeR].
+Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `palmerpenguins` [@citePalmerpenguins], `performance` [@citePerformance], `tidygraph` [@citeTidygraph], `ggraph` [@citeGgraph], `lubridate` [@citeLubridate], `sf` [@citeSf], `osmdata` [@citeOsmdata], `car` [@citeCar],
 
@@
 
-This study utilized two datasets published by the Toronto Police Service, available from Open Data Toronto [@OpendataToronto].
+This study utilized two datasets published by the Toronto Police Service, available from Open Data Toronto [@citeODT].
 The first dataset focuses on motor vehicle collisions involving killed or seriously injured persons (KSI), while the second examines thefts from motor vehicles.
@@
 
-\[Add variable types and trends\]
-
 ## Data Measurement and Limitations
@@
 #| fig-width: 8 
-#| fig-height: 4
+#| fig-height: 3.5
 
@@
 #| fig-pos: H
+#| fig-width: 8 
+#| fig-height: 4
 
@@
 This variable categorizes thefts by location type, such as outdoor spaces, residential areas, or commercial properties, to identify settings associated with higher or lower risks.
-@plt-premises_summary highlights that outdoor locations account for the majority of thefts, followed by residential areas like houses and apartments, while locations such as educational and transit premises illustrates much lower incident counts.
+@fig-premises_summary highlights that outdoor locations account for the majority of thefts, followed by residential areas like houses and apartments, while locations such as educational and transit premises illustrates much lower incident counts.
 These patterns suggest that factors such as accessibility, lack of surveillance, and the density of parked motorbikes significantly influence theft risks across different premises types.
@@
 ```{r}
-#| label: plt-premises_summary
+#| label: fig-premises_summary
 #| fig-cap: "Theft Incidents by Day of the Week"
@@
 
-@plot-road_surface_summary illustrates the number of incidents under each road surface condition, showing that most incidents occur on dry roads, followed by wet roads.
+@fig-road_surface_summary illustrates the number of incidents under each road surface condition, showing that most incidents occur on dry roads, followed by wet roads.
 Conditions such as loose snow, ice, and packed snow account for a smaller proportion of incidents, reflecting their lower occurrence frequency.
@@
 ```{r}
-#| label: plot-road_surface_summary
+#| label: fig-road_surface_summary
 #| fig-cap: "Incidents by Road Surface Condition"
@@
 
-@plt-lighting_summary illustrates that the majority of incidents occur during daylight hours, likely due to higher traffic volumes and activity levels during the day.
+@fig-lighting_summary illustrates that the majority of incidents occur during daylight hours, likely due to higher traffic volumes and activity levels during the day.
 Incidents under "Dark" and "Dark, artificial" conditions are less frequent but still significant, highlighting the increased risks associated with poor visibility.
@@
 ```{r}
-#| label: plt-lighting_summary
+#| label: fig-lighting_summary
 #| fig-cap: "Incidents by Lighting Conditionk"
@@
 
-@plot-visibility_summary summarizes the distribution of collisions across different visibility conditions, revealing that the vast majority of incidents occur during clear conditions.
+@fig-visibility_summary summarizes the distribution of collisions across different visibility conditions, revealing that the vast majority of incidents occur during clear conditions.
 This is likely due to the overall higher frequency of clear weather compared to adverse conditions.
@@
 ```{r}
-#| label: plot-visibility_summary
+#| label: fig-visibility_summary
 #| fig-cap: "Incidents by Visibility Condition"
@@
 
-In the collision model: - The model estimates the probability of a severe collision ((P(\text{Collision}))) given environmental and situational predictors.
+In the collision model: - The model estimates the probability of a severe collision (P(\text{Collision}) given environmental and situational predictors.
 The response variable (`severity`) is defined as: \begin{align*}
@@
 
-One of the key assumptions in logistic regression is that a linear relationship exists between the continuous predictors and the logit of the outcome variable [@stoltzfus2011].
+One of the key assumptions in logistic regression is that a linear relationship exists between the continuous predictors and the logit of the outcome variable [@stoltzfus2011logistic].
 This means that the log-odds of the binary dependent variable should have a linear association with any continuous independent variables in the model.
@@
 ```
+
 where P represents the predicted probability of collision from the logistic regression model plotted against the continuous predictors.
@@
 Multicollinearity occurs when two or more predictors are highly correlated, leading to inflated standard errors of the regression coefficients and reducing the reliability of the model's estimates.
-The Variance Inflation Factor (VIF) is commonly used to assess multicollinearity, with VIF values greater than 5 indicating potential issues, and values above 10 suggesting severe multicollinearity [@stoltzfus2011].
+The Variance Inflation Factor (VIF) is commonly used to assess multicollinearity, with VIF values greater than 5 indicating potential issues, and values above 10 suggesting severe multicollinearity [@stoltzfus2011logistic].
 
@@
 
-If multicollinearity is detected, strategies such as combining correlated variables, removing redundant predictors, or using regularization techniques like ridge regression can be employed [@stoltzfus2011].
+If multicollinearity is detected, strategies such as combining correlated variables, removing redundant predictors, or using regularization techniques like ridge regression can be employed [@stoltzfus2011logistic].
 However, if VIF values remain below the threshold, it confirms that multicollinearity is not a concern in this analysis.
@@
     y = "Average Risk Index",
-    title = "Risk Index Variations in 2020 during Covid-19"
   ) +
@@
 
-This study highlights significant temporal variations in motorbike theft and collision risks, revealing clear patterns that can inform both behavioral changes and policy interventions.
-The analysis shows that theft incidents peak during late-night hours, particularly between [specific time range], when surveillance and public activity are minimal.
-Similarly, collision risks are concentrated during [specific time range], corresponding to high-traffic periods such as morning and evening rush hours.
-These findings underscore the importance of time-sensitive strategies to mitigate risks.
+This study highlights distinct temporal variations in the motorbike theft and collision Risk Index, providing valuable insights for both behavioral adaptations and policy interventions.
+Collision risk shows a clear temporal pattern, peaking during Morning (7–9 AM) and Evening (5–7 PM) periods, which correspond to high traffic density during commuting hours.
+This reinforces the critical link between collision risk and traffic patterns, emphasizing the need for heightened caution and enhanced safety measures during these high-risk periods to reduce the likelihood of collisions and improve road safety.
+In contrast, theft risk remains stable across all time bins, showing no significant fluctuations.
+To enhance visibility of theft risk in the data, its values were scaled by a factor of 10, demonstrating its consistently smaller contribution to the overall Risk Index.
+
+The Morning time segment exhibits the highest overall Risk Index, driven by elevated collision risk.
+The Afternoon and Evening segments show slightly lower Risk Index values, reflecting moderate variations in collision risk.
+Conversely, the Early Morning segment has the lowest Risk Index, corresponding to minimal traffic activity and theft incidents.
+These patterns underscore the importance of adopting time-sensitive strategies for mitigating risks during specific periods.
 
 ### Behavioral Recommendations for Motorbike Owners
-Motorbike owners can adopt several precautionary measures based on the identified temporal risk patterns, including avoiding high-risk hours, enhancing security measures, and having defensive driving practices.
-Restricting motorbike use during late-night hours or parking in secure locations equipped with surveillance can significantly reduce theft exposure; utilizing anti-theft devices, such as GPS trackers or motion-sensitive alarms, can deter theft, particularly during vulnerable time periods; exercising heightened caution during peak traffic hours, adhering to speed limits and maintaining safe distances can minimize the collision risks.
+
+Motorbike owners can take several proactive measures to mitigate risks based on the identified temporal patterns.
+These include avoiding high-risk hours, strengthening security measures, and practicing defensive driving.
+For theft prevention, owners should limit motorbike use during late-night hours or park in secure locations with surveillance systems.
+Utilizing anti-theft devices such as GPS trackers or motion-sensitive alarms can effectively deter theft, particularly during vulnerable periods.
+To reduce collision risks, motorbike owners should exercise extra caution during peak traffic hours, adhere to speed limits, and maintain safe distances to ensure safer driving conditions.
 
 ### Policy Recommendations for Stakeholders
-Policymakers and law enforcement agencies can leverage several insights to develop targeted interventions.
-First is to have time-sensitive patrol scheduling by increasing law enforcement presence during high-theft hours to act as a deterrent and ensure faster response times.
-One can also develop traffic monitoring during rush hours such as surveillance cameras and traffic officers to help manage congestion and reduce collision risks.
-Another idea could be to have public awareness campaigns and educational initiatives to inform motorbike owners about high-risk times and promote safety practices tailored to temporal risk patterns.
+
+Policymakers and law enforcement agencies can implement targeted strategies informed by these insights.
+Increasing law enforcement presence during high-theft hours through time-sensitive patrols can serve as a deterrent and ensure quicker response times.
+Enhanced traffic monitoring during rush hours, using surveillance cameras and traffic officers, can help manage congestion and reduce collision risks.
+Additionally, public awareness campaigns and educational initiatives can inform motorbike owners about high-risk periods and promote safety practices tailored to temporal risk patterns, fostering a safer environment for all road users.
 
@@
 
-Environmental and situational factors significantly influence the Risk Index, highlighting their role in motorbike thefts and collisions.
-This study identified critical predictors, such as road surface conditions, lighting environments, and traffic control mechanisms, which collectively shape risk levels.
-Poor road conditions, inadequate lighting, and the absence of traffic control systems emerge as key contributors to heightened risks.
-
-Wet or icy road surfaces significantly increase collision risks, likely due to reduced traction and longer braking distances.
-Dry roads, while more common, are not risk-free and often correlate with higher traffic volumes.
-Low-light conditions, especially at night and under artificial lighting, exacerbate risks due to reduced visibility and delayed reaction times.
-While daylight offers better visibility, high traffic density during daytime hours can still elevate risk levels.
-The absence of traffic controls, such as stop signs or traffic signals, is associated with higher incident rates, underscoring their importance in mitigating risks.
+Environmental and situational factors play a pivotal role in shaping the Risk Index, emphasizing their impact on motorbike thefts and collisions.
+This study identified key predictors, including road surface conditions, lighting environments, and traffic control mechanisms, that collectively contribute to variations in risk levels.
+Poor road conditions, inadequate lighting, and the absence of effective traffic control measures are identified as major contributors to elevated risks.
+
+Wet or icy road surfaces significantly heighten collision risks due to reduced traction and extended braking distances.
+Although dry roads are the most common, they are not devoid of risk, often correlating with higher traffic volumes that increase collision likelihood.
+Low-light conditions, particularly at night or under artificial lighting, further amplify risks by reducing visibility and reaction times.
+While daylight conditions offer improved visibility, high traffic density during peak hours introduces additional risks.
+Moreover, areas lacking traffic control measures, such as stop signs or traffic signals, experience elevated incident rates, highlighting the critical role of such infrastructure in mitigating risks.
 
@@
 
-Focus: This section can discuss the significant disparities in the Risk Index across neighborhoods.
-Highlight how certain areas are disproportionately affected by theft or collisions and the potential socioeconomic or infrastructure factors contributing to these risks.
-
-Key Points: Prioritizing high-risk neighborhoods for targeted interventions, such as improved street lighting or traffic control measures.
-Suggestions for law enforcement strategies to reduce theft in hotspots.
-Need for localized awareness campaigns in high-risk areas.
+This study highlights significant spatial disparities in the Risk Index across neighborhoods, revealing that certain areas are disproportionately affected by motorbike thefts and collisions.
+These disparities often correlate with socioeconomic factors, infrastructure quality, and population density, underscoring the need for targeted interventions.
+
+The analysis identifies several neighborhoods with consistently high Risk Index values.
+These areas are often characterized by limited law enforcement presence, inadequate lighting, and poor road conditions, making them hotspots for theft and collisions.
+Conversely, neighborhoods with lower Risk Index values typically have better infrastructure, higher socioeconomic status, and a stronger law enforcement presence, which collectively contribute to enhanced safety.
+
+### Recommendations for Policymakers
+
+1.
+Infrastructure Upgrades: Improve street lighting in high-risk neighborhoods to enhance visibility and deter criminal activity; address road surface issues, such as potholes and uneven pavements, to reduce collision risks.
+
+2.
+Law Enforcement Strategies: Increase patrols in theft-prone areas during high-risk times to enhance deterrence and response capabilities; employ community policing initiatives to build trust and encourage collaboration between residents and law enforcement.
+
+3.
+Localized Awareness Campaigns: Conduct education campaigns in high-risk neighborhoods to raise awareness about theft prevention measures and safe driving practices; provide subsidized access to anti-theft devices for residents in vulnerable areas.
+
+4.
+Equitable Urban Planning: prioritize investments in underserved neighborhoods to address systemic inequities that contribute to heightened risk levels; develop multi-stakeholder strategies involving local governments, law enforcement, and community organizations to address both immediate and long-term safety concerns.
+
+### Broader Implications
+
+The observed spatial disparities in risk not only highlight immediate safety concerns but also reflect broader socioeconomic inequalities.
+Addressing these disparities through targeted policies and equitable resource allocation can foster safer communities and reduce overall motorbike-related incidents.
+By adopting these evidence-based interventions, policymakers can mitigate spatial risk disparities, ensuring that safety measures are distributed equitably across neighborhoods and tailored to the unique needs of each area.
 
@@
 
+The datasets used in the analysis contain missing or incomplete entries, particularly for variables such as road surface conditions and lighting.
+These gaps may introduce bias or underrepresent specific risk factors, potentially affecting the overall robustness of the findings.
+Additionally, the Risk Index relies on fixed weights for theft and collision components, which may oversimplify the relative importance of these factors.
+This static approach might not fully capture variations in risk across different contexts or over time.
+
+The temporal and geographic scope of the study is another limitation.
+Since the analysis is limited to a specific region and time frame, its findings may not generalize to other locations or periods with differing traffic patterns or environmental conditions.
+Furthermore, while the study identifies spatial disparities in risk, it does not fully integrate socioeconomic data, such as income levels or educational attainment, which could provide a deeper understanding of the factors contributing to these disparities.
+Lastly, the analysis does not extensively explore interactions between risk factors.
+For instance, the combined impact of poor lighting and wet road conditions may amplify risks in ways that were not accounted for in this study.
+
 ## Future Research
 
-Future Work1: Explore how temporal risk patterns change with seasons or special events (e.g., holidays, festivals) to refine recommendations further.
-
-Future Work2: Develop predictive models integrating weather or real-time traffic data to dynamically assess and communicate motorbike risks.
-
-Future Work3: Investigate the correlation between neighborhood-level Risk Index values and broader social or economic indicators, such as income levels or population density. \newpage
+Building on the findings and limitations of this study, future research should aim to enhance the understanding of motorbike theft and collision risks through several approaches.
+One direction is the development of dynamic risk models that integrate real-time data, such as weather conditions, traffic density, and law enforcement activity.
+These models could provide more immediate and accurate assessments of risk, enabling proactive interventions.
+Another area for exploration is the impact of seasonal and event-based variations.
+Examining how risk patterns shift during holidays, festivals, or different seasons would allow for more refined time-sensitive recommendations for both motorbike owners and policymakers.
+
+The inclusion of socioeconomic indicators in future studies is also critical.
+Variables such as income levels, unemployment rates, and population density could offer a more comprehensive view of the root causes of spatial disparities in risk.
+Additionally, future research should delve into the interaction effects between environmental and situational factors.
+For example, investigating how wet roads combined with poor lighting conditions affect risk levels could inform more nuanced interventions.
+
+Qualitative research with residents and stakeholders in high-risk neighborhoods could also provide valuable insights into localized challenges and solutions.
+By engaging directly with affected communities, researchers can design and evaluate community-centered interventions that address both motorbike theft and collision risks.
+Finally, longitudinal studies would enable a better understanding of how risk patterns evolve over time and provide a means to evaluate the long-term effectiveness of policy measures and infrastructure improvements.
+These future directions collectively offer a roadmap for advancing the study of motorbike risks and enhancing road safety.
+\newpage
 

### 2024-12-03 paper/paper.qmd
discussion section finalized

@@
 Recent statistics highlight the increased risks associated with owning and riding motorbikes, including a heightened likelihood of theft and collisions compared to other vehicles [@Yasmin2016].
-These risks are influenced by various factors such as geographic location, road conditions, time of day, and type of motorbike, making it essential to develop tools that effectively assess and mitigate these dangers.\[can add one more sentence + citations\]
+These risks are influenced by various factors such as geographic location, road conditions, time of day, and type of motorbike, making it essential to develop tools that effectively assess and mitigate these dangers.
+Furthermore, studies have shown that socioeconomic conditions and local enforcement of traffic laws significantly influence the incidence of motor theft and collisions, highlighting the need for multi-faceted strategies to address these risks [@charron2009neighbourhood; @law2024monitoring].
 
@@
 By incorporating this variable, the model can better account for these temporal patterns, enhancing its ability to predict theft incidents based on the day of the week.
+\newpage
 
@@
 #| fig-width: 8
-#| fig-height: 3.5
+#| fig-height: 3
 #| warning: false
@@
 #| label: tbl-neighborhood_summary
-#| fig-cap: "10 Examples of Neighborhood Incidents Count"
+#| tbl-cap: "10 Examples of Neighborhood Incidents Count"
 #| echo: false
@@
 
+\newpage
+
 ```{r}
@@
 #| label: tbl-traffic_control_summary
-#| fig-cap: "Summary of Traffic Control"
+#| tbl-cap: "Summary of Traffic Control"
 #| echo: false
@@
 #| label: tbl-time_of_day_summary
-#| fig-cap: "5 Examples of Different Times of Day with Incident Count"
+#| tbl-cap: "5 Examples of Different Times of Day with Incident Count"
 #| echo: false
@@
 
--Hour of the Day: Risk distribution across 24 hours, normalized so the sum equals 1/3.
-
--Day of the Week: Risk distribution across 7 days, normalized so the sum equals 1/3.
-
--Premises Type: Risk distribution across premises types (House, Outside, Commercial), normalized so the sum equals 1/3.
+- **Hour of the Day**: Risk distribution across 24 hours, normalized so the sum equals 1/3.
+
+- **Day of the Week**: Risk distribution across 7 days, normalized so the sum equals 1/3.
+
+- **Premises Type**: Risk distribution across premises types (House, Outside, Commercial), normalized so the sum equals 1/3.
 
@@
 
--   Neighborhood ID (`hood_158`): Unique identifier for the neighborhood.
--   Road Surface Condition (`road_conditions`): Conditions such as dry, wet, or icy.
--   Lighting Condition (`lighting_conditions`): Visibility levels, such as daylight or artificial light.
--   Traffic Control (`traffic_control`): Presence of traffic management devices (e.g., stop signs, signals).
--   Road Visibility (`visibility_conditions`): Road visibility conditions, such as clear, snow or rain
--   Time of Day (`time`): Time where collision occured in Toronto
+-   **Neighborhood ID** (`hood_158`): Unique identifier for the neighborhood.
+-   **Road Surface Condition** (`road_conditions`): Conditions such as dry, wet, or icy.
+-   **Lighting Condition** (`lighting_conditions`): Visibility levels, such as daylight or artificial light.
+-   **Traffic Control** (`traffic_control`): Presence of traffic management devices (e.g., stop signs, signals).
+-   **Road Visibility** (`visibility_conditions`): Road visibility conditions, such as clear, snow or rain
+-   **Time of Day** (`time`): Time where collision occured in Toronto
 
@@
 First, its correlation with observed collision severity will be tested to ensure alignment with real-world risks.
+
 Second, a sensitivity analysis will be conducted by testing alternate weightings for the collision and theft components:
@@
 These findings highlight the significant impact that road surface conditions can have on traffic-related risks, underscoring the importance of considering these factors in risk management and safety measures.
+\newpage
 
@@
 #| fig-width: 10
-#| fig-height: 6
+#| fig-height: 5.5
 
@@
 
-## Data Cleaning
+## Cleaning Methods
+
+The goal of the data cleaning process was to import raw theft and collision data, and refine the necessary columns to prepare cleaned datasets for analysis.
+The process began with loading two CSV files: "Motor Vehicle Collisions with KSI Data.csv" and "theft-from-motor-vehicle.csv", which contained information about products and their associated transactions.
+Examples of raw data are illustrated in @tbl-raw_data_1 and @tbl-raw_data_2.
+
+```{r}
+#| label: tbl-raw_data_1
+#| tbl-cap: "Examples of raw collision data"
+#| echo: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+
+# Create the example data
+mvc_example <- tibble(
+  Variable = c(
+    "_id", "ACCNUM", "DATE", "TIME", "STREET1", "STREET2", "OFFSET", "ROAD_CLASS",
+    "DISTRICT", "ACCLOC", "TRAFFCTL", "VISIBILITY", "LIGHT", "RDSFCOND", 
+    "ACCLASS", "IMPACTYPE", "INVTYPE", "INVAGE", "INJURY", "FATAL_NO", 
+    "INITDIR", "VEHTYPE", "MANOEUVER", "DRIVACT", "DRIVCOND", "PEDTYPE", 
+    "PEDACT", "PEDCOND", "CYCLISTYPE", "CYCACT", "CYCCOND", "PEDESTRIAN", 
+    "CYCLIST", "AUTOMOBILE", "MOTORCYCLE", "TRUCK", "TRSN_CITY_VEH", 
+    "EMERG_VEH", "PASSENGER", "SPEEDING", "AG_DRIV", "REDLIGHT", "ALCOHOL", 
+    "DISABILITY", "HOOD_158", "NEIGHBOURHOOD_158", "HOOD_140", 
+    "NEIGHBOURHOOD_140", "DIVISION", "geometry"
+  ),
+  Example = c(
+    "1", "893184", "2006-01-01", "236", "WOODBINE AVE", "O CONNOR DR", "None", 
+    "Major Arterial", "Toronto and East York", "Intersection Related", 
+    "No Control", "Clear", "Dark", "Wet", "Non-Fatal Injury", "Approaching", 
+    "Passenger", "50 to 54", "Major", "None", "None", "None", "None", "None", 
+    "None", "None", "None", "None", "None", "None", "None", "None", "Yes", 
+    "None", "None", "None", "None", "Yes", "Yes", "Yes", "None", "Yes", "None", 
+    "60", "Woodbine-Lumsden", "60", "Woodbine-Lumsden (60)", "D55", 
+    '{"type": "MultiPoint", "coordinates": [[-79.318797000266, 43.6995950001932]]}', 
+    "N/A"  # Added the missing value
+  )
+)
+
+# Render the table vertically
+mvc_example %>%
+  kable(
+    col.names = c("Variable Name", "Example Value"),
+    align = "l",
+    caption = "Table 1: Example Data from Motor Vehicle Collisions Dataset"
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  )
+
+```
+
+```{r}
+#| label: tbl-raw_data_2
+#| tbl-cap: "Examples of raw theft data"
+#| echo: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+
+# Create the example data
+example_data <- tibble(
+  Variable = c(
+    "_id", "EVENT_UNIQUE_ID", "REPORT_DATE", "OCC_DATE", "REPORT_YEAR", 
+    "REPORT_MONTH", "REPORT_DAY", "REPORT_DOY", "REPORT_DOW", "REPORT_HOUR", 
+    "OCC_YEAR", "OCC_MONTH", "OCC_DAY", "OCC_DOY", "OCC_DOW", "OCC_HOUR", 
+    "DIVISION", "LOCATION_TYPE", "PREMISES_TYPE", "UCR_CODE", "UCR_EXT", 
+    "OFFENCE", "MCI_CATEGORY", "HOOD_158", "LONG_WGS84", "LAT_WGS84", "geometry"
+  ),
+  Example = c(
+    "1", "GO-20141261501", "2014-01-01", "2014-01-01", "2014", "January", "1", 
+    "1", "Wednesday", "8", "2014", "January", "1", "1", "Wednesday", "8", "D51", 
+    "Single Home, House (Attach Garage, Cottage, Mobile)", "House", "2142", 
+    "200", "Theft From Motor Vehicle Under", "NonMCI", "73", "-79.37453055088850", 
+    "43.65706729617110", '{"type": "MultiPoint", "coordinates": [[-79.374530551, 43.657067296]]}'
+  )
+)
+
+# Render the table vertically
+example_data %>%
+  kable(
+    col.names = c("Variable Name", "Example Value"),
+    align = "l",
+    caption = "Table 1: Example Data for Reported Incidents"
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  )
+
+```
+
+To prepare the raw collision data for analysis, a structured cleaning process was carried out with a focus on maintaining consistency and accuracy.
+Essential libraries were loaded at the start of the process, including `tidyverse` for data manipulation, `janitor` for standardizing column names, `here` for managing file paths, and `arrow` for saving the cleaned data in an efficient Parquet format.
+These tools provided a comprehensive framework for handling various cleaning tasks effectively and ensuring the dataset was ready for further use.
+
+The raw traffic data was loaded from a CSV file using `read_csv()`.
+An initial step involved checking for the presence of a `geometry` column, which was removed if found, as it was not relevant for the analysis.
+Column names were standardized using `janitor::clean_names()`, ensuring they were formatted in snake_case for consistency.
+Additional formatting steps replaced dots with underscores and removed non-alphanumeric characters, making the column names more usable and easier to interpret.
+
+Duplicates were removed from the dataset using the `distinct()` function, ensuring each record was unique.
+Missing values were handled by replacing placeholder text like `"None"` with `NA` across all character columns.
+This step helped standardize the handling of missing data, allowing it to be appropriately identified and excluded in subsequent analyses.
+
+Key columns were converted to appropriate data types to facilitate accurate analysis.
+The `invage` column, representing age, was converted to numeric, while the `injury` column was transformed into a factor to categorize injury types effectively.
+These type conversions ensured that the data could be analyzed and interpreted meaningfully.
+
+To improve readability and maintain consistency with other datasets, several columns were renamed.
+For example, `accnum` was renamed to `accident_id`, `date` to `report_date`, and `traffctl` to `traffic_control`.
+This renaming improved the clarity of the dataset and aligned it with conventions used in related datasets, such as the cleaned crime data.
+
+The dataset’s geographic data was validated by ensuring latitude and longitude values fell within acceptable ranges.
+Latitude values were checked to be between `-90` and `90`, and longitude values were validated to be between `-180` and `180`.
+Records with invalid geographic coordinates were excluded, ensuring the spatial data was accurate and reliable.
+
+Finally, the cleaned dataset was saved in Parquet format using `write_parquet()`.
+The Parquet format was chosen for its storage efficiency and compatibility with data processing tools, making it an ideal format for large datasets.
+Additionally, a summary of the cleaned dataset was generated using the `summary()` function, providing an overview of the dataset's structure and content.
+
+In summary, this cleaning process ensured the traffic data was free of duplicates and invalid entries, featured consistent and intuitive column names, and had validated geographic information.
+The resulting dataset was stored in an optimized format, making it ready for downstream analysis tasks.
+
+```{r}
+```
+
+The data cleaning process for the raw theft data show in @tbl-raw_data_2 was designed to ensure the dataset is accurate, consistent, and ready for analysis.
+First, the required libraries such as `tidyverse` for data manipulation, `janitor` for standardizing column names, `here` for handling file paths, and `arrow` for managing Parquet files were loaded.
+Additionally, a directory structure was established to store the cleaned data in a designated folder.
+This setup provided a solid foundation for the cleaning process.
+
+To validate the data, predefined lists of valid values for certain fields were established.
+Specifically, a set of valid division codes (e.g., "D51", "D42") and offense categories (e.g., "Theft From Motor Vehicle Under", "Other Theft") was created to filter out any unrecognized or irrelevant entries.
+This ensured that only records meeting specific criteria would be retained in the final dataset.
+
+The raw crime data was loaded from a CSV file, and column names were standardized using the `clean_names()` function from the `janitor` package.
+Key columns, such as `report_date` and `occ_date`, were converted into date formats, while latitude and longitude fields were cast to numeric values for validation.
+The `report_hour` column was converted into integers, and a new column, `report_dow`, was added to capture the day of the week from the `report_date`.
+
+Rows with missing critical values, such as `event_unique_id`, `report_date`, `division`, `offence`, or geographic coordinates, were removed to maintain data quality.
+Additionally, latitude and longitude values were validated to ensure they fell within acceptable ranges (latitude: -90 to 90, longitude: -180 to 180).
+Text fields such as `division` and `offense` were standardized by trimming extra spaces and converting values to uppercase.
+Records with division codes or offenses not included in the predefined valid lists were excluded from the dataset.
+To further ensure data integrity, duplicate rows based on `event_unique_id` were removed.
+
+The dataset was further refined by selecting only relevant columns, such as `event_unique_id`, `report_date`, `division`, `location_type`, and geographic coordinates.
+Columns were renamed to make them more intuitive, such as renaming `event_unique_id` to `event_id` and `report_date` to `report`.
+This helped to simplify and clarify the structure of the data.
+
+Finally, the cleaned dataset was saved as a Parquet file (`cleaned_crime_data.parquet`) using the `write_parquet()` function.
+The Parquet format was chosen for its efficiency in storage and processing, making the dataset ready for further analysis.
+Examples of cleaned theft data is presented in @tbl-cleaned_theft.
+This comprehensive cleaning process resulted in a high-quality dataset that is well-prepared for any downstream tasks.
+
+```{r}
+```
 

### 2024-12-03 paper/paper.qmd
Add reference section
Finished the whole paper

@@
 abstract: "This study examines motor vehicle theft and traffic collisions in Toronto, focusing on spatial and temporal patterns,and environmental factors.
-A composite risk score model is used and reveals that collisions occur more frequently under unfavorable road conditions, such as poor lighting and wet or icy surfaces, and in areas with inadequate traffic control measures.
+A composite risk score model is used and highlights that collisions occur more frequently under unfavorable road conditions, such as poor lighting and wet or icy surfaces, and in areas with inadequate traffic control measures.
 These findings emphasize the need for targeted interventions, including enhanced road infrastructure and lighting in collision-prone zones.
@@
 @sec-model introduces the logistic regression models applied to estimate the probability of collision, as well as the method used to derive the theft index and combine these probabilities into a composite risk score.
-@sec-result then presents the main findings, including insights into how different factors contribute to the risks of owning and riding a motorbike.
+@sec-result then presents the main findings, including understandings into how different factors contribute to the risks of owning and riding a motorbike.
 Finally, @sec-discussion interprets the results, highlighting significant trends and implications for motorbike risk assessment, and concludes with a discussion on the limitations of the analysis and future research directions.
@@
 
-Details about the data cleaning process and the criteria for variable selection are provided in Appendix A.
+Details about the data cleaning process and the criteria for variable selection are provided in [Appendix -@sec-data_details].
 
@@
 The dataset excludes "unfounded" occurrences, adhering to Statistics Canada’s definition that these events were determined not to have occurred or been attempted.
-Like the KSI dataset, this dataset includes fields for both the old and new neighborhood structures, enabling comprehensive geographic analyses of theft trends.
+Like the KSI dataset, this dataset includes fields for both the old and new neighborhood structures, enabling solid geographic analyses of theft trends.
 
@@
 This indicates that most motorbike-related risks fall within a moderate range.
-The density curve overlay reveals a smooth progression in risk levels, with the peak occurring around a Risk Index value of 0.55, suggesting that this is the most common level of composite risk.
+The density curve overlay indicates a smooth progression in risk levels, with the peak occurring around a Risk Index value of 0.55, suggesting that this is the most common level of composite risk.
 The left tail, representing lower risk levels, is relatively small, while the right tail, corresponding to higher risks, extends further, indicating the presence of a smaller number of high-risk cases.
@@
 The extended tail on the higher end of the Risk Index suggests that certain environmental or situational factors disproportionately increase the risks in specific cases.
-This insight can inform targeted interventions, focusing on the outliers with high Risk Index values to mitigate the most critical risks.
+This idea can inform targeted interventions, focusing on the outliers with high Risk Index values to mitigate the most significant risks.
 
@@
 Each category represents a grouping of neighborhoods based on their average Risk Index values.
-The plot reveals distinct differences in the distribution of the Risk Index between these categories.
+The plot illustrates distinct differences in the distribution of the Risk Index between these categories.
 
@@
 This temporal variability highlights specific periods of increased theft risk, enabling the model to capture these patterns effectively.
-Understanding this trend allows for better predictions of theft dynamics and offers valuable insights for designing targeted interventions and optimizing resource allocation based on the time of day.
+Understanding this trend allows for better predictions of theft dynamics and offers meaningful ideas for designing targeted interventions and optimizing resource allocation based on the time of day.
 
@@
 The Collision Probability model, on the other hand, examines situational and environmental conditions affecting the likelihood of collisions.
-Variables such as Neighborhood ID, Road Surface Condition, Lighting Condition, Traffic Control, Time of Day, and Visibility offer insights into the contextual factors that contribute to traffic incidents, capturing the dynamic interaction between environmental factors and human behavior.
+Variables such as Neighborhood ID, Road Surface Condition, Lighting Condition, Traffic Control, Time of Day, and Visibility provides an understanding into the contextual factors that contribute to traffic incidents, capturing the dynamic interaction between environmental factors and human behavior.
 
@@
 
-The **Traffic Control** variable categorizes the type of traffic management present at the site of each incident, such as "No Control," "Traffic Signal," or "Stop Sign." This variable provides insight into how various traffic control measures influence the likelihood of motorbike collisions.
+The **Traffic Control** variable categorizes the type of traffic management present at the site of each incident, such as "No Control," "Traffic Signal," or "Stop Sign." This variable provides ideas into how various traffic control measures influence the likelihood of motorbike collisions.
 
@@
 
-Detailed model diagnostics, variable descriptions, and performance metrics are included in [Appendix @sec-model-details].
-
 ## Model Set-Up
@@
 The sub-index method for theft risk is tailored to the dataset’s characteristics, avoiding assumptions about unobserved cases, and the use of logistic regression for collision risk ensures robust and reliable predictions.
-Together, these components form a comprehensive and practical framework for evaluating motorbike ownership and usage risks, addressing both immediate safety concerns and long-term theft risks.
+Together, these components form a practical and practical framework for evaluating motorbike ownership and usage risks, addressing both immediate safety concerns and long-term theft risks.
 
@@
 where P represents the predicted probability of collision from the logistic regression model plotted against the continuous predictors.
-These plots are intended to visualize the relationship between each predictor and the logit of the outcome variable, providing insight into whether the relationship is approximately linear.
+These plots are intended to visualize the relationship between each predictor and the logit of the outcome variable, providing thoughts into whether the relationship is approximately linear.
 
@@
 Conversely, Road Conditions\` account for the physical state of the driving surface, such as wet, icy, or damaged roads, which influence vehicle handling and stopping distance.
-Together, these variables capture complementary aspects of collision risk, ensuring that the model provides a comprehensive assessment.
+Together, these variables capture complementary aspects of collision risk, ensuring that the model provides a solid assessment.
 
@@
 
-The year 2020 coincides with the beginning of the COVID-19 pandemic, which had a profound impact on urban mobility patterns.
+The year 2020 coincides with the beginning of the COVID-19 pandemic, which had a strong impact on urban mobility patterns.
 Lockdowns, social distancing measures, and reduced economic activities likely led to significant decreases in traffic volumes and changes in behavior related to motor vehicle usage [@buehler2021covid].
@@
 This section examines how factors such as road surface conditions, lighting conditions, and traffic control types contribute to changes in the Risk Index.
-Categorizing the data based on these factors reveals specific conditions under which the likelihood of collisions or theft increases.
+Categorizing the data based on these factors illustrates specific conditions under which the likelihood of collisions or theft increases.
 
@@
 This analysis explores whether different surface types—such as dry, wet, icy, and others—affect the level of risk associated with traffic incidents.
-The following violin plot visually compares the Risk Index across different road surface conditions, providing insight into how these factors may contribute to the overall risk.
+The following violin plot visually compares the Risk Index across different road surface conditions, providing understandings into how these factors may contribute to the overall risk.
 ```{r}
@@
 
-@fig-risk-distribution-roadcondition reveals notable differences in the Risk Index distribution across various road surface conditions.
+@fig-risk-distribution-roadcondition highlights notable differences in the Risk Index distribution across various road surface conditions.
 Wet and icy conditions, in particular, show higher variability in the Risk Index, suggesting that these surfaces may lead to more unpredictable and elevated risk levels.
@@
 
-Dark Conditions in (@fig-risk-index-lighting-conditions-1) highlights a concentration of Risk Index values in the moderate-to-high range, indicating that low visibility during dark conditions significantly contributes to elevated collision risks.
+Dark Conditions in @fig-risk-index-lighting-conditions-1 highlights a concentration of Risk Index values in the moderate-to-high range, indicating that low visibility during dark conditions significantly contributes to elevated collision risks.
 These findings underscore the importance of improved street lighting or reflective road markers in mitigating risks under dark conditions.
 
-Daylight Conditions distrubition in (@fig-risk-index-lighting-conditions-2) shows a peak at slightly lower Risk Index values compared to dark conditions, reflecting the safety advantages of enhanced visibility.
+Daylight Conditions distrubition in @fig-risk-index-lighting-conditions-2 shows a peak at slightly lower Risk Index values compared to dark conditions, reflecting the safety advantages of enhanced visibility.
 Daylight conditions reduce collision risks, likely due to clear visibility and the predictability of traffic patterns during the day.
@@
 
-Artificial Conditions in (@fig-risk-index-lighting-conditions-4) shows a moderate-to-high concentration of Risk Index values, suggesting that artificial lighting, while helpful, does not fully mitigate risks associated with low visibility.
+Artificial Conditions in @fig-risk-index-lighting-conditions-4 shows a moderate-to-high concentration of Risk Index values, suggesting that artificial lighting, while helpful, does not fully mitigate risks associated with low visibility.
 Enhanced lighting technologies and maintenance of artificial lights could further improve safety outcomes under these conditions.
@@
 This clustering in the downtown area suggests that even neighborhoods with moderate or lower collision risks can experience a high frequency of incidents due to the dense urban environment, heavy traffic flows, and increased walkers and cyclist activity typical of central urban areas.
-This observation highlights the importance of comprehensive urban traffic management strategies across risk levels to improve safety in both high- and low-risk neighborhoods.
+This observation highlights the importance of urban traffic management strategies across risk levels to improve safety in both high- and low-risk neighborhoods.
 
@@
 The datasets used in the analysis contain missing or incomplete entries, particularly for variables such as road surface conditions and lighting.
-These gaps may introduce bias or underrepresent specific risk factors, potentially affecting the overall robustness of the findings.
+These limitations may introduce bias or underrepresent specific risk factors, potentially affecting the overall realiability of the findings.
 Additionally, the Risk Index relies on fixed weights for theft and collision components, which may oversimplify the relative importance of these factors.
@@
 These models could provide more immediate and accurate assessments of risk, enabling proactive interventions.
-Another area for exploration is the impact of seasonal and event-based variations.
+Another area for research is the impact of seasonal and event-based variations.
 Examining how risk patterns shift during holidays, festivals, or different seasons would allow for more refined time-sensitive recommendations for both motorbike owners and policymakers.
@@
 The inclusion of socioeconomic indicators in future studies is also critical.
-Variables such as income levels, unemployment rates, and population density could offer a more comprehensive view of the root causes of spatial disparities in risk.
-Additionally, future research should delve into the interaction effects between environmental and situational factors.
+Variables such as income levels, unemployment rates, and population density could offer a more thorough view of the root causes of spatial disparities in risk.
+Additionally, future research should explore the interaction effects between environmental and situational factors.
 For example, investigating how wet roads combined with poor lighting conditions affect risk levels could inform more nuanced interventions.
@@
 #| label: tbl-raw_data_1
-#| tbl-cap: "Examples of raw collision data"
+#| tbl-cap: "Examples of Raw Collision Data"
 #| echo: false
@@
     "60", "Woodbine-Lumsden", "60", "Woodbine-Lumsden (60)", "D55", 
-    '{"type": "MultiPoint", "coordinates": [[-79.318797000266, 43.6995950001932]]}', 
+    '{"type": "MultiPoint", "coordinates": [[-79.318797, 43.699595]]}', 
     "N/A"  # Added the missing value
@@
     align = "l",
-    caption = "Table 1: Example Data from Motor Vehicle Collisions Dataset"
   ) %>%
@@
 #| label: tbl-raw_data_2
-#| tbl-cap: "Examples of raw theft data"
+#| tbl-cap: "Examples of Raw Theft Data"
 #| echo: false
@@
     "200", "Theft From Motor Vehicle Under", "NonMCI", "73", "-79.37453055088850", 
-    "43.65706729617110", '{"type": "MultiPoint", "coordinates": [[-79.374530551, 43.657067296]]}'
+    "43.65706729617110", '{"type": "MultiPoint", "coordinates": [[-79.3745305, 43.6570672]]}'
   )
@@
     col.names = c("Variable Name", "Example Value"),
-    align = "l",
-    caption = "Table 1: Example Data for Reported Incidents"
+    align = "l"
   ) %>%
@@
 Essential libraries were loaded at the start of the process, including `tidyverse` for data manipulation, `janitor` for standardizing column names, `here` for managing file paths, and `arrow` for saving the cleaned data in an efficient Parquet format.
-These tools provided a comprehensive framework for handling various cleaning tasks effectively and ensuring the dataset was ready for further use.
+These tools provided a solid framework for handling various cleaning tasks effectively and ensuring the dataset was ready for further use.
 
@@
 
-In summary, this cleaning process ensured the traffic data was free of duplicates and invalid entries, featured consistent and intuitive column names, and had validated geographic information.
-The resulting dataset was stored in an optimized format, making it ready for downstream analysis tasks.
-
-```{r}
+In summary, this cleaning process ensured the collision data was free of duplicates and invalid entries, featured consistent and intuitive column names, and had validated geographic information.
+The resulting cleaned dataset show in @tbl-cleaned_collision_data was stored in an optimized format, making it ready for downstream analysis tasks.
+
+```{r}
+#| label: tbl-cleaned_collision_data
+#| tbl-cap: "Example of Cleaned Collision Data"
+#| echo: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+
+# Create the example data
+collision_data_example <- tibble(
+  Variable = c(
+    "id", "accident_id", "report_date", "time_of_day", "street_primary", 
+    "street_secondary", "offset", "road_type", "district", "accident_location", 
+    "traffic_control", "visibility_conditions", "lighting_conditions", 
+    "road_conditions", "accident_classification", "impact_type", "invtype", 
+    "invage", "injury", "fatal_no", "initdir", "vehicle_type", "maneuver", 
+    "driver_action", "driver_condition", "pedtype", "pedact", "pedcond", 
+    "cyclistype", "cycact", "cyccond", "pedestrian", "cyclist", "automobile", 
+    "motorcycle", "truck", "trsn_city_veh", "emerg_veh", "passenger", "speeding", 
+    "ag_driv", "redlight", "alcohol", "disability", "hood_158", "neighborhood", 
+    "hood_140", "neighbourhood_140", "division"
+  ),
+  Example = c(
+    "1", "893184", "2006-01-01", "236", "WOODBINE AVE", "O CONNOR DR", "NA", 
+    "Major Arterial", "Toronto and East York", "Intersection Related", 
+    "No Control", "Clear", "Dark", "Wet", "Non-Fatal Injury", "Approaching", 
+    "Passenger", "NA", "Major", "NA", "NA", "NA", "NA", "NA", "NA", "NA", 
+    "NA", "NA", "NA", "NA", "NA", "NA", "NA", "Yes", "NA", "NA", "NA", "NA", 
+    "Yes", "Yes", "Yes", "NA", "Yes", "NA", "60", "Woodbine-Lumsden", "60", 
+    "Woodbine-Lumsden (60)", "D55"
+  )
+)
+
+# Render the table vertically
+collision_data_example %>%
+  kable(
+    col.names = c("Variable Name", "Example Value"),
+    align = "l",
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  ) %>%
+  column_spec(2, width = "20em")  # Adjust width to handle longer text
+
 ```
@@
 Examples of cleaned theft data is presented in @tbl-cleaned_theft.
-This comprehensive cleaning process resulted in a high-quality dataset that is well-prepared for any downstream tasks.
-
-```{r}
-```
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-
-```
-
+This cleaning process resulted in a high-quality dataset that is well-prepared for any downstream tasks.
+
+```{r}
+#| label: tbl-cleaned_theft
+#| tbl-cap: "Example of Cleaned Theft Data"
+#| echo: false
+
+# Load necessary libraries
+library(tidyverse)
+library(kableExtra)
+
+# Create the example data
+cleaned_data_example <- tibble(
+  Variable = c(
+    "event_id", "report", "occurrence", "hour", "day_of_week", "division",
+    "location", "premises_type", "offense", "mci_category", "hood_158", 
+    "longitude", "latitude"
+  ),
+  Example = c(
+    "GO-20141261501", "2014-01-01", "2014-01-01", "8", "Wednesday", "D51", 
+    "Single Home, House (Attach Garage, Cottage, Mobile)", "House", 
+    "Theft From Motor Vehicle Under", "NonMCI", "73", "-79.37453055088850", 
+    "43.65706729617110"
+  )
+)
+
+# Render the table vertically
+cleaned_data_example %>%
+  kable(
+    col.names = c("Variable Name", "Example Value"),
+    align = "l"
+  ) %>%
+  kable_styling(
+    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
+    full_width = FALSE
+  ) %>%
+  column_spec(2, width = "20em")  # Adjust width for better text wrapping
+
+```
+
+## Predictor Variable Choice
+
+### Theft Index
+
+In constructing the theft risk sub-indexes, the choice of predictor variables: Hour of the Day, Day of the Week, and Premises Type was informed by both theoretical and practical considerations rooted in spatial, temporal, and environmental factors associated with theft incidents.
+These variables were selected from the cleaned theft dataset to create a strong and interpretable model that captures key patterns of theft occurrences while aligning with the study's goal of identifying targeted interventions.
+
+1. Hour of the Day
+
+The inclusion of the Hour of the Day as a predictor variable reflects the strong temporal dependency of theft occurrences.
+Theft risk often varies significantly across a 24-hour period due to human activity patterns, such as peak commuting times, work hours, and late-night inactivity.
+For example, motor vehicle thefts are more likely to occur during nighttime hours when there is reduced surveillance and fewer bystanders to deter criminal behavior.
+By incorporating the Hour of the Day into the theft risk sub-index, the model accounts for this temporal dimension, enabling stakeholders to identify high-risk hours and implement time-specific interventions, such as increased nighttime patrols or surveillance in parking lots and residential areas.
+
+2. Day of the Week
+
+The Day of the Week was selected to capture the variations in theft risk across different days, which often reflect behavioral patterns linked to work, leisure, and mobility.
+Weekends, for instance, may see heightened theft activity in areas with nightlife or during events that attract large crowds, while weekdays might exhibit theft patterns associated with commuter parking zones or commercial districts.
+The inclusion of this variable allows the model to identify day-specific risks and tailor interventions, such as targeted surveillance during high-risk days or public awareness campaigns about theft prevention.
+
+3. Premises Type
+
+Premises Type represents the spatial dimension of theft risk, capturing the environmental characteristics where thefts are most likely to occur.
+For example, thefts in residential areas (e.g., houses, apartments) often occur when vehicles are parked in driveways or on the street with limited security measures.
+Similarly, commercial areas and outdoor public spaces may pose unique risks due to their high foot traffic, limited oversight during certain hours, or accessibility of parked vehicles.
+Including Premises Type in the sub-index allows the model to differentiate risk levels based on spatial characteristics and recommend location-specific strategies, such as installing better lighting in outdoor parking areas or promoting secure storage options for residential settings.
+
+### Collision Probability
+
+The predictors chosen for the logistic regression model were carefully selected to capture key spatial, temporal, and environmental factors that influence the likelihood of severe collisions.
+These variables were derived from the cleaned collision dataset, ensuring data quality and relevance to the study's focus on urban mobility and safety in Toronto.
+
+1. Neighborhood ID
+
+The inclusion of hood_158, a unique identifier for neighborhoods, allows the model to capture spatial patterns in collision occurrences.
+Different neighborhoods may exhibit varying levels of collision risk due to factors such as traffic density, road design, infrastructure quality, and socioeconomic conditions.
+By including this variable, the model can account for geographic heterogeneity, identifying high-risk areas and enabling targeted interventions such as infrastructure improvements or traffic calming measures in specific neighborhoods.
+
+2. Road Surface Condition
+
+Road surface conditions, such as dry, wet, or icy surfaces, significantly impact vehicle control and braking effectiveness, making this variable a critical predictor.
+Collisions are more likely under adverse surface conditions, such as rain, snow, or ice, due to reduced traction and visibility.
+Including this variable in the model ensures that environmental factors influencing vehicle dynamics are accounted for, providing insights into how weather and road maintenance contribute to collision risk.
+This information can guide policies such as improved road drainage systems or timely snow and ice removal.
+
+3. Lighting Condition
+
+Lighting conditions, which describe whether collisions occur during daylight, nighttime, or under artificial light, are essential for understanding the role of visibility in collisions.
+Poor lighting can reduce drivers' ability to perceive hazards, increasing the likelihood of accidents.
+By incorporating this variable, the model can assess how lighting contributes to collision risk, highlighting the importance of enhanced street lighting in poorly lit areas or during nighttime hours.
 \newpage
 
+4. Traffic Control
+
+The presence and type of traffic control devices, such as stop signs, traffic signals, and pedestrian crossovers, play a important role in preventing collisions by regulating vehicle flow and ensuring predictable driver behavior.
+The inclusion of this variable allows the model to evaluate the effectiveness of existing traffic control measures and identify gaps, such as intersections without adequate control devices, where collisions are more likely to occur.
+This idea can inform infrastructure upgrades and the installation of additional traffic management systems.
+
+5. Road Visibility
+
+Road visibility conditions, influenced by factors like fog, rain, or snow, are directly linked to drivers' ability to perceive their surroundings and react to potential hazards.
+Poor visibility increases the risk of collisions by reducing reaction time and decision-making effectiveness.
+Including this variable ensures that the model captures the impact of adverse weather and environmental conditions on collision likelihood, emphasizing the need for weather-responsive traffic management strategies, such as variable speed limits or enhanced signage.
+
+6. Time of Day
+
+The Time of Day is a key temporal variable that captures variations in collision risk throughout a 24-hour period.
+Collision likelihood often peaks during specific hours, such as morning and evening rush hours, due to increased traffic volume, or late at night, when factors like fatigue or impaired driving may come into play.
+Including this variable allows the model to account for temporal patterns in collision occurrences, supporting the implementation of time-specific measures, such as increased traffic enforcement during high-risk hours or adjustments to traffic light timings.
+
+\newpage
+
 # Idealized Methodology for a Survey on Motor Risk {#sec-survey}
 
-## Survey Overview
-
-This survey aims to assess factors contributing to motorbike risks, including theft and collisions, by gathering data from motorbike owners and riders.
-The survey will explore demographic characteristics, riding behavior, motorbike usage patterns, and perceptions of environmental and situational risks.
-
-## Sampling Approach
-
-The survey will use a stratified random sampling method to ensure diverse representation across:
-
--   Geographic regions (urban, suburban, rural areas).
--   Demographic groups (age, gender, income, and education levels).
--   Riding experience (novices, intermediate, and experienced riders).
-
-The target population includes licensed motorbike riders and owners. A sample size of approximately 1,000 respondents is proposed to ensure statistical validity across strata.
-
-## Survey structure
-
-The survey will consist of five main sections:
-
--   Demographics: Basic information on respondents (e.g., age, gender, income, education, geographic location).
-
--   Riding Behavior: Frequency, duration, and purpose of motorbike usage.
-
--   Risk Awareness and Perception: Personal experiences with collisions or theft and perceptions of environmental risks.
-
--   Situational Factors: Details of riding conditions such as time of day, weather, road surface, and lighting.
-
--   Preventive Measures: Actions taken by riders to mitigate theft or collision risks (e.g., use of locks, helmets, or avoiding high-risk areas).
+## Survey Objectives
+
+The primary goal of this survey is to investigate the factors contributing to motorbike risks, including theft and collisions, by collecting data from motorbike owners and riders.
+This research aims to understand the interplay of demographic, behavioral, and environmental factors that influence risk levels.
+Insights gained from the survey will help develop targeted interventions to enhance motorbike safety and security.
+
+## Sampling Methodology
+
+The target population for this survey includes licensed motorbike riders and owners across diverse geographic regions.
+The survey will focus on individuals who have used motorbikes within the past year to ensure relevance and recency of responses.
+
+A stratified random sampling approach will be employed to capture diverse perspectives and ensure statistical validity. Stratification will be based on:
+
+Geographic regions: Urban, suburban, and rural areas to understand variations in risks and behaviors.
+Demographic characteristics: Age, gender, income, and education levels to capture socioeconomic diversity.
+Riding experience: Novice, intermediate, and experienced riders to assess how skill and familiarity influence risks.
+A sample size of approximately 1,000 respondents is proposed, distributed proportionally across the strata to ensure adequate representation for comparative analysis.
+
+## Survey structure and Content
+
+The survey is structured into six main sections, each designed to gather detailed information about factors influencing motorbike risks.
+
+**1. Demographics**
+
+This section collects essential background information to contextualize the respondent's behaviors and risks:
+
+- Age
+- Gender
+- Education level
+- Income range
+- Geographic location (urban, suburban, rural)
+- Motorbike ownership status
+
+**2. Motorbike Usage Patterns**
+
+Questions in this section aim to understand how respondents use their motorbikes:
+
+- Frequency of use (e.g., daily, weekly, monthly)
+- Primary purpose (e.g., commuting, recreation, delivery work)
+- Average trip duration
+- Type of roads frequently used (e.g., highways, city streets, rural roads)
+- Riding seasonality (e.g., year-round, summer-only)
+
+**3. Riding Behavior**
+
+This section looks into behaviors that may influence collision and theft risks:
+
+- Average speed during rides
+- Use of safety gear (e.g., helmets, gloves, reflective clothing)
+- Adherence to traffic laws (e.g., stopping at signals, avoiding speeding)
+- Riding under adverse conditions (e.g., rain, low light)
+- Group riding versus solo riding
+
+**4. Situational Risk Factors**
+
+This section explores external factors that contribute to motorbike risks:
+
+- Time of day most frequently ridden (morning, afternoon, night)
+- Common weather conditions experienced (e.g., sunny, rainy, foggy)
+- Road surface conditions during rides (e.g., dry, wet, icy)
+- Presence of traffic control measures on roads used
+- Experience with poorly lit or visibility-compromised areas
+
+**5. Risk Awareness and Experience**
+
+This section assesses respondents' awareness of risks and their personal experiences:
+
+- Perception of collision and theft risks in their area
+- Personal experience with motorbike theft (e.g., number of thefts, recovered cases)
+- Personal experience with collisions (e.g., frequency, severity, injuries sustained)
+- Awareness of high-risk areas for theft or collisions
+- Perception of environmental risks (e.g., poor road design, lack of lighting)
+
+**6. Preventive Measures and Interventions**
+
+This section investigates actions taken by riders to mitigate risks and their opinions on potential interventions:
+
+- Security measures used to prevent theft (e.g., locks, alarms, GPS trackers)
+- Riding adjustments to avoid collisions (e.g., avoiding high-risk areas, reducing     speed)
+- Willingness to participate in safety programs (e.g., high-level riding courses)
+  Opinions on public safety measures (e.g., improved lighting, better road   
+  infrastructure
+- Suggestions for reducing theft and collisions in their area
 
@@
 
--   Closed-ended questions: For quantitative data (e.g., multiple choice, Likert scales, ranking).
-
--   Open-ended questions: To capture nuanced insights and personal experiences.
-
--   Matrix questions: To evaluate attitudes across multiple dimensions efficiently.
-
-### Question List
-
-Here are examples of survey questions:
-
-Demographics: What is your age group? (e.g., 18–24, 25–34, etc.) What is your highest level of education completed?
-
-Riding Behavior: How often do you ride your motorbike? (Daily, Weekly, Monthly, Rarely) For what purposes do you primarily use your motorbike? (Commuting, Recreation, Delivery/Work, Other)
-
-Risk Awareness and Perception: On a scale of 1–5, how would you rate the theft risk in your neighborhood? Have you experienced a motorbike theft or collision in the past year? (Yes/No)
-
-Situational Factors: What time of day do you usually ride? (Morning, Afternoon, Evening, Night) In what weather conditions do you typically ride? (Clear, Rainy, Snowy)
-
-Preventive Measures: What measures do you take to secure your motorbike against theft?
-(Select all that apply: Lock, Alarm, GPS Tracker, Parking in Secure Locations) Do you avoid specific times or areas due to perceived collision risks?
-(Yes/No)
+To ensure the survey collects meaningful, diverse, and actionable data, a combination of question types will be used.
+Each type serves a specific purpose, enabling both quantitative analysis and qualitative understanding.
+Below are the key question types included in the survey:
+
+### Questionnaire Design
+
+The questionnaire will include a mix of closed-ended and open-ended questions.
+Closed-ended questions will use Likert scales, multiple-choice options, and ranking formats, while open-ended questions will provide opportunities for respondents to share detailed thoughts.
+
+1. Closed-Ended Questions
+
+These questions offer predefined answer choices and are used to gather quantitative data. They are easy to analyze and are suitable for identifying patterns and trends.
+
+2. Open-Ended Questions
+
+Open-ended questions allow respondents to express their thoughts and experiences in detail. They are useful for collecting qualitative data and identifying informatiosn
+
+3. Matrix Questions
+
+Matrix questions allow multiple related items to be rated on the same scale, making it easier to assess various factors within a single question.
+
+4. Ranking Questions
+
+Ranking questions require respondents to prioritize options based on their preferences or perceived importance. They are effective for understanding preferences and trade-offs.
+
+5. Dichotomous Questions
+
+These questions have only two answer choices and are typically used to gather straightforward, binary data.
+
+6. Drop-Down Questions
+
+Drop-down questions are useful for limiting the display of options while allowing respondents to choose from a broad range of answers.
+
+7. Conditional/Follow-Up Questions
+
+Conditional questions are used to customize the survey based on previous responses, ensuring relevance and efficiency.
+
+#### Question List
+
+**Demographics**
+
+1. What is your age group?
+
+- Under 18
+
+- 18-24
+
+- 25-34
+
+- 35-44
+
+- 45-54
+
+- 55+
+
+2. What is your highest level of education?
+
+- High school or less
+
+- Some college or university
+
+- Bachelor’s degree
+
+- Master’s degree
+
+- Doctorate or professional degree
+
+**Motorbike Usage Patterns**
+
+3. How frequently do you use your motorbike?
+
+- Daily
+
+- Weekly
+
+- Monthly
+
+- Rarely
+
+- Other
+
+4. How important are the following reasons for riding a motorbike to you?
+
+- Commuting
+
+- Recreation
+
+- Delivery work
+
+- Other
+
+(Rank each reason: Not important at all, Slightly important, Moderately important, Very important, Extremely important)
+
+**Riding Behavior**
+
+5. How often do you wear a helmet while riding?
+
+- Never
+
+- Rarely
+
+- Sometimes
+
+- Often
+
+- Always
+
+6. What is your average speed while riding?
+
+- Below 40 km/h
+
+- 40-60 km/h
+
+- 60-80 km/h
+
+- Above 80 km/h
+
+**Situational Risk Factors**
+
+7. What time of day do you most often ride?
+
+- Morning (6 AM–12 PM)
+
+- Afternoon (12 PM–6 PM)
+
+- Night (6 PM–12 AM)
+
+- Early morning (12 AM–6 AM)
+
+8. How often do you encounter poor road visibility during rides?
+
+- Never
+
+- Rarely
+
+- Sometimes
+
+- Often
+
+**Risk Awareness and Experience**
+
+9. Have you ever had your motorbike stolen?
+
+- Yes
+
+- No
+
+(If "Yes," follow-up: How many times has your motorbike been stolen?)
+
+- Once
+
+- 2-3 times
+
+- More than 3 times
+
+10. How would you rate the risk of collisions in your area?
+
+- Very low
+
+- Low
+
+- Moderate
+
+- High
+
+- Very high
+
+11. How important are the following factors for motorbike safety?
+
+- Helmet use
+
+- Speed limits
+
+- Visibility
+
+(Rate each factor: Not important at all, Slightly important, Moderately important, Very important, Extremely important)
+
+**Preventive Measures**
+
+12. What security measures do you currently use to prevent motorbike theft?
+(Select all that apply)
+
+- Locks
+
+- Alarms
+
+- GPS trackers
+
+- Other
+
+13. How effective do you think these changes would be in reducing motorbike collisions in your area?
+
+(Rate each factor: Not effective at all, Slightly effective, Moderately effective, Very effective, Extremely effective)
+
+- Improved road lighting
+
+-Better road maintenance (e.g., fixing potholes, clearing debris)
+
+- Increased police patrols in high-risk areas
+
+- Public education on motorbike safety
+
+- Enforcing speed limits
+
+14. What changes do you think would most effectively reduce motorbike collisions in your area?
+(Open-ended)
 
@@
 
-Participants will be recruited through a combination of online and offline channels:
-
-Online platforms: Motorbike owner forums, social media groups, and email lists of motorbike organizations.
-
-Offline channels: Flyers at motorbike dealerships, repair shops, and riding schools. Incentives such as small gift cards or entry into a raffle may be provided to encourage participation.
+To ensure a diverse and representative sample, participants will be recruited through a combination of online and offline channels, targeting motorbike riders and owners across various demographics, regions, and riding experiences.
+Online recruitment will utilize motorbike-related communities and platforms, including forums, social media groups, and email lists maintained by motorbike organizations.
+These platforms allow direct engagement with individuals who are likely to be interested in the survey, ensuring efficient outreach to a large pool of potential respondents.
+Social media advertisements and posts in popular groups dedicated to motorbike enthusiasts will also be utilized to broaden the reach and attract participants from different backgrounds.
+
+Offline recruitment will complement the online efforts by targeting physical spaces frequented by motorbike riders.
+Flyers and posters will be distributed at motorbike dealerships, repair shops, riding schools, and popular gathering spots for riders.
+These locations are strategic, as they attract individuals actively engaged in motorbike use and maintenance, making them ideal candidates for the survey.
+To further incentivize participation, small rewards such as gift cards or entries into a raffle will be offered.
+These incentives are designed to encourage participation while demonstrating appreciation for the respondents’ time and effort.
+The combination of online and offline recruitment strategies ensures that the survey captures a wide range of perspectives, improving the reliability and generalizability of the findings.
 
@@
 
-The survey design is informed by prior studies on motor vehicle risk perception, road safety, and crime prevention. Key references include:
-
-Research on environmental and situational predictors of road accidents. Studies on the effectiveness of theft prevention measures. Literature on sampling methods and survey design for risk assessment.
+The design of this survey is grounded in a solid structure of literature on motor vehicle risk perception, road safety, and crime prevention.
+Studies on environmental and situational predictors of road accidents have informed the inclusion of variables such as time of day, weather conditions, road surface quality, and lighting [@ling2020cyclist].
+Research has consistently highlighted these factors as critical determinants of collision risk, guiding the formulation of survey questions aimed at understanding how riders perceive and respond to these risks in their daily activities [@fridman2020effect].
+
+Additionally, studies on theft prevention measures have shaped the focus on security practices and riders’ awareness of theft risks.
+Prior research has emphasized the importance of behavioral and environmental factors in preventing motorbike theft, such as the use of locks, secure parking spaces, and community awareness programs [@anderson2014steal].
+These ideas have been incorporated into the survey’s sections on preventive measures and risk awareness, ensuring the collection of data relevant to both individual behaviors and broader community-based interventions.
+
+Finally, the survey’s sampling methodology and design are informed by established literature on survey-based risk assessments.
+Stratified random sampling, a method widely regarded as effective for ensuring representativeness, was chosen based on evidence from prior studies [@harlow1988motor].
+The survey structure, combining closed-ended, open-ended, and conditional questions, is modeled on best practices in survey design to capture both quantitative and qualitative data.
+By integrating findings from previous research, the survey aims to contribute to the ongoing discourse on motorbike safety and theft prevention, while addressing gaps in the literature specific to the Toronto context.
+
+This careful alignment with existing literature not only validates the survey design but also ensures its findings will be relevant and valuable for developing targeted interventions to enhance motorbike safety and security.
+By linking the survey to established research, the study builds on a foundation of evidence, contributing new thoughts into the complex interplay of factors influencing motorbike risks.
+\newpage
 

### 2024-12-03 paper/paper.qmd
changes color of the predictor variable plots

@@
     y = "Latitude",
-    color = "Neighborhood Type",
-    title = "Neighborhood Collision Points in Toronto"
+    color = "Neighborhood Type"
   ) +

### 2024-12-03 other/datasheet/Toronto_Motor_Risk_datasheet.qmd
datasheet rename

---
title: "Datasheet for 'Toronto Motorbike Collision and Theft Risk'"
subtitle: "Supplementary document to \"When Risk Hits the Road: Does Danger Drive a Return?\""
author: 
  - Yingke He
thanks: "Code and data are available at: [https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main](https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main)."
date: Today
date-format: "D MMMM YYYY"
abstract: "This supplementary datasheet accompanies the motorbike risk analysis paper. Adopting the standardized format proposed by Gebru et al. (2021), it documents the motivation, composition, collection process, recommended applications, maintenance, and distribution of the cleaned dataset, available in the analysis data folder of the [GitHub Repository](https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main). The goal is to enhance communication between dataset creators and users while promoting transparency and accountability within the statistical and risk analysis communities."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

# Acknowledgement

All the following questions are extracted from @gebru2021datasheets.

# Motivation

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -  The dataset was created to document and analyze incidents of motor vehicle collisions (KSI dataset) and thefts from motor vehicles within Toronto. Its primary purpose is to assist law enforcement, policymakers, and researchers in identifying trends, risk factors, and areas of concern. Specifically, it addresses gaps in understanding the temporal, spatial, and contextual factors contributing to serious road safety incidents and vehicle thefts. This data also provides a foundation for neighborhood-level analysis and supports the development of targeted interventions for public safety.
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   The datasets were created and published by the Toronto Police Service as part of their commitment to transparency and public accountability. The data is made available through Open Data Toronto, an initiative aimed at providing open access to public datasets for community use and research.
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   The funding for the dataset's creation likely stems from the operating budget of the Toronto Police Service, supported by the City of Toronto. There is no specific mention of external grants or funding in the documentation, suggesting that it is a part of their routine operational and public outreach activities.
4.  *Any other comments?*
    -   The datasets serve as critical tools for urban planning, public safety strategies, and academic research. However, users should note the limitations of the dataset, including potential discrepancies due to privacy adjustments and the lack of guarantees for completeness or timeliness. The initiative reflects broader efforts to leverage open data for improving community outcomes and fostering collaborative problem-solving.

# Composition

1.  *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
    -   The dataset includes two types of instances: motor vehicle collisions and thefts from motor vehicles. Instances in the Motor Vehicle Collisions (KSI dataset) represent individual events where a person was killed or seriously injured, recorded along with details such as severity, type of collision, and location. Instances in the Theft from Motor Vehicles dataset represent theft events categorized by the value of stolen items and recorded with attributes like the date, time, and geographic location. These instances collectively enable an analysis of both theft and collision risks.
2.  *How many instances are there in total (of each type, if appropriate)?*
    -   There are in total 32425 incidents for the theft data, and 18957 incidents for the collision data
3.  *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
    -   The dataset is a sample of reported incidents and does not include all possible instances. It represents events documented by the Toronto Police Service, meaning unreported or undocumented collisions and thefts are excluded. While it may not capture every incident, the dataset broadly reflects patterns across Toronto and includes diverse geographic coverage.
4.  *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
    -   Each instance consists of structured features rather than raw data. For collisions, features include variables such as collision type, severity, and location. For thefts, features include attributes like the value category of stolen items, date, time, and location. The data has undergone cleaning and formatting to ensure usability and consistency.
5.  *Is there a label or target associated with each instance? If so, please provide a description.*
    -   The raw dataset does not include explicit labels or targets. However, in this study, the composite risk score is used as an outcome variable, calculated by integrating probabilities of theft and collisions derived from the dataset’s predictor variables, such as time of day, road conditions, and neighborhood characteristics.
6.  *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
    -   Some instances have missing or incomplete information, such as geographic coordinates recorded as "0, 0" or locations labeled as "Not Specified Area (NSA)." These gaps occur due to unvalidated data or unavailable details at the time of reporting, and while they do not render the dataset unusable, they may affect the precision of certain analyses.
7.  *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
    -   Relationships between instances are not explicitly defined in the dataset. However, implicit relationships can be inferred, such as temporal or spatial clustering of thefts and collisions within specific neighborhoods or during certain time periods, which offer opportunities for broader contextual analysis.
8.  *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
    -   The dataset does not provide predefined splits for training, validation, or testing. Any splits for analysis would need to be created by the user, potentially based on temporal factors (e.g., separating data by year) or geographic boundaries to maintain logical consistency and ensure representativeness.
9.  *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
    -   The dataset contains potential sources of noise and errors, such as location offsets introduced for privacy and placeholder values like "None" or "NSA." Additionally, duplicates and inconsistencies in field entries were addressed during the cleaning process, but some minor errors may persist due to the limitations of the original data collection.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
    -   The dataset is largely self-contained but relies on external frameworks such as Toronto neighborhood structures and definitions of theft categories. While these external resources are publicly available, their long-term stability is not guaranteed. The dataset itself is accessible through Open Data Toronto, ensuring its immediate usability without additional fees or licensing restrictions.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
    -   The dataset does not contain explicitly confidential information, as it is publicly available through Open Data Toronto. However, some details, such as locations of incidents, have been adjusted for privacy (e.g., offset to the nearest intersection) to prevent the identification of exact addresses or individuals. While this reduces the risk of revealing sensitive data, users must still handle the dataset responsibly to avoid inadvertent breaches of privacy.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
    -   The dataset might cause anxiety or discomfort for some individuals, as it involves serious incidents like fatalities, injuries, and thefts. While the data is presented in an aggregated, structured format, its subject matter could evoke emotional responses, particularly for those affected by similar events.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
    -   The dataset includes information that can indirectly identify sub-populations, such as geographic regions (e.g., neighborhoods) and temporal patterns (e.g., time of day, day of the week). However, it does not explicitly identify sub-populations based on demographic factors such as age or gender, as these variables are either not included or anonymized.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
    -   It is not possible to directly identify individuals from the dataset, as personal identifiers are excluded. However, indirect identification might be possible in rare cases by combining detailed temporal, geographic, or contextual information with external data sources. The privacy adjustments (e.g., location offsets) mitigate this risk significantly.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
    -   The dataset includes potentially sensitive information, such as geographic locations of incidents, which could reveal patterns related to crime or traffic safety in specific areas. While no direct identifiers are present, the nature of the data—especially involving fatalities or serious injuries—could be considered sensitive due to its implications for public safety and community well-being.
16. *Any other comments?*
    -   The dataset serves as an invaluable resource for public safety analysis and urban planning, but users must approach it with care, especially regarding privacy and ethical considerations. While efforts have been made to anonymize and adjust sensitive information, the potential for misuse underscores the importance of responsible data handling and interpretation.

# Collection process

1.  *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
    -   The data was collected through reports made to the Toronto Police Service. For motor vehicle collisions, the data reflects incidents documented by police officers based on on-site investigations or reports submitted by the public. For thefts from vehicles, the data was gathered from incident reports submitted by vehicle owners or other witnesses. The data is considered directly observable in terms of the reported events but may include some subjectivity or errors in reporting. Validation processes are not explicitly described, but police documentation likely involved standard procedures to ensure accuracy.
2.  *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
    -   Data collection relied on manual reporting by law enforcement officers, public submissions (e.g., online forms or emergency calls), and administrative data entry systems used by the Toronto Police Service. These mechanisms are part of standard law enforcement operations, which are typically validated through internal quality assurance processes, though specific details about validation are not provided.

3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    -   The dataset is a comprehensive sample of reported incidents rather than a probabilistic or random sample. It includes all documented collisions and thefts meeting specific criteria (e.g., collisions resulting in fatalities or serious injuries). However, unreported incidents are excluded, meaning the dataset represents only the subset of events known to law enforcement.
4.  *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
    -   Data collection was conducted by Toronto Police Service personnel, including officers and administrative staff. Their compensation would be part of their standard salaries as law enforcement and administrative employees. No external crowdworkers or contractors were involved in the collection process.
5.  *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
    -   The data for motor vehicle collisions spans from 2006 to the present, while the theft data covers a defined reporting period, typically several years. The creation timeframe aligns with the reporting and documentation of these events, as the data was recorded contemporaneously with the incidents.
6.  *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   There is no mention of specific ethical review processes for the dataset, as it is operational data collected as part of law enforcement duties. Ethical considerations likely adhere to standard privacy and data protection protocols, such as anonymizing sensitive information and offsetting geographic locations.
7.  *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
    -   The data was collected directly by the Toronto Police Service from individuals reporting incidents, either in person, through emergency calls, or via online submissions. It was subsequently aggregated into structured datasets for public use through Open Data Toronto.
8.  *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
    -   Individuals reporting incidents were likely aware of the data collection as part of the reporting process. While explicit notification language is not described, it is standard for law enforcement to inform individuals that their reports may be used for documentation and analysis.
9.  *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
    -   Consent for data collection is implicit in the act of reporting an incident to law enforcement. Explicit consent for public release of anonymized data was likely not sought, but the dataset complies with privacy regulations, including removing or offsetting personally identifiable information.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
    -   There is no mechanism for individuals to revoke consent for the use of anonymized, aggregated data in this context. Since the data is de-identified and part of public records, it is not subject to withdrawal by individuals under typical privacy frameworks.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
    -   There is no explicit mention of a formal data protection impact analysis (DPIA) or similar review for this dataset. However, the Toronto Police Service has implemented privacy measures, such as anonymizing data and offsetting locations, to minimize risks to individuals. The dataset’s availability through Open Data Toronto implies adherence to public transparency and data protection policies, though the lack of a detailed impact assessment leaves room for further ethical analysis.
12. *Any other comments?*
    -   While the dataset provides valuable insights for public safety and urban planning, a more detailed impact analysis could address potential unintended consequences, such as misrepresentation of neighborhoods or misuse of sensitive information. Documentation of such an analysis would enhance confidence in the dataset’s ethical use.

# Preprocessing/cleaning/labeling

1.  *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
    -   Yes, extensive preprocessing was performed to clean and standardize the data. This included renaming columns for consistency, removing duplicates, validating geographic coordinates, handling missing values, and converting data types (e.g., numeric or categorical). Specific examples include replacing "None" with NA, validating latitude and longitude ranges, and standardizing textual fields like division codes.
2.  *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
    -   The raw data is not explicitly described as being saved alongside the cleaned version. However, given the dataset's origins with Open Data Toronto, the raw data may still be accessible through their repository, albeit in a less processed form.
3.  *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
    -   Yes, the cleaning process relied on open-source software, particularly R [@citeR] and associated libraries, such as tidyverse [@citeTidyverse], Dplyr [@citeDplyr], and here [@citeHere]. These tools are freely available, and detailed preprocessing scripts can likely be reproduced or adapted using their documentation.

4.  *Any other comments?*
    -   The structured cleaning ensures that the dataset is ready for analysis, but consumers should verify cleaning steps for their specific use cases. Maintaining access to the raw data would also support greater flexibility and reproducibility in future research.

# Uses

1.  *Has the dataset been used for any tasks already? If so, please provide a description.*
    -   Yes, the dataset has been used to develop a composite risk score model in this study, integrating probabilities of theft and collisions to assess motorbike-related risks in Toronto. It has also likely been used for public safety analyses, urban planning, and policy development.
2.  *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
    -   There is no specific repository linking papers or systems that use this dataset. However, Open Data Toronto serves as the primary source for accessing the dataset, and related studies may cite it in their publications.
3.  *What (other) tasks could the dataset be used for?*
    -   The dataset could be applied to tasks such as traffic safety evaluations, urban mobility studies, insurance risk assessments, crime prevention strategies, and machine learning applications for predictive modeling of safety incidents.
4.  *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
    -   The dataset's reliance on reported incidents means it may underrepresent unreported events, particularly in neighborhoods with lower reporting rates. Privacy adjustments, such as location offsets, could affect the precision of geographic analyses. Users should carefully interpret results to avoid bias or misrepresentation of communities and ensure ethical use.
5.  *Are there tasks for which the dataset should not be used? If so, please provide a description.*
    -   The dataset should not be used for tasks that could harm individuals or communities, such as creating unfair stereotypes about specific neighborhoods or using the data for discriminatory purposes. It is also unsuitable for highly granular analyses requiring exact locations due to privacy adjustments.
6.  *Any other comments?*
    -   This dataset is a valuable resource for research and policy but requires careful handling to respect privacy and avoid misuse. Clear documentation of limitations and preprocessing steps would further support ethical and effective use.

# Distribution

1.  *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
    -   Yes, the dataset is distributed to the public through Open Data Toronto, a platform designed to share datasets collected by the City of Toronto and its agencies. It is accessible to third parties, including researchers, policymakers, and the general public.
2.  *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
    -   The dataset is distributed via the Open Data Toronto website, where it can be downloaded in formats such as CSV or accessed through APIs for integration into applications. It does not appear to have a specific Digital Object Identifier (DOI), but a permanent link is provided for access.
3.  *When will the dataset be distributed?*
    -   The dataset is already available for public access on the Open Data Toronto platform. Updates to the dataset are likely released periodically, depending on the reporting frequency and data processing timelines of the Toronto Police Service.
4.  *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
    -   Yes, the dataset is distributed under the Open Government License - Toronto, which allows users to copy, modify, publish, and distribute the data, provided proper attribution is given. There are no fees associated with accessing or using the dataset under this license. 
5.  *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
    -   No third-party IP restrictions are imposed on the dataset. However, the terms of the Open Government License require users to acknowledge the source of the data when using it in derivative works or publications.
6.  *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
    -   No export controls or regulatory restrictions apply to the dataset. It is openly available for access and use by anyone, provided they adhere to the terms of the Open Government License.
7.  *Any other comments?*
    -   The dataset’s open distribution model ensures broad accessibility, promoting transparency and enabling its use in research, policy-making, and public initiatives. However, users should remain mindful of privacy and ethical considerations when analyzing and sharing insights derived from the dataset. Providing regular updates and improving metadata (e.g., DOI assignment) could enhance its usability and traceability.

#Maintenance

1.  *Who will be supporting/hosting/maintaining the dataset?*
    -   The dataset is hosted, supported, and maintained by Open Data Toronto, a platform managed by the City of Toronto. The data originates from the Toronto Police Service, which is responsible for collecting and curating the raw data.
2.  *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
    -   Open Data Toronto provides contact information on its website for inquiries. General questions can often be directed to the City of Toronto's Open Data team via email at opendata@toronto.ca.
3.  *Is there an erratum? If so, please provide a link or other access point.*
    -   There is no specific erratum provided for this dataset. However, any issues or corrections might be addressed in updates to the dataset, and users can report discrepancies through Open Data Toronto’s feedback mechanisms.
4.  *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
    -   Yes, the dataset is periodically updated to include new data, correct errors, or modify existing entries. Updates are managed by the Toronto Police Service and communicated through the Open Data Toronto platform, typically indicated in the dataset’s metadata or version history.
5.  *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
    -   There are no explicit retention limits mentioned. However, since the dataset is anonymized and aggregated, it is unlikely to be subject to strict retention rules. Individual privacy is protected through anonymization and privacy offsets, making long-term retention less of a concern.
6.  *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
    -   Open Data Toronto does not explicitly provide access to historical versions of the dataset. Updates replace older versions, and users are encouraged to download and archive versions if historical consistency is required. Obsolescence is not formally communicated beyond the release of updated data.
7.  *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
    -   There is no formal mechanism for third-party contributions to the dataset itself. Users can create derivative works or analyses based on the data, but these are independent of the official dataset. Open Data Toronto does not currently validate or integrate external contributions.
8.  *Any other comments?*
    -   The dataset is maintained to ensure relevance and utility, but users must account for updates and potential changes in structure or format over time. Establishing clearer update policies, version control, and a mechanism for community contributions could enhance its usability and foster collaboration.

\newpage

# References

### 2024-12-03 paper/paper.qmd
Made corrections on grammar in appendix section

@@
 
-This project is motivated and guided by Rohan Alexander and his book [@tellingstories].
+This project is motivated and guided by Rohan Alexander and his book [@citeTbook].
 Data used in this paper was cleaned, analyzed and modeled with the programming language R [@citeR].
-Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `palmerpenguins` [@citePalmerpenguins], `performance` [@citePerformance], `tidygraph` [@citeTidygraph], `ggraph` [@citeGgraph], `lubridate` [@citeLubridate], `sf` [@citeSf], `osmdata` [@citeOsmdata], `car` [@citeCar],
+Also with support of additional packages in R: `readr` [@citeReadr], `ggplot2` [@citeGgplot2], `tidyverse` [@citeTidyverse], `dplyr` [@citeDplyr], `here` [@citeHere], `knitr` [@citeKnitr], `kableExtra` [@citeKableExtra], `palmerpenguins` [@citePalmerpenguins], `performance` [@citePerformance], `tidygraph` [@citeTidygraph], `ggraph` [@citeGgraph], `ggridges` [@ggridges], `lubridate` [@citeLubridate], `sf` [@citeSf], `osmdata` [@citeOsmdata], and `car` [@citeCar].
 
@@
 
-Additional details about the dataset are available in the datasheet, accessible through the repository linked to this paper.
+Additional details about the dataset are available in the [datasheet](https://github.com/ohyykk/Toronto_Motor_Viehicle/blob/main/other/datasheet/Toronto_Motor_Risk_datasheet.pdf), accessible through the repository linked to this paper.
 
@@
 ```{r}
-#| label: fig-risk-index-boxplot
+#| label: fig-risk-index-ridge
 #| echo: false
@@
 #| fig-width: 8 
-#| fig-height: 4
-
+#| fig-height: 3.5
+
+# Load necessary libraries
+library(ggridges)
 library(ggplot2)
-library(dplyr)
-
+library(here)
+library(readr)
+library(dplyr)  # For data manipulation
+
+# Load the data
 risk_index_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
 
-# Assume `risk_index_data` contains the `risk_index` values for each neighborhood
-# Group neighborhoods by average Risk Index
-grouped_data <- risk_index_data %>%
-  group_by(hood_158) %>%
-  summarize(avg_risk_index = mean(risk_index, na.rm = TRUE)) %>%
-  mutate(risk_category = case_when(
-    avg_risk_index <= quantile(avg_risk_index, 0.25) ~ "Low Risk",
-    avg_risk_index > quantile(avg_risk_index, 0.25) & avg_risk_index <= quantile(avg_risk_index, 0.75) ~ "Moderate Risk",
-    avg_risk_index > quantile(avg_risk_index, 0.75) ~ "High Risk"
-  ))
-
-# Add risk categories directly to the `risk_index_data`
+# Create risk_category based on risk_index
 risk_index_data <- risk_index_data %>%
-  left_join(grouped_data %>% select(hood_158, risk_category), by = "hood_158")
-
-# Box plot of Risk Index by Risk Category with custom colors
-ggplot(risk_index_data, aes(x = risk_category, y = risk_index, fill = risk_category)) +
-  geom_boxplot(outlier.color = "darkred", outlier.shape = 16, alpha = 0.7) +
+  mutate(
+    risk_category = case_when(
+      risk_index < 0.33 ~ "Low Risk",
+      risk_index < 0.66 ~ "Moderate Risk",
+      TRUE ~ "High Risk"
+    )
+  )
+
+# Ridge plot of Risk Index by Risk Category
+ggplot(risk_index_data, aes(x = risk_index, y = risk_category, fill = risk_category)) +
+  geom_density_ridges(alpha = 0.7, color = "black") +
   scale_fill_manual(
@@
   labs(
-    x = "Risk Category",
-    y = "Risk Index"
+    x = "Risk Index",
+    y = "Risk Category"
   ) +
@@
 
-Building on the distribution of the Risk Index in @fig-risk-index-histogram, @fig-risk-index-boxplot highlights the distribution of the Risk Index across three defined categories: High Risk, Low Risk and Moderate Risk.
+Building on the distribution of the Risk Index in @fig-risk-index-ridge highlights the density of the Risk Index across three defined categories: High Risk, Moderate Risk, and Low Risk.
 Each category represents a grouping of neighborhoods based on their average Risk Index values.
-The plot illustrates distinct differences in the distribution of the Risk Index between these categories.
-
-The High Risk category illustrates a higher median Risk Index, with a relatively narrow interquartile range (IQR), indicating consistent high-risk values across neighborhoods in this group.
-Conversely, the Low Risk category has a significantly lower median, with a similarly narrow IQR, suggesting stable low-risk conditions in these neighborhoods.
-The Moderate Risk category shows greater variability in its distribution, with a wider IQR and overlapping values with both High Risk and Low Risk categories.
-This suggests that neighborhoods in the Moderate Risk group exhibit diverse risk profiles, likely influenced by varying environmental and situational factors.
+The plot illustrates distinct patterns in the distribution of the Risk Index between these categories.
+
+The High Risk category demonstrates a single, narrow peak at relatively higher Risk Index values, indicating a concentrated group of neighborhoods with consistently elevated risk levels.
+This density suggests a homogeneity in the high-risk group, where most neighborhoods share similar risk characteristics.
+Conversely, the Low Risk category shows a bimodal distribution, with two distinct peaks at much lower Risk Index values.
+This indicates variability within this category, with some neighborhoods experiencing very low risk and others clustering near the moderate range.
+The Moderate Risk category displays a more diffuse distribution, with a broad range of Risk Index values extending into both the High Risk and Low Risk ranges.
+This variability suggests that neighborhoods in the Moderate Risk category experience diverse risk profiles, likely shaped by a combination of environmental, temporal, and situational factors.
 
@@
 #| label: tbl-time_of_day_summary
-#| tbl-cap: "5 Examples of Different Times of Day with Incident Count"
+#| tbl-cap: "Examples of Different Times of Day with Incident Count"
 #| echo: false
@@
 
-Figure @fig-linear-time-of-day illustrates the relationship between variable Time of Day and the predicted values of the collision probability model.
+@fig-linear-time-of-day illustrates the relationship between variable Time of Day and the predicted values of the collision probability model.
 The horizontal axis represents the reported time of day in minutes since midnight, while the vertical axis displays the logit values.
@@
 
-The results of the VIF analysis for the collision risk model are presented in Table @tbl-vif.
+The results of the VIF analysis for the collision risk model are presented in @tbl-vif.
 Most predictors exhibit VIF values well below the commonly used threshold of 5, indicating no significant multicollinearity among them.

### 2024-12-03 scripts/03-clean_data.R
data cleaning script final
edit data cleaning preamble

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Cleans the raw theft dataset for further analysis and includes the day of the week (`REPORT_DOW`) and saves the data in CSV format.
-# Author: [Your Name]
-# Date: [Current Date]
-# Contact: [Your Email]
+# Purpose: Cleans the raw theft and collision dataset for further analysis
+# Author: Yingke He
+# Date: 25 Nov 2024
+# Contact: kiki.he@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
 # - The `tidyverse`, `janitor`, `arrow`, and `here` packages must be installed.

### 2024-12-03 scripts/05-exploratory_data_analysis.R
model and exploratory data analysis final

@@ -1,9 +1,9 @@
 
 #### Preamble ####
 # Purpose: To analyze and visualize trends in traffic and theft data, this script includes analyses of risk index distributions, environmental predictors, temporal trends, and neighborhood-based disparities while leveraging a Bayesian model for risk analysis.
-# Author: [Your Name]
-# Date: [Today's Date]
-# Contact: [Your Email]
+# Author: Yingke He
+# Date: 25 Nov
+# Contact: kiki.he@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
 # - Required R packages installed

### 2024-12-03 scripts/06-model_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Calculate theft sub-indexes and output as CSV
 # Author: Yingke He
-# Date: [Insert Date]
-# Contact: [Insert Email]
+# Date: 25 Nov 2024
+# Contact: kiki.he@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
 # - Data cleaned using 03-clean_data.R

### 2024-12-13 paper/paper.qmd
Add link of the questionnaire to appendix

@@
 
-The primary estimand of the analysis is the composite risk score, derived from the individual probabilities of motorbike theft and collision.
+The **primary estimand** of the analysis is the composite risk score, derived from the individual probabilities of motorbike theft and collision.
 This score is calculated using predictor variables such as neighborhood characteristics, road and lighting conditions, time of day, and other contextual factors, which were selected for their documented relevance to motorbike-related risks.
@@
 #| fig-cap: "Risk Index by Lighting Conditions"
-#| fig-subcap: ["Low Light (Dark)", "Dawn Light", "Bright Daylight and Artificial Light", "Artificial Lighting Conditions"]
+#| fig-subcap: ["Dark Light Conditions", "Daylight Conditions", "Dawn Light", "Artificial Lighting Conditions"]
 #| echo: false
@@
 
-Dark Conditions in @fig-risk-index-lighting-conditions-1 highlights a concentration of Risk Index values in the moderate-to-high range, indicating that low visibility during dark conditions significantly contributes to elevated collision risks.
+Dark Light Conditions in @fig-risk-index-lighting-conditions-1 highlights a concentration of Risk Index values in the moderate-to-high range, indicating that low visibility during dark conditions significantly contributes to elevated collision risks.
 These findings underscore the importance of improved street lighting or reflective road markers in mitigating risks under dark conditions.
@@
 
-\newpage
-
 # Idealized Methodology for a Survey on Motor Risk {#sec-survey}
@@
 
-## Survey structure and Content
-
-The survey is structured into six main sections, each designed to gather detailed information about factors influencing motorbike risks.
-
-**1. Demographics**
-
-This section collects essential background information to contextualize the respondent's behaviors and risks:
-
-- Age
-- Gender
-- Education level
-- Income range
-- Geographic location (urban, suburban, rural)
-- Motorbike ownership status
-
-**2. Motorbike Usage Patterns**
-
-Questions in this section aim to understand how respondents use their motorbikes:
-
-- Frequency of use (e.g., daily, weekly, monthly)
-- Primary purpose (e.g., commuting, recreation, delivery work)
-- Average trip duration
-- Type of roads frequently used (e.g., highways, city streets, rural roads)
-- Riding seasonality (e.g., year-round, summer-only)
-
-**3. Riding Behavior**
-
-This section looks into behaviors that may influence collision and theft risks:
-
-- Average speed during rides
-- Use of safety gear (e.g., helmets, gloves, reflective clothing)
-- Adherence to traffic laws (e.g., stopping at signals, avoiding speeding)
-- Riding under adverse conditions (e.g., rain, low light)
-- Group riding versus solo riding
-
-**4. Situational Risk Factors**
-
-This section explores external factors that contribute to motorbike risks:
-
-- Time of day most frequently ridden (morning, afternoon, night)
-- Common weather conditions experienced (e.g., sunny, rainy, foggy)
-- Road surface conditions during rides (e.g., dry, wet, icy)
-- Presence of traffic control measures on roads used
-- Experience with poorly lit or visibility-compromised areas
-
-**5. Risk Awareness and Experience**
-
-This section assesses respondents' awareness of risks and their personal experiences:
-
-- Perception of collision and theft risks in their area
-- Personal experience with motorbike theft (e.g., number of thefts, recovered cases)
-- Personal experience with collisions (e.g., frequency, severity, injuries sustained)
-- Awareness of high-risk areas for theft or collisions
-- Perception of environmental risks (e.g., poor road design, lack of lighting)
-
-**6. Preventive Measures and Interventions**
-
-This section investigates actions taken by riders to mitigate risks and their opinions on potential interventions:
-
-- Security measures used to prevent theft (e.g., locks, alarms, GPS trackers)
-- Riding adjustments to avoid collisions (e.g., avoiding high-risk areas, reducing     speed)
-- Willingness to participate in safety programs (e.g., high-level riding courses)
-  Opinions on public safety measures (e.g., improved lighting, better road   
-  infrastructure
-- Suggestions for reducing theft and collisions in their area
-
-### Question Types
-
-To ensure the survey collects meaningful, diverse, and actionable data, a combination of question types will be used.
-Each type serves a specific purpose, enabling both quantitative analysis and qualitative understanding.
-Below are the key question types included in the survey:
+## Survey Structure and Content
+
+Link to the survey: [Motorbike Safety and Risk Assessment Survey](https://forms.office.com/Pages/ResponsePage.aspx?id=DQSIkWdsW0yxEjajBLZtrQAAAAAAAAAAAANAAWz9FNNUN1pHR0haWjJUSU1LTVlDVkFFVVVMQjJUTy4u)
 
@@
 These questions have only two answer choices and are typically used to gather straightforward, binary data.
-
-6. Drop-Down Questions
-
-Drop-down questions are useful for limiting the display of options while allowing respondents to choose from a broad range of answers.
-
-7. Conditional/Follow-Up Questions
-
-Conditional questions are used to customize the survey based on previous responses, ensuring relevance and efficiency.
-
-#### Question List
-
-**Demographics**
-
-1. What is your age group?
-
-- Under 18
-
-- 18-24
-
-- 25-34
-
-- 35-44
-
-- 45-54
-
-- 55+
-
-2. What is your highest level of education?
-
-- High school or less
-
-- Some college or university
-
-- Bachelor’s degree
-
-- Master’s degree
-
-- Doctorate or professional degree
-
-**Motorbike Usage Patterns**
-
-3. How frequently do you use your motorbike?
-
-- Daily
-
-- Weekly
-
-- Monthly
-
-- Rarely
-
-- Other
-
-4. How important are the following reasons for riding a motorbike to you?
-
-- Commuting
-
-- Recreation
-
-- Delivery work
-
-- Other
-
-(Rank each reason: Not important at all, Slightly important, Moderately important, Very important, Extremely important)
-
-**Riding Behavior**
-
-5. How often do you wear a helmet while riding?
-
-- Never
-
-- Rarely
-
-- Sometimes
-
-- Often
-
-- Always
-
-6. What is your average speed while riding?
-
-- Below 40 km/h
-
-- 40-60 km/h
-
-- 60-80 km/h
-
-- Above 80 km/h
-
-**Situational Risk Factors**
-
-7. What time of day do you most often ride?
-
-- Morning (6 AM–12 PM)
-
-- Afternoon (12 PM–6 PM)
-
-- Night (6 PM–12 AM)
-
-- Early morning (12 AM–6 AM)
-
-8. How often do you encounter poor road visibility during rides?
-
-- Never
-
-- Rarely
-
-- Sometimes
-
-- Often
-
-**Risk Awareness and Experience**
-
-9. Have you ever had your motorbike stolen?
-
-- Yes
-
-- No
-
-(If "Yes," follow-up: How many times has your motorbike been stolen?)
-
-- Once
-
-- 2-3 times
-
-- More than 3 times
-
-10. How would you rate the risk of collisions in your area?
-
-- Very low
-
-- Low
-
-- Moderate
-
-- High
-
-- Very high
-
-11. How important are the following factors for motorbike safety?
-
-- Helmet use
-
-- Speed limits
-
-- Visibility
-
-(Rate each factor: Not important at all, Slightly important, Moderately important, Very important, Extremely important)
-
-**Preventive Measures**
-
-12. What security measures do you currently use to prevent motorbike theft?
-(Select all that apply)
-
-- Locks
-
-- Alarms
-
-- GPS trackers
-
-- Other
-
-13. How effective do you think these changes would be in reducing motorbike collisions in your area?
-
-(Rate each factor: Not effective at all, Slightly effective, Moderately effective, Very effective, Extremely effective)
-
-- Improved road lighting
-
--Better road maintenance (e.g., fixing potholes, clearing debris)
-
-- Increased police patrols in high-risk areas
-
-- Public education on motorbike safety
-
-- Enforcing speed limits
-
-14. What changes do you think would most effectively reduce motorbike collisions in your area?
-(Open-ended)
 

### 2024-12-14 paper/paper.qmd
shortened length of the paper, adjust size of the plots

@@
 #| fig-width: 8 
-#| fig-height: 3.5
+#| fig-height: 2.5
 
@@
 #| fig-width: 8 
-#| fig-height: 3.5
+#| fig-height: 2.7
 
@@
 ```
+\newpage
 
@@
 #| message: false
+#| fig-width: 8
+#| fig-height: 3.7
 # Load necessary libraries
@@
 By incorporating this variable, the model can better account for these temporal patterns, enhancing its ability to predict theft incidents based on the day of the week.
-\newpage
 
@@
 #| fig-width: 8
-#| fig-height: 3
+#| fig-height: 2
 #| warning: false
@@
 These patterns suggest that factors such as accessibility, lack of surveillance, and the density of parked motorbikes significantly influence theft risks across different premises types.
-Incorporating this variable allows the model to account for environmental factors, improving its ability to predict theft occurrences based on location characteristics.
 
@@
 This variable enables the analysis to assess how different surface conditions contribute to motorbike collision risks.
-
-@fig-road_surface_summary illustrates the number of incidents under each road surface condition, showing that most incidents occur on dry roads, followed by wet roads.
-Conditions such as loose snow, ice, and packed snow account for a smaller proportion of incidents, reflecting their lower occurrence frequency.
-However, these different surface conditions may represent a disproportionately high risk per unit of exposure.
-This distribution underscores the importance of accounting for surface conditions in collision prevention strategies, especially during adverse weather, to enhance road safety and reduce risks.
+@fig-road_surface_summary illustrates the number of incidents associated with different road surface conditions.
+Most incidents occur on dry roads, followed by wet roads, likely due to their higher frequency of use.
+In contrast, conditions such as loose snow, ice, and packed snow account for a smaller share of incidents, possibly because they occur less frequently.
 
@@
 #| fig-width: 8
-#| fig-height: 3.5
+#| fig-height: 2.9
 
@@
 
-\newpage
-
 ```{r}
@@
 #| fig-width: 8
-#| fig-height: 3.5
+#| fig-height: 2.7
 #| 
@@
 ```
+
+\newpage
 
@@
 #| fig-width: 8
-#| fig-height: 3.5
+#| fig-height: 3
 
@@
 Visibility Conditions directly reflects environmental factors like fog, heavy rain, or low light, which impair drivers' ability to see hazards.
-Conversely, Road Conditions\` account for the physical state of the driving surface, such as wet, icy, or damaged roads, which influence vehicle handling and stopping distance.
+Conversely, Road Conditions account for the physical state of the driving surface, such as wet, icy, or damaged roads, which influence vehicle handling and stopping distance.
 Together, these variables capture complementary aspects of collision risk, ensuring that the model provides a solid assessment.
@@
 #| message: false
-#| fig-cap: "Risk Index by Time of Day, with bars showing Overall Risk Index, the red line showing Collision Risk, and green dots showing Theft Risk."
+#| fig-cap: "Risk Index by Time of Day, with bars showing Overall Risk Index, the red line showing average Collision Risk, and green dots showing average Theft Risk."
 #| fig-pos: H
@@
 
+# Load necessary libraries
+library(ggplot2)
+library(dplyr)
+library(readr)
+
 # Load necessary data
 final_risk_data <- read_csv(here("data", "02-analysis_data", "final_risk_index.csv"))
-
 cleaned_traffic_data <- read_csv(here("data", "02-analysis_data", "cleaned_traffic_data.csv"))
-
 collision_risk_scores <- read_csv(here("data", "02-analysis_data", "collision_risk_scores.csv"))
 
-# Convert 'id' to character in both datasets if necessary
+# Convert 'id' to character in all datasets if necessary
 final_risk_data$id <- as.character(final_risk_data$id)
@@
 
-# Create 'time_of_day_bin' based on 'time_of_day' in cleaned_traffic_data
-merged_data$time_of_day_bin <- case_when(
-  merged_data$time_of_day >= 0 & merged_data$time_of_day < 600 ~ "Early Morning",
-  merged_data$time_of_day >= 600 & merged_data$time_of_day < 1200 ~ "Morning",
-  merged_data$time_of_day >= 1200 & merged_data$time_of_day < 1800 ~ "Afternoon",
-  merged_data$time_of_day >= 1800 & merged_data$time_of_day < 2400 ~ "Evening"
+# Create 'time_of_day_bin' based on 'time_of_day'
+merged_data <- merged_data %>%
+  mutate(
+    time_of_day_bin = case_when(
+      time_of_day >= 0 & time_of_day < 600 ~ "Early Morning",
+      time_of_day >= 600 & time_of_day < 1200 ~ "Morning",
+      time_of_day >= 1200 & time_of_day < 1800 ~ "Afternoon",
+      time_of_day >= 1800 & time_of_day < 2400 ~ "Evening"
+    )
   )
@@
     avg_risk_index = mean(risk_index, na.rm = TRUE),
-    avg_collision_index = mean(collision_probability
-, na.rm = TRUE)
+    avg_collision_index = mean(collision_probability, na.rm = TRUE)
   )
@@
   geom_bar(aes(y = avg_risk_index), stat = "identity", fill = "#967145") +
-  geom_line(aes(y = avg_collision_index), group = 1, color = "darkred", size = 1) + # Collision index line
+  geom_line(aes(y = avg_collision_index, group = 1), color = "darkred", size = 1) + # Collision index line
   geom_text(aes(y = avg_risk_index, label = round(avg_risk_index, 2)), vjust = -0.5) +  # Add labels for risk index
   geom_text(aes(y = avg_collision_index, label = round(avg_collision_index, 2)), vjust = 1.5, color = "black") +  # Add labels for collision index
-  geom_line(aes(y = avg_theft_index * 10), color = "green", size = 1, linetype = "dashed") +  # Theft index line (scaled by 100)
+  geom_line(aes(y = avg_theft_index * 10), color = "green", size = 1, linetype = "dashed") +  # Theft index line (scaled by 10)
   geom_point(aes(y = avg_theft_index * 10), color = "green", size = 2) +  # Theft index points
+  # Label for Collision Risk
+  geom_label(
+    data = data.frame(
+      time_of_day_bin = "Evening", 
+      y = max(final_plot_data$avg_collision_index, na.rm = TRUE) * 1.05
+    ),
+    aes(x = time_of_day_bin, y = y, label = "Average Collision Risk"),
+    inherit.aes = FALSE,
+    fill = "pink",
+    alpha = 0.3,
+    color = "darkred",
+    size = 3,
+    label.size = 0
+  ) +
+  # Label for Theft Risk (adjusted position to the right)
+  geom_label(
+    data = data.frame(
+      time_of_day_bin = "Morning", 
+      y = final_plot_data$avg_theft_index[final_plot_data$time_of_day_bin == "Morning"] * 10 + 0.1
+    ),
+    aes(x = time_of_day_bin, y = y, label = "Average Theft Risk"),
+    inherit.aes = FALSE,
+    fill = "lightgreen",
+    alpha = 0.3,
+    color = "green",
+    size = 3,
+    label.size = 0,
+    hjust = 0.5  # Adjust horizontal alignment slightly
+  ) +
   labs(
-
     x = "Time of Day",
@@
 
-This time series analysis explores how the Risk Index evolves over time, offering a long-term view across the years and a more focused examination of a specific year.
+This time series analysis explores how the Risk Index evolves over time, offering a long-term view across the years and a more focused examination of the year 2020 during covid outbreaks.
 By visualizing these temporal trends, we can reveal potential seasonal effects, fluctuations, and patterns that might not be immediately apparent in a snapshot of data.
-The following section presents the trends in the Risk Index over both the full timespan (2006-2021) and the specific year 2020.
+The following section presents the trends in the Risk Index over both the full timespan (2006-2021) and the year 2020.
 
@@
 #| fig-pos: H
-#| fig-width: 8
+#| fig-width: 9
 #| fig-height: 3.5
@@
 library(lubridate)
+library(zoo)
 library(here)
@@
 
-# Convert `date` to Date format if not already
+# Convert `report_date` to Date format
 cleaned_traffic_data <- cleaned_traffic_data %>%
@@
 
-# Aggregate data by date to calculate the average Risk Index
+# Aggregate data by month to calculate the average Risk Index
 risk_trends <- cleaned_traffic_data %>%
-  group_by(date) %>%
-  summarize(avg_risk_index = mean(risk_index, na.rm = TRUE))
-
-# Create a 6-month rolling average for clearer trend
-risk_trends <- risk_trends %>%
-  arrange(date) %>%
-  mutate(rolling_avg = zoo::rollmean(avg_risk_index, 6, fill = NA, align = "center"))
+  mutate(month = floor_date(date, unit = "month")) %>%  # Aggregate at the monthly level
+  group_by(month) %>%
+  summarize(avg_risk_index = mean(risk_index, na.rm = TRUE)) %>%
+  arrange(month) %>%
+  mutate(rolling_avg = zoo::rollmean(avg_risk_index, 6, fill = NA, align = "center"))  # 6-month rolling average
 
 # Create the time series plot for the entire period
-ggplot(risk_trends, aes(x = date, y = avg_risk_index)) +
-  geom_line(color = "#967145", size = 1, alpha = 0.7) +
-  geom_line(aes(x = date, y = rolling_avg), color = "grey", size = 1, linetype = "dashed") +
+ggplot(risk_trends, aes(x = month, y = avg_risk_index)) +
+  geom_line(color = "#967145", size = 1, alpha = 0.7) +  # Original average trend
+  geom_line(aes(y = rolling_avg), color = "red", size = 1, linetype = "dashed") +  # Rolling average
+  # Add annotation for the six-month rolling average
+  annotate(
+    "label", 
+    x = as.Date("2022-06-01"),  # Adjust to place the label inside the plot area
+    y = 0.6,  # Place it at a visible y-coordinate
+    label = "Six-Month Rolling Average", 
+    color = "red", 
+    fill = "lightpink", 
+    alpha = 0.4, 
+    size = 4, 
+    label.size = 0
+  ) +
   labs(
-
     x = "Year",
@@
     date_labels = "%Y"
+  ) +
+  scale_y_continuous(
+    limits = c(0.52, 0.62),  # Ensure the label fits within the y-axis range
+    breaks = seq(0.52, 0.62, by = 0.02)
   ) +
@@
 
-@fig-risk-index-timeseries-longterm illustrates the long-term temporal trends in the Average Risk Index, measured daily and visualized from mid 2006 to mid 2023.
-The solid line represents the daily Average Risk Index, while the grey dashed line shows a six-month rolling average, providing a smoothed representation of the underlying trends.
-
-The daily Average Risk Index illustrates considerable variability, reflecting fluctuations in risk factors such as collision and theft incidents over time.
-These short-term variations may be influenced by factors like weather conditions, traffic density, or seasonal changes.
-In contrast, the rolling average highlights more stable, long-term patterns, suggesting relatively stable long-term trends in the Risk Index, with no sudden increases or decreases over the years.
-However, several periodic dips and peaks in late 2016 and mid 2018 underscore the importance of addressing short-term change in risk to enhance overall safety.
+@fig-risk-index-timeseries-longterm depicts the long-term trends in the Average Risk Index, measured daily and displayed from mid-2006 to mid-2023.
+The brown solid line represents the daily Average Risk Index, while the red dashed line illustrates a six-month rolling average, providing a smoother depiction of overall patterns.
+
+The daily Average Risk Index fluctuates significantly, reflecting changes in factors such as collision and theft occurrences over time.
+These short-term variations could be associated with seasonal patterns, weather changes, or traffic conditions.
+Meanwhile, the rolling average highlights broader, more consistent patterns, indicating stable long-term trends without dramatic increases or decreases.
+Periodic dips and peaks, such as those observed in late 2016 and mid-2018, emphasize the importance of addressing temporary shifts in risk to support safer conditions.
 
@@
 These findings highlight the significant impact that road surface conditions can have on traffic-related risks, underscoring the importance of considering these factors in risk management and safety measures.
-\newpage
 
@@
 library(ggplot2)
-
 library(osmdata)
@@
   scale_color_manual(values = c("High Collision Risk Neighborhoods" = "red", "Others" = "yellow")) +
+  # Add the arrow with adjusted position
+  geom_segment(
+    aes(x = -79.38, y = 43.61, xend = -79.38, yend = 43.63),  # Move arrow to the right
+    arrow = arrow(length = unit(0.2, "cm")),
+    color = "#363231"
+  ) +
+  # Add the label with adjusted position
+  annotate(
+    "label",
+    x = -79.38,  # Move label to the right
+    y = 43.60,
+    label = "Downtown Area",
+    fill = "lightgrey",
+    color = "#363231",
+    alpha = 0.7,
+    size = 4,
+    label.size = 0
+  ) +
   labs(
@@
 # Appendix {.unnumbered}
+
+# Shiny Application
 
@@
 
-## Predictor Variable Choice
-
-### Theft Index
-
-In constructing the theft risk sub-indexes, the choice of predictor variables: Hour of the Day, Day of the Week, and Premises Type was informed by both theoretical and practical considerations rooted in spatial, temporal, and environmental factors associated with theft incidents.
-These variables were selected from the cleaned theft dataset to create a strong and interpretable model that captures key patterns of theft occurrences while aligning with the study's goal of identifying targeted interventions.
-
-1. Hour of the Day
-
-The inclusion of the Hour of the Day as a predictor variable reflects the strong temporal dependency of theft occurrences.
-Theft risk often varies significantly across a 24-hour period due to human activity patterns, such as peak commuting times, work hours, and late-night inactivity.
-For example, motor vehicle thefts are more likely to occur during nighttime hours when there is reduced surveillance and fewer bystanders to deter criminal behavior.
-By incorporating the Hour of the Day into the theft risk sub-index, the model accounts for this temporal dimension, enabling stakeholders to identify high-risk hours and implement time-specific interventions, such as increased nighttime patrols or surveillance in parking lots and residential areas.
-
-2. Day of the Week
-
-The Day of the Week was selected to capture the variations in theft risk across different days, which often reflect behavioral patterns linked to work, leisure, and mobility.
-Weekends, for instance, may see heightened theft activity in areas with nightlife or during events that attract large crowds, while weekdays might exhibit theft patterns associated with commuter parking zones or commercial districts.
-The inclusion of this variable allows the model to identify day-specific risks and tailor interventions, such as targeted surveillance during high-risk days or public awareness campaigns about theft prevention.
-
-3. Premises Type
-
-Premises Type represents the spatial dimension of theft risk, capturing the environmental characteristics where thefts are most likely to occur.
-For example, thefts in residential areas (e.g., houses, apartments) often occur when vehicles are parked in driveways or on the street with limited security measures.
-Similarly, commercial areas and outdoor public spaces may pose unique risks due to their high foot traffic, limited oversight during certain hours, or accessibility of parked vehicles.
-Including Premises Type in the sub-index allows the model to differentiate risk levels based on spatial characteristics and recommend location-specific strategies, such as installing better lighting in outdoor parking areas or promoting secure storage options for residential settings.
-
-### Collision Probability
-
-The predictors chosen for the logistic regression model were carefully selected to capture key spatial, temporal, and environmental factors that influence the likelihood of severe collisions.
-These variables were derived from the cleaned collision dataset, ensuring data quality and relevance to the study's focus on urban mobility and safety in Toronto.
-
-1. Neighborhood ID
-
-The inclusion of hood_158, a unique identifier for neighborhoods, allows the model to capture spatial patterns in collision occurrences.
-Different neighborhoods may exhibit varying levels of collision risk due to factors such as traffic density, road design, infrastructure quality, and socioeconomic conditions.
-By including this variable, the model can account for geographic heterogeneity, identifying high-risk areas and enabling targeted interventions such as infrastructure improvements or traffic calming measures in specific neighborhoods.
-
-2. Road Surface Condition
-
-Road surface conditions, such as dry, wet, or icy surfaces, significantly impact vehicle control and braking effectiveness, making this variable a critical predictor.
-Collisions are more likely under adverse surface conditions, such as rain, snow, or ice, due to reduced traction and visibility.
-Including this variable in the model ensures that environmental factors influencing vehicle dynamics are accounted for, providing insights into how weather and road maintenance contribute to collision risk.
-This information can guide policies such as improved road drainage systems or timely snow and ice removal.
-
-3. Lighting Condition
-
-Lighting conditions, which describe whether collisions occur during daylight, nighttime, or under artificial light, are essential for understanding the role of visibility in collisions.
-Poor lighting can reduce drivers' ability to perceive hazards, increasing the likelihood of accidents.
-By incorporating this variable, the model can assess how lighting contributes to collision risk, highlighting the importance of enhanced street lighting in poorly lit areas or during nighttime hours.
-\newpage
-
-4. Traffic Control
-
-The presence and type of traffic control devices, such as stop signs, traffic signals, and pedestrian crossovers, play a important role in preventing collisions by regulating vehicle flow and ensuring predictable driver behavior.
-The inclusion of this variable allows the model to evaluate the effectiveness of existing traffic control measures and identify gaps, such as intersections without adequate control devices, where collisions are more likely to occur.
-This idea can inform infrastructure upgrades and the installation of additional traffic management systems.
-
-5. Road Visibility
-
-Road visibility conditions, influenced by factors like fog, rain, or snow, are directly linked to drivers' ability to perceive their surroundings and react to potential hazards.
-Poor visibility increases the risk of collisions by reducing reaction time and decision-making effectiveness.
-Including this variable ensures that the model captures the impact of adverse weather and environmental conditions on collision likelihood, emphasizing the need for weather-responsive traffic management strategies, such as variable speed limits or enhanced signage.
-
-6. Time of Day
-
-The Time of Day is a key temporal variable that captures variations in collision risk throughout a 24-hour period.
-Collision likelihood often peaks during specific hours, such as morning and evening rush hours, due to increased traffic volume, or late at night, when factors like fatigue or impaired driving may come into play.
-Including this variable allows the model to account for temporal patterns in collision occurrences, supporting the implementation of time-specific measures, such as increased traffic enforcement during high-risk hours or adjustments to traffic light timings.
-
 # Idealized Methodology for a Survey on Motor Risk {#sec-survey}

### 2024-12-15 other/shiny/app.R
update files for shiny app

@@ -1,0 +1,173 @@
+# Install necessary packages
+if (!require("shiny")) install.packages("shiny")
+if (!require("leaflet")) install.packages("leaflet")
+if (!require("geojsonsf")) install.packages("geojsonsf")
+if (!require("dplyr")) install.packages("dplyr")
+if (!require("sf")) install.packages("sf")
+if (!require("here")) install.packages("here")
+
+library(shiny)
+library(leaflet)
+library(geojsonsf)
+library(dplyr)
+library(sf)
+library(here)
+
+# Read and process the data using here() for proper path resolution
+# 1. Read GeoJSON data and convert to SF object
+collision_data <- st_read(here("data", "02-analysis_data", "collision.geojson"))
+
+# 2. Read risk index data
+risk_data <- read.csv(here("data", "02-analysis_data", "final_risk_index.csv")) %>%
+  distinct(hood_158, risk_index) %>%  # Remove duplicates to get unique hood/risk pairs
+  filter(!is.na(risk_index))  # Remove NA values
+
+# 3. Create neighborhood centroids from collision data
+neighborhood_data <- collision_data %>%
+  group_by(HOOD_158, NEIGHBOURHOOD_158) %>%
+  summarise(
+    collision_count = n(),
+    # Convert to centroid, ensuring point geometry
+    geometry = st_centroid(st_union(geometry))
+  ) %>%
+  ungroup() %>%
+  # Ensure geometry is properly set
+  st_cast("POINT")
+
+# 4. Merge with risk data
+merged_data <- neighborhood_data %>%
+  left_join(risk_data, by = c("HOOD_158" = "hood_158")) %>%
+  filter(!is.na(risk_index))  # Remove any remaining NA values
+
+# Define UI with background image, dark brown title, and Georgia font
+ui <- fluidPage(
+  tags$head(
+    tags$style(HTML("
+      body {
+        background-image: url('1.jpg');
+        background-size: cover;
+        background-repeat: no-repeat;
+        background-attachment: fixed;
+        font-family: Georgia, serif;
+      }
+      .container-fluid {
+        background-color: rgba(255, 255, 255, 0.3);  /* Reduced opacity from 0.9 to 0.6 */
+        padding: 20px;
+        border-radius: 10px;
+        font-family: Georgia, serif;
+      }
+      h2 {
+        color: black;
+        font-weight: bold;
+        font-family: Georgia, serif;
+      }
+      .selectize-input, .selectize-dropdown, .slider-animate-container {
[…]

### 2024-12-15 other/shiny/app.R
app finalized

@@ -109,22 +109,37 @@ ui <- fluidPage(
 # Define server
 server <- function(input, output, session) {
   
-  # Create filtered reactive dataset
+  # Create filtered reactive dataset with error handling
   filtered_data <- reactive({
+    req(input$risk_range)  # Ensure risk range input exists
+    
     data <- merged_data
     
+    # Handle neighborhood selection with validation
     if (!is.null(input$neighborhood) && length(input$neighborhood) > 0) {
-      data <- data[data$NEIGHBOURHOOD_158 %in% input$neighborhood, ]
+      data <- data %>%
+        filter(NEIGHBOURHOOD_158 %in% input$neighborhood)
+      
+      # Validate that we have data after filtering
+      if (nrow(data) == 0) {
+        return(merged_data)  # Return all data if filter results in empty set
+      }
     }
     
-    data <- data[data$risk_index >= input$risk_range[1] & 
-                 data$risk_index <= input$risk_range[2], ]
+    # Filter by risk range with validation
+    data <- data %>%
+      filter(risk_index >= input$risk_range[1],
+             risk_index <= input$risk_range[2])
     
     return(data)
   })
   
-  # Create the map
+  # Create the map with error handling
   output$riskMap <- renderLeaflet({
+    # Validate that we have data to show
+    data <- filtered_data()
+    req(nrow(data) > 0)  # Ensure we have data
+    
     # Create color palette
     pal <- colorNumeric(
       palette = "YlOrRd",
@@ -132,13 +147,12 @@ server <- function(input, output, session) {
     )
     
     # Get coordinates for filtered data
-    data <- filtered_data()
     coords <- st_coordinates(data)
     
     # Create base map
     leaflet(data) %>%
       addTiles() %>%
-      setView(lng = -79.3832, lat = 43.6532, zoom = 11) %>%  # Toronto coordinates
+      setView(lng = -79.3832, lat = 43.6532, zoom = 11) %>%
       addCircleMarkers(
         lng = coords[,1],
         lat = coords[,2],

### 2024-12-15 other/datasheet/Toronto_Motor_Risk_datasheet.qmd
datasheet finalized

@@
 4.  *Any other comments?*
-    -   The datasets serve as critical tools for urban planning, public safety strategies, and academic research.
+    -   The datasets serve as important tools for urban planning, public safety strategies, and academic research.
 However, users should note the limitations of the dataset, including potential discrepancies due to privacy adjustments and the lack of guarantees for completeness or timeliness.
-The initiative reflects broader efforts to leverage open data for improving community outcomes and fostering collaborative problem-solving.
+The initiative reflects broader efforts to utilize open data to enhance community outcomes and promote collaborative problem-solving.
 
@@
 16. *Any other comments?*
-    -   The dataset serves as an invaluable resource for public safety analysis and urban planning, but users must approach it with care, especially regarding privacy and ethical considerations.
+    -   The dataset serves as an important resource for public safety analysis and urban planning, but users must approach it with care, especially regarding privacy and ethical considerations.
 While efforts have been made to anonymize and adjust sensitive information, the potential for misuse underscores the importance of responsible data handling and interpretation.
@@
 3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-    -   The dataset is a comprehensive sample of reported incidents rather than a probabilistic or random sample.
+    -   The dataset represents an extensive collection of reported incidents rather than a probabilistic or random sample.
 It includes all documented collisions and thefts meeting specific criteria (e.g., collisions resulting in fatalities or serious injuries).
@@
 12. *Any other comments?*
-    -   While the dataset provides valuable insights for public safety and urban planning, a more detailed impact analysis could address potential unintended consequences, such as misrepresentation of neighborhoods or misuse of sensitive information.
+    -   While the dataset provides important ideas for public safety and urban planning, a more detailed impact analysis could address potential unintended consequences, such as misrepresentation of neighborhoods or misuse of sensitive information.
 Documentation of such an analysis would enhance confidence in the dataset’s ethical use.
@@
 6.  *Any other comments?*
-    -   This dataset is a valuable resource for research and policy but requires careful handling to respect privacy and avoid misuse.
+    -   This dataset is a important resource for research and policy but requires careful handling to respect privacy and avoid misuse.
 Clear documentation of limitations and preprocessing steps would further support ethical and effective use.
@@
     -   The dataset’s open distribution model ensures broad accessibility, promoting transparency and enabling its use in research, policy-making, and public initiatives.
-However, users should remain mindful of privacy and ethical considerations when analyzing and sharing insights derived from the dataset.
+However, users should remain mindful of privacy and ethical considerations when analyzing and sharing ideas derived from the dataset.
 Providing regular updates and improving metadata (e.g., DOI assignment) could enhance its usability and traceability.

### 2024-12-15 paper/paper.qmd
added shiny app
paper finalized
made minor change in wordings

@@
 By combining these elements, the Risk Index provides a unified measure of risk, enabling the identification of high-risk scenarios and areas.
+An interactive visualization of the final risk index can be found in [Appendix -@sec-shiny].
 
@@
 This design ensures that the model reflects the specific data characteristics and practical considerations when assessing motorbike-related risks.
-Logistic regression is not utilized for theft risk due to the absence of negative (non-theft) cases in the dataset.
+Logistic regression is **not utilized** for theft risk due to the absence of negative (non-theft) cases in the dataset.
 Instead, theft risk is represented through sub-index calculations for three critical factors: hour of the day, day of the week, and premises type.
@@
 
-The final Risk Index integrates the theft and collision components using a weighted formula.
-The collision probability component is weighted at 0.7, reflecting its higher immediate impact on safety, while the theft component is weighted at 0.3.
-This weighting scheme prioritizes collision risk while ensuring that theft risk is not overlooked.
-By combining these components, the Risk Index provides a unified measure of motorbike-related risks, enabling stakeholders to assess and compare safety conditions across different contexts.
+The final Risk Index integrates the theft and collision components using a weighted scheme of 0.3 for theft and 0.7 for collision, emphasizing collision risk while ensuring theft risk is considered.
+By combining these components, the Risk Index offers a unified measure of motorbike-related risks, enabling stakeholders to evaluate and compare safety conditions across different contexts.
+
+The model assumes independence between theft and collision risks and relies on proportional theft representations, which may limit its ability to capture interactions or account for unmeasured confounders.
+**Alternative methods** like generalized additive models or machine learning techniques could address these limitations but were not used to prioritize interpretability and simplicity.
 
 This modeling approach is justified by its ability to adapt to the data’s constraints while maintaining statistical rigor and interpretability.
-The sub-index method for theft risk is tailored to the dataset’s characteristics, avoiding assumptions about unobserved cases, and the use of logistic regression for collision risk ensures robust and reliable predictions.
-Together, these components form a practical and practical framework for evaluating motorbike ownership and usage risks, addressing both immediate safety concerns and long-term theft risks.
+The sub-index approach for theft risk is designed to align with the characteristics of the dataset, avoiding assumptions about unobserved cases, and the use of logistic regression for collision risk ensures solid and reliable predictions.
+Together, these components form a practical framework for evaluating motorbike ownership and usage risks, addressing both immediate safety concerns and long-term theft risks.
 
@@
 
-to evaluate the robustness and reliability of the final Risk Index.
+to evaluate the reliability of the final Risk Index.
 
@@
 The observations are not repeated or correlated, ensuring that the logistic regression model provides unbiased estimates of the relationships between predictor variables and the response variable.
-By meeting this assumption, the collision model maintains its validity for estimating probabilities of severe collisions and contributes robustly to the composite Risk Index.
+By meeting this assumption, the collision model remains valid for estimating probabilities of severe collisions and provides a reliable contribution to the composite Risk Index.
 
@@
 
-# Shiny Application
+# Shiny Application {#sec-shiny}
+
+The Risk Index by neighborhood can be visualized [here](https://ohyykk.shinyapps.io/TorontoMotorVehicle/).
 

### 2024-12-15 paper/paper.qmd
finalized paper

@@
 ---
-title: "When Risk Hits the Road: Does Danger Drive a Return?"
+title: "Motor Vehicle Risks in Toronto: Uncovering Collision and Theft Patterns"
 subtitle: "Risk Peaks in the Morning on Dark, Snow-Packed Roads and in High Urban Density Areas"

### 2024-12-15 other/shiny/app.R
add preambles

@@ -1,10 +1,9 @@
-# Install necessary packages
-if (!require("shiny")) install.packages("shiny")
-if (!require("leaflet")) install.packages("leaflet")
-if (!require("geojsonsf")) install.packages("geojsonsf")
-if (!require("dplyr")) install.packages("dplyr")
-if (!require("sf")) install.packages("sf")
-if (!require("here")) install.packages("here")
+# Purpose: Creates shiny app as a companion to the paper
+# Author: Yingke He
+# Date: 14 Dec 2024
+# Contact: kiki.he@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Have downloaded all the required libraries including shiny, leaflet and geojsonsf.
 
 library(shiny)
 library(leaflet)

### 2024-12-19 other/datasheet/Toronto_Motor_Risk_datasheet.qmd
datasheet fix
minor error on title

@@
 title: "Datasheet for 'Toronto Motorbike Collision and Theft Risk'"
-subtitle: "Supplementary document to \"When Risk Hits the Road: Does Danger Drive a Return?\""
+subtitle: "Supplementary document to \"Motor Vehicle Risks in Toronto: Uncovering Collision and Theft Patterns\""
 author: 
@@
 thanks: "Code and data are available at: [https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main](https://github.com/ohyykk/Toronto_Motor_Viehicle/tree/main)."
-date: Today
-date-format: "D MMMM YYYY"
+date: today 
 abstract: "This supplementary datasheet accompanies the motorbike risk analysis paper.
